---
title: "ISGlobal charts"
author: "www.databrew.cc"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: "hide"
---


```{r setup, include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               # echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.width = 8.64,
               fig.height = 4.86,
               fig.path = 'figures/')
```

```{r}
## Load libraries
library(covid19)
library(ggplot2)
library(lubridate)
library(dplyr)
library(ggplot2)

# Define vector of countries
countries <- c('China','Korea, South',
               'Italy',
               'Spain',
               'UK',
               'Germany',
               'South Africa',
               'Ghana',
               'Kenya')
```

## All countries

### Cumulative cases

#### "Start date" of 150 cases

Logarithmic: 

```{r}
plot_day_zero(countries = countries, day0 = 150)
```

Linear: 

```{r}
plot_day_zero(countries = countries, day0 = 150, ylog = F)
```


#### "Start date" of 25 cases

Logarithmic:

```{r}
plot_day_zero(countries = countries, day0 = 25)
```

Linear:

```{r}
plot_day_zero(countries = countries, day0 = 25, ylog = F)
```





### Cumulative deaths


#### "Start date" of 50 deaths

Logarithmic: 

```{r}
plot_day_zero(countries = countries, day0 = 50, deaths = T)
```

Linear: 

```{r}
plot_day_zero(countries = countries, day0 = 50, ylog = F, deaths = T)
```


#### "Start date" of 15 deaths

Logarithmic: 

```{r}
plot_day_zero(countries = countries, day0 = 15, deaths = T)
```

Linear: 

```{r}
plot_day_zero(countries = countries, day0 = 15, ylog = F, deaths = T)
```


#### "Start date" of 5 deaths

Logarithmic:

```{r}
plot_day_zero(countries = countries, day0 = 25, 5, deaths = T)
```

Linear:

```{r}
plot_day_zero(countries = countries, day0 = 5, ylog = F, deaths = T)
```



## Spain's CCAAs

### Cumulative cases

#### "Start date" of 50 cases

Logarithmic:

```{r}
plot_day_zero_esp(ccaa = sort(unique(esp_df$ccaa)),
                  day0 = 50)
```

Linear:

```{r}
plot_day_zero_esp(ccaa = sort(unique(esp_df$ccaa)),
                  day0 = 50,
                  ylog = F)
```


### Cumulative deaths

#### "Start date" of 5 deaths

Logarithmic:

```{r}
plot_day_zero_esp(ccaa = sort(unique(esp_df$ccaa)),
                  day0 = 5,
                  deaths = T)
```

Linear:

```{r}
plot_day_zero_esp(ccaa = sort(unique(esp_df$ccaa)),
                  day0 = 5,
                  deaths = T,
                  ylog = F)
```

## CCAAs, deaths adjusted by population

```{r, fig.height = 6.5, fig.width = 9}
adj <- esp_df %>% filter(date == max(date))
pop <- covid19::esp_pop
joined <- left_join(adj, pop)
# make vars
joined <- joined %>%
  mutate(cases_per_k = (cases / pop) * 100000,
         deaths_per_k = (deaths / pop) * 100000)

joined$value <- joined$deaths_per_k
joined <- joined %>% arrange(value)
joined$p <- joined$value / sum(joined$value) * 100
# joined$ccaa <- factor(joined$ccaa, 
#                       levels = joined$ccaa)
text_size <- 10
ggplot(data = joined,
       aes(x = ccaa,
           y = value)) + 
  geom_bar(stat = 'identity',
           fill = 'black') +
  theme_simple() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = '',
       y = 'Deaths',
       title = 'COVID-19 deaths in Spain (per 100,000)',
       # subtitle = 'Data as of evening of Sunday, March 15',
       caption = 'github.com/databrew/covid19 | joe@databrew.cc') +
  theme(legend.position = 'top',
        legend.text = element_text(size = text_size * 2),
        axis.title = element_text(size = text_size * 2),
        plot.title = element_text(size = text_size * 2.3),
        axis.text.x = element_text(size = text_size * 1.5)) +
  geom_text(data = joined,
            aes(x = ccaa,
                y = value,
                label = paste0(deaths, ' deaths\n(',
                               round(value, digits = 1), ' per\n100,000)')),
                # label = paste0(deaths, '\n(',
                #                round(p, digits = 1), '%)')),
            size = text_size * 0.25,
            nudge_y = 0.35)
# ggsave('~/Desktop/spain.png')
```



## CCAAs, cases adjusted by population

```{r, fig.height = 6.5, fig.width = 9}
adj <- esp_df %>% filter(date == max(date))
pop <- covid19::esp_pop
joined <- left_join(adj, pop)
# make vars
joined <- joined %>%
  mutate(cases_per_k = (cases / pop) * 100000,
         deaths_per_k = (deaths / pop) * 100000)

joined$value <- joined$cases_per_k
joined <- joined %>% arrange(value)
joined$p <- joined$value / sum(joined$value) * 100
# joined$ccaa <- factor(joined$ccaa, 
#                       levels = joined$ccaa)
text_size <- 10
ggplot(data = joined,
       aes(x = ccaa,
           y = value)) + 
  geom_bar(stat = 'identity',
           fill = 'black') +
  theme_simple() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = '',
       y = 'Cases',
       title = 'COVID-19 cases in Spain (per 100,000 people)',
       # subtitle = 'Data as of evening of Sunday, March 15',
       caption = 'github.com/databrew/covid19 | joe@databrew.cc') +
  theme(legend.position = 'top',
        legend.text = element_text(size = text_size * 2),
        axis.title = element_text(size = text_size * 2),
        plot.title = element_text(size = text_size * 2.3),
        axis.text.x = element_text(size = text_size * 1.5)) +
  geom_text(data = joined,
            aes(x = ccaa,
                y = value,
                label = paste0(cases, ' cases\n(',
                               round(value, digits = 1), ' per\n100,000)')),
                # label = paste0(deaths, '\n(',
                #                round(p, digits = 1), '%)')),
            size = text_size * 0.25,
            nudge_y = 12)
# ggsave('~/Desktop/spain.png')
```



## CCAAs, UCI by population

```{r, fig.height = 6.5, fig.width = 9}
adj <- esp_df %>% filter(date == max(date))
pop <- covid19::esp_pop
joined <- left_join(adj, pop)
# make vars
joined <- joined %>%
  mutate(cases_per_k = (cases / pop) * 100000,
         deaths_per_k = (deaths / pop) * 100000,
         uci_per_k = (uci / pop) * 100000)

joined$value <- joined$uci_per_k
joined <- joined %>% arrange(value)
joined$p <- joined$value / sum(joined$value) * 100
# joined$ccaa <- factor(joined$ccaa, 
#                       levels = joined$ccaa)
text_size <- 10
ggplot(data = joined,
       aes(x = ccaa,
           y = value)) + 
  geom_bar(stat = 'identity',
           fill = 'black') +
  theme_simple() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = '',
       y = 'Cases',
       title = 'COVID-19: intensive care patients per 100,000',
       # subtitle = 'Data as of evening of Sunday, March 15',
       caption = 'github.com/databrew/covid19 | joe@databrew.cc') +
  theme(legend.position = 'top',
        legend.text = element_text(size = text_size * 2),
        axis.title = element_text(size = text_size * 2),
        plot.title = element_text(size = text_size * 2.3),
        axis.text.x = element_text(size = text_size * 1.5)) +
  geom_text(data = joined,
            aes(x = ccaa,
                y = value,
                label = paste0(uci, '\nadmissions per\n(',
                               round(value, digits = 1), ' per\n100,000)')),
                # label = paste0(deaths, '\n(',
                #                round(p, digits = 1), '%)')),
            size = text_size * 0.25,
            nudge_y = 0.4)
# ggsave('~/Desktop/spain.png')
```

## Spain and Italy

### National comparison

Start date of 150 cases, cases, logarithmic scale

```{r}
plot_day_zero(countries = c('Spain', 'Italy'))
```


Start date of 150 cases, cases, linear scale

```{r}
plot_day_zero(countries = c('Spain', 'Italy'), ylog = F)
```

Start date of 50 deaths, deahs, logarithmic scale

```{r}
plot_day_zero(countries = c('Spain', 'Italy'), deaths = T, day0 = 50)
```


Start date of 150 cases, cases, linear scale

```{r}
plot_day_zero(countries = c('Spain', 'Italy'), deaths = T, day0  = 50, ylog = F)
```


#### Spain

Ratio of deaths to cases

```{r}
pd <- esp_df %>%
  filter(date == '2020-03-15') %>%
  dplyr::select(ccaa, deaths, cases) %>%
  mutate(deaths_per_case = deaths / cases * 1000)
text_size = 10
ggplot(data = pd,
       aes(x = ccaa,
           y = deaths_per_case)) +
    geom_bar(stat = 'identity',
           fill = 'black') +
  theme_simple() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = '',
       y = 'Deaths per 1.000 confirmed cases',
       title = 'COVID-19: deaths per 1.000 cases',
       # subtitle = 'Data as of evening of Sunday, March 15',
       caption = 'github.com/databrew/covid19 | joe@databrew.cc') +
  theme(legend.position = 'top',
        legend.text = element_text(size = text_size * 2),
        axis.title = element_text(size = text_size * 2),
        plot.title = element_text(size = text_size * 2.3),
        axis.text.x = element_text(size = text_size * 1.5)) +
  geom_text(data = pd,
            aes(x = ccaa,
                y = deaths_per_case,
                label = paste0(deaths, ' deaths\n(', round(deaths_per_case, digits = 1), 
                                ' per\n1.000)')),
                # label = paste0(deaths, '\n(',
                #                round(p, digits = 1), '%)')),
            size = text_size * 0.35,
            nudge_y = 5)
```

