---
title: "All charts"
author: "www.databrew.cc"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: "hide"
---


```{r setup, include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               # echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.width = 9.64,
               fig.height = 5.9,
               fig.path = 'figures/')
```

```{r}
## Load libraries
library(covid19)
library(ggplot2)
library(lubridate)
library(dplyr)
library(ggplot2)
library(sp)
library(raster)
library(viridis)
library(ggthemes)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer)
library(readr)
library(zoo)
library(tidyr)
options(scipen = '999')
```

```{r}
# Matrix for Aleix Prat

# All comarques
pd <- muni %>% ungroup %>%
  group_by(date, ComarcaDescripcio) %>%
  summarise(new_cases = sum(confirmed_cases_non_cum),
            pop = sum(pop)) %>%
  ungroup %>%
  arrange(date) %>%
  group_by(ComarcaDescripcio) %>%
  mutate(cases_roll = zoo::rollmeanr(new_cases, k = 7, fill = NA)) %>%
  mutate(yp = new_cases / pop * 100000,
         ypr = cases_roll / pop * 100000) %>%
  mutate(ypl = sqrt(yp),
         yprl = sqrt(ypl))

# Make wide matrix
pdw <- pd %>% dplyr::select(date, ComarcaDescripcio, yp) %>%
  tidyr::spread(key = date, value = yp)
write_csv(pdw, 'aleix_prat.csv')



# Create weekly (instead of pd)
weekly <- pd %>%
   mutate(week = lubridate::week(date)) %>%
  group_by(week, ComarcaDescripcio) %>%
  mutate(week_start = min(date)) %>%
  mutate(n_days = length(unique(date))) 

ggplot(data = pd %>% filter(date >= '2020-03-11',
                            date <= '2020-08-20'),
       aes(x = date,
           y = ComarcaDescripcio,
           fill = ypl)) +
  geom_tile(
    # color = 'black',size = 0.2
    ) +
  databrew::theme_simple() +
  scale_fill_gradientn(name = '', colors = colorRampPalette(c('beige',
                                                              # 'grey',
                                                              # 'yellow',
                                                              # 'white',
                                                              # 'orange',
                                                              'darkorange', 
                                                              'red',
                                                              'darkred',
                                                              'black'))(20),
                       breaks = labeler, labels = round(labeler^2)) +
  labs(x = '',
       y = '',
       title = 'Casos confirmats nous (per 100.000 habitants)',
       caption = 'Aix-x: dia; Aix-y: comarca; Color: casos nous.\nDades de https://analisi.transparenciacatalunya.cat/. Gràfic: @joethebrew') +
  theme(axis.text.y = element_text(size = 10))
ggsave('~/Desktop/daily.png', width = 10, height = 6)

```

```{r}
covid19::plot_day_zero(ylog = F,
                       day0 = 1, 
                       by_district = TRUE,
                       calendar = TRUE,
                       pop = TRUE,
                       roll = 7,
                       alpha = 0.7,
                       point_size = 0.5,
                       point_alpha =0.5,
                       countries = c('Spain', 'Italy', 'US'),
                       districts = c('Cataluña',  'Aragón', 'Madrid',  'Lombardia', 'Florida', 'New York', 'Emilia-Romagna'),
                       cumulative = FALSE) +
  labs(x = 'Date',
       y = 'Cases per million',
       title = 'Confirmed Covid-19 cases per 1,000,000 population\nRolling average of previous 7 days',
       subtitle = 'Data from Italy / USA as of Aug 21 2020; Data from Spain as of Aug 16 2020') +
  scale_color_colorblind(name = '',
                         guide = guide_legend(reverse = TRUE))
ggsave('~/Desktop/plot.png')
```

```{r}
# US vs Spain

pd <- owid %>%
  filter(location %in% c('United States', 'Spain', 'Italy')) %>%
  mutate(location = ifelse(location == 'United States', 'USA', location))
ggplot(data = pd,
       aes(x = date,
           y = new_cases_smoothed_per_million,
           color = location,
           shape = location)) +
  geom_line(size = 1) +
  geom_point(data = pd %>% dplyr::filter(weekdays(date) == 'Friday') , size = 3) +
  databrew::theme_simple() +
  labs(x = 'Date',
       y = 'Cases per million (smoothed)',
       title = 'Daily cases per million population',
       subtitle = '7-day rolling average',
       caption = 'Chart: @joethebrew. Data source: Our World in Data.\nhttps://github.com/owid/covid-19-data/tree/master/public/data') +
  theme(legend.position = 'right',
        legend.title = element_blank()) +
  scale_color_colorblind(name = '',
                         guide = guide_legend(reverse = TRUE))
  # scale_color_manual(name = '',
  #                    values = c('black', 'red', 'blue'),
  #                    guide = guide_legend(reverse = TRUE))
ggsave('~/Desktop/comparsion.png', height = 4, width = 7.5)
```

```{r}
# ECDC
pd <- ecdc %>%
  mutate(date = as.Date(dateRep, format = '%d/%m/%Y')) %>%
  mutate(cum14 = Cumulative_number_for_14_days_of_COVID.19_cases_per_100000) %>%
  mutate(country = countriesAndTerritories) %>%
  arrange(date) %>%
  filter(popData2019 >= 100000,
         continentExp == 'Europe') %>%
  group_by(country) %>%
  mutate(deaths14 = zoo::rollsumr(deaths, 14, fill = NA)) %>%
  ungroup %>%
  mutate(deathsp = deaths14 / popData2019 * 100000) %>%
  filter(date == '2020-08-19') %>%
  arrange(desc(cum14)) 
pd[1:5,]


pd <- ecdc %>%
  mutate(date = as.Date(dateRep, format = '%d/%m/%Y')) %>%
  mutate(cum14 = Cumulative_number_for_14_days_of_COVID.19_cases_per_100000) %>%
  mutate(country = countriesAndTerritories) %>%
  arrange(date) %>%
  filter(popData2019 >= 1000000) %>%
  filter(continentExp == 'Europe') %>%
  group_by(country) %>%
  mutate(deaths14 = zoo::rollsumr(deaths, 14, fill = NA)) %>%
  ungroup %>%
  mutate(deathsp = deaths14 / popData2019 * 100000) %>%
  arrange(desc(cum14)) %>%
  filter(country %in% c('Spain', 'Italy')) 
  # mutate(spain = ifelse(country == 'Spain', 'Spain',
  #                       'Others'))

ggplot(data = pd,
       aes(x = date,
           y = cum14,
           color = country,
           group = country)) +
  geom_line(alpha = 0.7) + 
  # geom_line(data = pd %>% filter(country == 'Spain'),
  #      aes(x = date,
  #          y = cum14,
  #          color = spain,
  #          group = country),
  #      size = 2) +
  scale_color_manual(name = '',
                     values = c('black', 'red')) +
  # scale_y_log10() +
  databrew::theme_simple() +
  labs(x = 'Fecha',
        y = 'Casos por 100.000, últimos 14 días',
        title = 'Incidencia Covid-19, Italia y España')
ggsave('~/Desktop/esp_vs_ita.png')


ggplot(data = pd,
       aes(x = date,
           y = deathsp,
           color = country,
           group = country)) +
  geom_line(alpha = 0.7) + 
  # geom_line(data = pd %>% filter(country == 'Spain'),
  #      aes(x = date,
  #          y = cum14,
  #          color = spain,
  #          group = country),
  #      size = 2) +
  scale_color_manual(name = '',
                     values = c('black', 'red')) +
  # scale_y_log10() +
  databrew::theme_simple() +
  labs(x = 'Fecha',
        y = 'Muertos por 100.000, últimos 14 días',
        title = 'Muertes Covid-19, Italia y España')
ggsave('~/Desktop/deaths_esp_vs_ita.png')

ggplot(data = pd,
       aes(x = date,
           y = deaths,
           color = country,
           group = country)) +
  geom_line(alpha = 0.7) + 
  # geom_line(data = pd %>% filter(country == 'Spain'),
  #      aes(x = date,
  #          y = cum14,
  #          color = spain,
  #          group = country),
  #      size = 2) +
  scale_color_manual(name = '',
                     values = c('black', 'red')) +
  # scale_y_log10() +
  databrew::theme_simple() +
  labs(x = 'Fecha',
        y = 'Muertes diarias',
        title = 'Muertes Covid-19, Italia y España')
```


```{r}
pd <- ecdc %>%
  mutate(date = as.Date(dateRep, format = '%d/%m/%Y')) %>%
  mutate(cum14 = Cumulative_number_for_14_days_of_COVID.19_cases_per_100000) %>%
  mutate(country = countriesAndTerritories) %>%
  arrange(date) %>%
  filter(popData2019 >= 10000000) %>%
  group_by(country) %>%
  mutate(deaths14 = zoo::rollsumr(deaths, 14, fill = NA)) %>%
  ungroup %>%
  mutate(deathsp = deaths14 / popData2019 * 100000) %>%
  arrange(desc(cum14)) 


ggplot(data = pd,
       aes(x = date,
           y = cum14,
           group = country,
           color = continentExp)) +
  geom_line(alpha = 0.6) +
  geom_line(data = pd %>% filter(country == 'Spain'),
       aes(x = date,
           y = cum14,
           color = 'red',
           group = country),
       size = 2)
```

```{r}
# Leaflet for municipi
pd <- abs %>%
  # filter(RegioSanitariaDescripcio %in% c('Barcelona Ciutat')) %>%
    filter(RegioSanitariaDescripcio %in% c('Barcelona Ciutat',
                                         'Metropolità Nord',
                                         'Metropolità Sud')) %>%
  arrange(date) %>%
  group_by(ABSCodi) %>%
  mutate(cases_roll = zoo::rollsumr(confirmed_cases_non_cum, k = 7, fill = NA)) %>%
  mutate(incidence_week_100k = cases_roll / pop * 100000)

map <- abs_shp
right <- pd %>% filter(date == (Sys.Date()-3))
map@data$ABSCodi <- map@data$CODIABS

map@data <- left_join(map@data, right)
# map <- map[!map@data$CODIABS %in% pd$ABSCodi,]
map <- map[map@data$CODIABS %in% pd$ABSCodi |
             map@data$NOMABS %in% c('Montcada i Reixac - 2',
                                      'Montcada i Reixac - 1',
                                    'Castellbisbal'),]

map <- spTransform(map, "+init=epsg:4326")

bins <- c(-1, 10, 20, 50, 100, 200, 500,Inf)
library(RColorBrewer)
library(leaflet)
pal <- colorBin("YlOrRd", domain = map@data$incidence_week_100k, bins = bins)

leaflet() %>%
  addProviderTiles(providers$Stamen.TonerBackground) %>%
  addPolygons(data = map,
              
  fillColor = ~pal(incidence_week_100k),
  weight = 0.5,#2,
  opacity = 0.2,
  color = 'black', #"white",
  # dashArray = "3",
  popup = paste0(map@data$NOMABS, ' ', map@data$CODIABS),
  # popup = paste0(map@data$ABSDescripcio, ': ', round(map@data$incidence_week_100k)),
  fillOpacity = 0.9
  )
  
```

```{r}

```

US deaths and cases

```{r}
pd <- owid %>% filter(location == 'United States') %>%
   mutate(deaths_roll = zoo::rollmeanr(new_deaths, k = 7, fill = NA)) %>%
    mutate(cases_roll = zoo::rollmeanr(new_cases, k = 7, fill = NA)) %>%
  dplyr::select(date,new_deaths, new_cases, deaths_roll, cases_roll) %>%
  filter(date >= '2020-03-10')
  # tidyr::gather(key, value, new_deaths:cases_roll)

ggplot(data = pd,
       aes(x = date)) +
  # geom_bar(stat = 'identity', aes(y = new_cases, color = 'Cases', fill = 'Cases')) +
  geom_line(aes(y = cases_roll, color = 'Cases')) +
  geom_line(aes(y = deaths_roll * 15, color = 'Deaths')) +
  geom_point(aes(y = new_cases, color = 'Cases'), alpha = 0.3) +
  geom_point(aes(y = new_deaths * 15, color = 'Deaths'), alpha = 0.3) +
  scale_y_continuous(sec.axis = sec_axis(~./15, name = "Deaths")) +
  databrew::theme_simple() +
  scale_color_manual(name = '',
                     values = c('black', 'red')) +
  # scale_fill_manual(name = '',
  #                    values = c('black', 'red')) +
  labs(y = 'Cases',
       x = 'Date',
       title = 'US Covid-19 Cases and deaths',
       subtitle = 'Dot = daily observation; Line = 7 day rolling average') +
  theme(legend.position = 'bottom')
ggsave('~/Desktop/us_cases_and_deaths.png')
```

```{r}
pd <- covid19::muni
pd$ComarcaDescripcio <- trimws(as.character(pd$ComarcaDescripcio))
keep5 <- function(x){substr(x, 1, 5)}

# muni$ComarcaDescripcio <- gsub('à', 'a', muni$ComarcaDescripcio)
lleida_comarques <- c('Aran', 'Alta Ribagorça', 'Pallars Sobirà',
                      'Pallars Jussà', 'Alt Urgell', 'Cerdanya',
                      'Solsonès', 'Noguera', "Pla d'Urgell",
                      'Segrià', 'Garrigues', 'Urgell', 'Segarra')
pd$ComarcaDescripcio <- keep5(pd$ComarcaDescripcio)
lleida_comarques <- keep5(lleida_comarques)

agg <- pd %>%
  group_by(ComarcaDescripcio, date) %>%
  summarise(cases = sum(confirmed_cases_non_cum, na.rm = TRUE),
            pop = sum(pop)) %>%
  ungroup %>%
  arrange(date) %>% 
  group_by(ComarcaDescripcio) %>%
  mutate(cases_roll = zoo::rollmeanr(cases, k = 7, fill = NA)) %>%
  ungroup %>%
  mutate(cases100k = cases / pop * 100000,
         cases_roll100k = cases_roll / pop * 100000)
ggplot(data = agg %>% filter(ComarcaDescripcio %in% lleida_comarques),
       aes(x = date,
           y = cases100k)) +
  geom_point(alpha = 0.3) +
  geom_line(alpha =0.4) +
  geom_line(aes(y = cases_roll100k), color = 'red') +
  facet_wrap(~ComarcaDescripcio) +
  cowplot::theme_cowplot()  +
  xlim(as.Date('2020-05-01'),
       (Sys.Date()-1))
```

```{r}
# Santa Coloma
scq <- muni %>%
  filter(MunicipiDescripcio == 'Santa Coloma de Queralt') %>%
  dplyr::select(date, confirmed_cases_non_cum)

ggplot(data = scq,
       aes(x = date,y = confirmed_cases_non_cum )) +
  geom_bar(stat = 'identity') +
  databrew::theme_simple() +
  xlim(Sys.Date() - 40, Sys.Date() - 1) +
  labs(x = 'Data',
       y = 'Casos confirmats')

conca <- muni %>%
  filter(grepl('Conca de Barbe', ComarcaDescripcio)) 

ggplot(data = conca,
       aes(x = date,
           y = confirmed_cases_non_cum)) +
  # geom_point() +
  facet_wrap(~MunicipiDescripcio) +
  geom_line()

# All comarques
pd <- muni %>% ungroup %>%
  group_by(date, ComarcaDescripcio) %>%
  summarise(new_cases = sum(confirmed_cases_non_cum),
            pop = sum(pop)) %>%
  ungroup %>%
  arrange(date) %>%
  group_by(ComarcaDescripcio) %>%
  mutate(cases_roll = zoo::rollmeanr(new_cases, k = 7, fill = NA)) %>%
  mutate(yp = new_cases / pop * 100000,
         ypr = cases_roll / pop * 100000) %>%
  mutate(ypl = sqrt(yp),
         yprl = sqrt(ypl))

# ggplot(data = pd,
#        aes(x = date,
#            y = cases_roll)) +
#   geom_point(aes(y = new_cases), alpha = 0.3,
#              size = 0.3) +
#   geom_line(color = 'red', size = 0.3) +
#   facet_wrap(~ComarcaDescripcio, scales = 'free_y')


labeler <- sqrt(c(0, 10, 50, 100, 200))
# pd <- pd %>% filter(date <= '2020-07-16')
levs <- pd %>% filter(date == max(date)) %>% arrange(desc(ypr)) %>% .$ComarcaDescripcio
pd$ComarcaDescripcio <- factor(pd$ComarcaDescripcio, levels = rev(levs))

# Create weekly (instead of pd)
weekly <- pd %>%
   mutate(week = lubridate::week(date)) %>%
  group_by(week, ComarcaDescripcio) %>%
  mutate(week_start = min(date)) %>%
  mutate(n_days = length(unique(date))) %>%

ggplot(data = pd %>% filter(date >= '2020-03-11',
                            date <= '2020-08-18'),
       aes(x = date,
           y = ComarcaDescripcio,
           fill = ypl)) +
  geom_tile(
    # color = 'black',size = 0.2
    ) +
  databrew::theme_simple() +
  scale_fill_gradientn(name = '', colors = colorRampPalette(c('beige',
                                                              # 'grey',
                                                              # 'yellow',
                                                              # 'white',
                                                              # 'orange',
                                                              'darkorange', 
                                                              'red',
                                                              'darkred',
                                                              'black'))(20),
                       breaks = labeler, labels = round(labeler^2)) +
  labs(x = '',
       y = '',
       title = 'Casos confirmats nous (per 100.000 habitants)',
       caption = 'Aix-x: dia; Aix-y: comarca; Color: casos nous.\nDades de https://analisi.transparenciacatalunya.cat/. Gràfic: @joethebrew') +
  theme(axis.text.y = element_text(size = 10))
ggsave('~/Desktop/daily.png', width = 10, height = 6)

```

```{r}
# Weekly version
pd <- muni %>% ungroup %>%
  group_by(date, ComarcaDescripcio) %>%
  summarise(new_cases = sum(confirmed_cases_non_cum),
            pop = sum(pop)) %>%
  ungroup %>%
  arrange(date) %>%
  group_by(ComarcaDescripcio) %>%
  mutate(cases_roll = zoo::rollmeanr(new_cases, k = 7, fill = NA)) %>%
  mutate(yp = new_cases / pop * 100000,
         ypr = cases_roll / pop * 100000) %>%
  mutate(ypl = sqrt(yp),
         yprl = sqrt(ypl))


labeler <- sqrt(c(0, 10, 50, 100, 200))
# pd <- pd %>% filter(date <= '2020-07-16')
levs <- pd %>% filter(date == max(date)) %>% arrange(desc(ypr)) %>% .$ComarcaDescripcio
pd$ComarcaDescripcio <- factor(pd$ComarcaDescripcio, levels = rev(levs))


ggplot(data = pd %>% filter(date >= '2020-03-11',
                            date <= '2020-08-18'),
       aes(x = date,
           y = ComarcaDescripcio,
           fill = ypl)) +
  geom_tile(
    # color = 'black',size = 0.2
    ) +
  databrew::theme_simple() +
  scale_fill_gradientn(name = '', colors = colorRampPalette(c('beige',
                                                              # 'grey',
                                                              # 'yellow',
                                                              # 'white',
                                                              # 'orange',
                                                              'darkorange', 
                                                              'red',
                                                              'darkred',
                                                              'black'))(20),
                       breaks = labeler, labels = round(labeler^2)) +
  labs(x = '',
       y = '',
       title = 'Casos confirmats nous (per 100.000 habitants)',
       caption = 'Aix-x: dia; Aix-y: comarca; Color: casos nous.\nDades de https://analisi.transparenciacatalunya.cat/. Gràfic: @joethebrew') +
  theme(axis.text.y = element_text(size = 10))
ggsave('~/Desktop/weekly.png', width = 10, height = 6)

```

```{r}
# 50 as the cut-off
pd$fifty <- ifelse(pd$yp >= 50/7, '>50',
                   ifelse(pd$yp >= 25/7, '25-50',
                   '<25'))
pd$fifty <- factor(pd$fifty,
                   levels = c('<25', '25-50', '>50'))
ggplot(data = pd %>% filter(date >= '2020-03-11'),
       aes(x = date,
           y = ComarcaDescripcio,
           fill = fifty)) +
  geom_tile(
    # color = 'black',size = 0.2
    ) +
  databrew::theme_simple() +
  scale_fill_manual(name = 'Incidència\nper 100.000',
                    values = c('beige', 'darkorange', 'darkred')) +
  labs(x = '',
       y = '',
       title = 'Casos nous\n(convertit en incidència setmanal per 100.000)',
       caption = 'Eix-x: dia; Eix-y: comarca; Color: casos nous.\nDades de https://analisi.transparenciacatalunya.cat/. Gràfic: @joethebrew') +
  theme(axis.text.y = element_text(size = 10))
ggsave('~/Desktop/daily_binary.png', width = 10, height = 6)

```

```{r}
# Not population-adjusted
ggplot(data = pd,
       aes(x = date,
           y = ComarcaDescripcio,
           fill = new_cases)) +
  geom_tile(color = 'white', size = 0.2) +
  databrew::theme_simple() +
  scale_fill_gradientn(name = '', colors = colorRampPalette(c('white', 
                                                              'grey',
                                                              # 'white',
                                                              'darkorange', 
                                                              'red',
                                                              'black'))(20)) +#,
                       # breaks = labeler, labels = round(labeler^2)) +
  labs(x = '',
       y = '',
       title = 'Casos confirmats nous') +
  theme(axis.text.y = element_text(size = 10))
```

```{r}
# Anoia
pd <- muni %>%
  filter(ComarcaDescripcio == 'Anoia') %>%
  group_by(date_month = lubridate::month(date),
           date_day = lubridate::day(date)) %>%
  summarise(cases = sum(confirmed_cases_non_cum)) %>%
  filter(date_month %in% c(3, 7)) %>%
  mutate(date_month = factor(date_month,
                             levels = c('3', '7'),
                             labels = c('Març', 'Juliol')))

ggplot(data = pd,
       aes(x = date_day,
           y = cases,
           group = date_month,
           color = date_month)) +
  geom_line()

```

```{r}
# Top few
# All comarques
pd <- muni %>% ungroup %>%
  group_by(date, ComarcaDescripcio) %>%
  summarise(new_cases = sum(confirmed_cases_non_cum),
            pop = sum(pop)) %>%
  ungroup %>%
  arrange(date) %>%
  group_by(ComarcaDescripcio) %>%
  mutate(cases_roll = zoo::rollmeanr(new_cases, k = 7, fill = NA)) %>%
  mutate(yp = new_cases / pop * 100000,
         ypr = cases_roll / pop * 100000) %>%
  mutate(ypl = sqrt(yp),
         yprl = sqrt(ypl))

# ggplot(data = pd,
#        aes(x = date,
#            y = cases_roll)) +
#   geom_point(aes(y = new_cases), alpha = 0.3,
#              size = 0.3) +
#   geom_line(color = 'red', size = 0.3) +
#   facet_wrap(~ComarcaDescripcio, scales = 'free_y')


labeler <- sqrt(c(0, 10, 50, 100, 200))
# pd <- pd %>% filter(date <= '2020-07-16')
levs <- pd %>% filter(date == max(date)) %>% arrange(desc(ypr)) %>% .$ComarcaDescripcio
pd$ComarcaDescripcio <- factor(pd$ComarcaDescripcio, levels = (levs))
# Keep only top few
pd <- pd %>%
  filter(ComarcaDescripcio %in% levs[1:10]) %>%
  filter(date >= '2020-07-01')

ggplot(data = pd,
       aes(x = date,
           y = yp,
           color = ComarcaDescripcio)) +
  geom_line()
```

# Small municipalities with cases in the last week

```{r}
# Above the 50 per 100.000 week
pd <- muni %>%
  filter(date >= (Sys.Date() - 7),
         date <= (Sys.Date()-1)) %>%
  # filter(pop < 4000) %>%
  group_by(MunicipiDescripcio) %>%
  summarise(cases = sum(confirmed_cases_non_cum),
            pop = mean(pop),
            id = dplyr::first(MunicipiCodi)) %>%
  ungroup %>%
  mutate(per100000week = (cases/pop) * 100000) %>%
  arrange(desc(per100000week), pop)
pd

```

```{r}
library(maptools)
catalan_codes <- c('43', '08', '25', '17')

map <- municipios
map <- map[substr(map@data$id, 1, 2) %in% catalan_codes, ]
map_fort <- fortify(map, id = 'id' )# %>% mutate(id = as.numeric(as.character(id)))
# Dot plot animation
dot_plot <- function(the_date = Sys.Date(),
                     out_dir = '~/Desktop/animation'){
  if(!dir.exists(out_dir)){
    dir.create(out_dir)
  }
  
  map <- municipios
  map <- map[substr(map@data$id, 1, 2) %in% catalan_codes, ]
  
  right <- muni %>%
    mutate(id = MunicipiCodi) %>%
    # filter(date == the_date) %>%
    # filter(date %in% seq( (the_date - 6), the_date, 1)) %>%
    group_by(id) %>%
    summarise(cases = sum(confirmed_cases_non_cum[date == the_date], na.rm = TRUE),
              cases1 = sum(confirmed_cases_non_cum[date == (the_date -1)], na.rm = TRUE),
              cases2 = sum(confirmed_cases_non_cum[date == (the_date -2)], na.rm = TRUE))
  
  map@data <- map@data %>%
    left_join(right) %>%
    mutate(cases = ifelse(is.na(cases), 0, cases),
           cases1 = ifelse(is.na(cases1), 0, cases1),
           cases2 = ifelse(is.na(cases2), 0, cases2))
  final <- map_fort %>%
    left_join(right) %>%
    mutate(cases = ifelse(is.na(cases), 0, cases),
           cases1 = ifelse(is.na(cases1), 0, cases1),
           cases2 = ifelse(is.na(cases2), 0, cases2))
  
  dots <- dotsInPolys(map, as.integer(map@data$cases))
  dots1 <- dotsInPolys(map, as.integer(map@data$cases1))
  dots2 <- dotsInPolys(map, as.integer(map@data$cases2))
  
  dots_df <- tibble(long = coordinates(dots)[,1],
                    lat = coordinates(dots)[,2],
                    ID  = dots$ID) 
  dots_df1 <- tibble(long = coordinates(dots1)[,1],
                    lat = coordinates(dots1)[,2],
                    ID  = dots1$ID) 
  dots_df2 <- tibble(long = coordinates(dots2)[,1],
                    lat = coordinates(dots2)[,2],
                    ID  = dots2$ID) 
  g <- ggplot() +
    geom_polygon(data = map_fort,
                 aes(x = long,
                     y = lat,
                     group = group),
                 fill = 'black',
                 color = 'darkgrey',
                 size = 0.1) +
    geom_point(data = dots_df,
               aes(x = long,
                   y = lat),
               color = 'red',
               alpha = 0.9,
               size = 0.15) +
    geom_point(data = dots_df1,
               aes(x = long,
                   y = lat),
               color = 'red',
               alpha = 0.5,
               size = 0.15) +
        geom_point(data = dots_df2,
               aes(x = long,
                   y = lat),
               color = 'red',
               alpha = 0.2,
               size = 0.15) +
    labs(title = the_date) +
    databrew::theme_simple() +
    theme_map() +
    coord_map() +
    theme(plot.title = element_text(size = 22))
  ggsave(paste0(out_dir, '/', the_date, '.png'),
                height = 8,
                width = 10)
  
}

# dot_plot(the_date = Sys.Date() -5 )

dates <- seq(as.Date('2020-03-10'), Sys.Date()-1, 1)
# dates <- dates[weekdays(dates) == 'Tuesday']
for(i in 1:length(dates)){
  message(dates[i])
  dot_plot(dates[i])
}
```

```{r}
# choropleth
map <- municipios
map <- map[substr(map@data$id, 1, 2) %in% catalan_codes, ]
map_fort <- fortify(map, region  = 'id' )# %>% mutate(id = as.numeric(as.character(id)))
# Dot plot animation
choro_plot <- function(the_date = Sys.Date(),
                     out_dir = '~/Desktop/animation_choro'){
  if(!dir.exists(out_dir)){
    dir.create(out_dir)
  }

  right <- muni %>%
    mutate(id = MunicipiCodi) %>%
    # filter(date == the_date) %>%
    # filter(date %in% seq( (the_date - 6), the_date, 1)) %>%
    group_by(id) %>%
    summarise(cases = sum(confirmed_cases_non_cum[date == the_date], na.rm = TRUE),
              cases1 = sum(confirmed_cases_non_cum[date == (the_date -1)], na.rm = TRUE),
              cases2 = sum(confirmed_cases_non_cum[date == (the_date -2)], na.rm = TRUE),
              cases_lag = sum(confirmed_cases_non_cum[date <= the_date & date >= (the_date-6)]),
              pop = dplyr::first(pop)) %>%
    ungroup %>%
    mutate(cases_k = cases_lag / pop * 100000) %>%
    mutate(cases = ifelse(is.na(cases), 0, cases),
           cases1 = ifelse(is.na(cases1), 0, cases1),
           cases2 = ifelse(is.na(cases2), 0, cases2),
           cases_lag = ifelse(is.na(cases_lag), 0, cases_lag),
           cases_k = ifelse(is.na(cases_k), 0, cases_k))

  final <- map_fort %>%
    left_join(right)
  
  final$category <- ifelse(final$cases_k == 0, 
                                  '0',
                                  ifelse(final$cases_k <= 20,
                           '0-20',
                           ifelse(final$cases_k <= 50,
                                  '20-50',
                                  ifelse(final$cases_k <= 200,
                                         '50-200',
                                         '200+'))))
  final$category <- ifelse(is.na(final$category), 0, final$category)
  final$category <- factor(final$category, 
                           levels = c('0', '0-20','20-50', '50-200', '200+'))
  g <- ggplot() +
    geom_polygon(data = final,
                 aes(x = long,
                     y = lat,
                     group = group,
                     fill = category),
                 color = 'black',
                 size = 0.1) +
    labs(title = the_date,
         subtitle = 'Casos confirmats (darrers 7 dies) per 100.000 habitants' ) +
    databrew::theme_simple() +
    theme_map() +
    coord_map() +
    theme(plot.title = element_text(size = 24),
          plot.subtitle = element_text(size = 20),
          legend.text = element_text(size = 20)) +
    scale_fill_manual(name = '',
                         values = c( 'beige', 'yellow', 'darkorange', 'red', 'darkred')) +
    # scale_fill_gradientn(name = 'Incidència setmanal\n(per 100.000)',
    #                      colors = c('white', 'beige', 'darkorange', 'red', 'darkred'),
    #                      breaks = c(0, 50, 200, 500),
    #                      limits = c(0, (round(max(muni$confirmed_cases_non_cum / muni$pop * 100000)))),
  # na.value = 'white')
    theme(legend.position = 'bottom')
  ggsave(paste0(out_dir, '/', the_date, '.png'),
                height = 8,
                width = 10)
  
}

# dot_plot(the_date = Sys.Date() -5 )

dates <- seq(as.Date('2020-03-10'), Sys.Date()-3, 1)
dates <- dates[c(1:5, (length(dates) -20):length(dates))]
# dates <- dates[weekdays(dates) == 'Tuesday']
for(i in 1:length(dates)){
  message(dates[i])
  choro_plot(dates[i])
}
```

```
# Command line
cd ~/Desktop/animation
mogrify -resize 50% *.png
convert -delay 20 -loop 0 *.png result.gif

```

# Time chart 

```{r}
pd <- muni %>% ungroup %>%
  group_by(date, ComarcaDescripcio) %>%
  summarise(new_cases = sum(confirmed_cases_non_cum),
            pop = sum(pop)) %>%
  ungroup %>%
  arrange(date) %>%
  group_by(ComarcaDescripcio) %>%
  mutate(cases_roll = zoo::rollsumr(new_cases, k = 7, fill = NA)) %>%
  mutate(yp = new_cases / pop * 100000,
         ypr = cases_roll / pop * 100000) %>%
  mutate(danger = ypr >= 50)

library(ggrepel)
make_plot <- function(the_date,
                      out_dir = 'ts_chart'){
  if(!dir.exists('ts_chart')){
    dir.create('ts_chart')
  }
  sub_data <- pd %>%
    filter(date <= the_date) %>%
    filter(!is.na(danger))
  label_df <- sub_data %>% filter(date == the_date) %>%
    arrange(desc(ypr))
  label_df <- label_df[1:3,]
  label_df <- label_df %>% filter(ypr >= 50)
  top3 <- sub_data %>% filter(ComarcaDescripcio %in% label_df$ComarcaDescripcio)
  sub_data <- sub_data %>%
    filter(!ComarcaDescripcio %in% label_df$ComarcaDescripcio)
  ggplot() +
    geom_line(alpha = 0.5,
              data = sub_data,
         aes(x = date,
             y = ypr,
             group = ComarcaDescripcio)) +
    geom_line(alpha = 0.9,
              data = top3,
         aes(x = date,
             y = ypr,
             group = ComarcaDescripcio),
         color = 'darkred') +
    # scale_y_log10() +
    geom_hline(yintercept = 50, color = 'blue') +
    databrew::theme_simple() +
    geom_point(data = label_df,
              aes(x = date ,
                  y = ypr),
              color = 'darkred') +
    geom_text(data = label_df,
              aes(x = date + 0.5,
                  y = ypr,
                  label = ComarcaDescripcio)) +
    labs(title = the_date,
         x = 'Data',
         y = 'Casos per 100.000 (darrers 7 dies)') +
    theme(legend.position = 'none') +
    xlim(min(sub_data$date),
         max(sub_data$date) + 2) +
    ylim(0, max(pd$ypr, na.rm = T)+1)
  ggsave(paste0(out_dir, '/', the_date, '.png'),
                height = 8,
                width = 10) 
}
# dates <- sort(unique(pd$date))
dates <- seq(as.Date('2020-03-10'),
             Sys.Date()-1,
             by = 1)
for(i in 1:length(dates)){
  make_plot(the_date = dates[i])
}
```

```
# Command line
cd ts_chart
mogrify -resize 50% *.png
convert -delay 15 -loop 0 *.png result.gif

```

```{r}
pd <- muni %>%
  filter(date >= (Sys.Date() - 8)) %>%
  filter(pop < 4000) %>%
  group_by(MunicipiDescripcio) %>%
  summarise(cases = sum(confirmed_cases_non_cum),
            pop = mean(pop)) %>%
  arrange(desc(cases), pop) %>%
  filter(cases > 5)

pd <- muni %>%
  filter(date >= (Sys.Date() - 8)) %>%
  group_by(MunicipiDescripcio) %>%
  summarise(cases = sum(confirmed_cases_non_cum),
            pop = mean(pop)) %>%
  arrange(desc(cases), pop) %>%
  filter(cases > 1)

fit <- lm(cases ~ pop, data = pd)
ggplot(data = pd,
       aes(x = pop,
           y = cases)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(formula= y~x)

```

```{r}
# Now municipalities
pd <- muni %>% ungroup %>%
  group_by(date, MunicipiDescripcio) %>%
  summarise(new_cases = sum(confirmed_cases_non_cum),
            pop = sum(pop)) %>%
  ungroup %>%
  arrange(date) %>%
  group_by(MunicipiDescripcio) %>%
  mutate(cases_roll = zoo::rollmeanr(new_cases, k = 7, fill = NA)) %>%
  mutate(yp = new_cases / pop * 100000,
         ypr = cases_roll / pop * 100000) %>%
  mutate(ypl = sqrt(yp),
         yprl = sqrt(ypl))

# ggplot(data = pd,
#        aes(x = date,
#            y = cases_roll)) +
#   geom_point(aes(y = new_cases), alpha = 0.3,
#              size = 0.3) +
#   geom_line(color = 'red', size = 0.3) +
#   facet_wrap(~ComarcaDescripcio, scales = 'free_y')


labeler <- sqrt(c(0, 10, 50, 100, 200))
pd <- pd %>% filter(date <= (Sys.Date()-1))
levs <- pd %>% filter(date == max(date)) %>% arrange(desc(ypr)) %>% .$MunicipiDescripcio
pd$MunicipiDescripcio <- factor(pd$MunicipiDescripcio, levels = rev(levs))

pd$fifty <- ifelse(pd$yp >= 50, '>50',
                   ifelse(pd$yp >= 40, '40-50',
                          ifelse(pd$yp >= 30, '30-40',
                          ifelse(pd$yp >= 20, '20-30',
                                 ifelse(pd$yp >= 10, '10-20',
                                        ifelse(pd$yp > 0, '<10', '0'))))))
pd$fifty <- factor(pd$fifty,
                   levels = c('0',
                              '<10',
                              '10-20',
                              '20-30',
                              '30-40',
                              '40-50',
                              '>50'))

comarques <- sort(unique(muni$ComarcaDescripcio))
dir.create('~/Desktop/comarques')
for(i in 1:length(comarques)){
  this_comarca <- comarques[i]
  these_municipis <- muni$MunicipiDescripcio[muni$ComarcaDescripcio == this_comarca]
  sub_pd <- pd %>%
    filter(MunicipiDescripcio %in% these_municipis)
  sub_pd$MunicipiDescripcio <- factor(sub_pd$MunicipiDescripcio, levels = rev(unique(sub_pd$MunicipiDescripcio)))
  ggplot(data = sub_pd,
         aes(x = date,
             y = MunicipiDescripcio,
             fill = yp
             # fill = fifty
             )) +
    geom_tile(
      color = 'white', size = 0.02
      ) +
    databrew::theme_simple() +
    scale_fill_gradientn(name = '', colors = colorRampPalette(c('beige',                                                              'darkorange', 'red', 'darkred','black'))(20)) +
    # scale_fill_manual(name = 'Incidència',
    #                   values = colorRampPalette(c('white',
    #                                               'beige',
    #                              'darkorange',
    #                              'darkred'))(length(unique(pd$fifty)))) +
    labs(x = '',
         y = '',
         title = paste0(this_comarca, ': Casos nous confirmats per 100.000'))
  ggsave(paste0('~/Desktop/comarques/', this_comarca, '.png'),
         height = 6, width = 11.5)
}


```

```{r}
# Number of municipios with any cases
pd <- muni %>%
  group_by(date) %>%
  summarise(any_case = length(which(confirmed_cases_non_cum > 0)),
            new_cases = sum(confirmed_cases_non_cum)) %>%
  ungroup 

ggplot(data = pd,
       aes(x = date,
           y = any_case)) +
  geom_point() +
  geom_line()
```






```{r}
# Date of comarcas first case
x <- covid19::muni %>%
  group_by(MunicipiDescripcio) %>%
  summarise(`Casos confirmats` = sum(confirmed_cases_non_cum),
            `Població` = mean(pop),
            date = min(date[confirmed_cases_non_cum > 0])) %>%
  ungroup
y <- x %>%
  filter(`Casos confirmats` == 0) %>%
  arrange(desc(`Població`)) %>% dplyr::select(-date)
```


```{r}
# Catalonia age
edat <- read_csv('https://analisi.transparenciacatalunya.cat/api/views/qwj8-xpvk/rows.csv?accessType=DOWNLOAD&sorting=true')
pd <- edat %>%
  mutate(date = as.Date(TipusCasData, format = '%d/%m/%Y'),
         age_group = EdatRang) %>%
  filter(!TipusCasDescripcio %in% c('Sospitós'),
         !EdatRang %in% c('No classificat')) %>%
  group_by(date, age_group) %>%
  summarise(n = sum(NumCasos, na.rm = TRUE))

left <- expand.grid(date = seq(min(pd$date),
                               max(pd$date),
                               by = 1),
                    age_group = sort(unique(pd$age_group)))
joined <- left_join(left, pd) %>%
  mutate(n = ifelse(is.na(n), 0, n)) %>%
  filter(date >= '2020-03-04')

# Get by week
bw <- joined %>%
  mutate(week = lubridate::week(date)) %>%
  group_by(week) %>%
  mutate(week_start = min(date)) %>%
  mutate(n_days = length(unique(date))) %>%
  ungroup %>%
  dplyr::select(-date) %>%
  group_by(date = week_start, age_group) %>%
  summarise(n = sum(n, na.rm = TRUE),
            n_days = dplyr::first(n_days)) %>%
  ungroup %>%
  mutate(per_day = n / n_days)

ggplot(data = bw %>% filter(date != max(date)),
       aes(x = date,
           y = age_group,
           fill = per_day)) +
    geom_tile() +
    databrew::theme_simple() +
    scale_fill_gradientn(name = '', colors =
                           colorRampPalette(c('lightyellow', 
                                                                # 'yellow',
                                                                # 'lightblue', 
                                                                # 'white',
                                                                'darkorange', 
                                                                'darkred'))(20)) +
    labs(x = '',
         y = '',
         title = 'Catalunya: casos confirmats per edat',
         subtitle = 'Catalonia: confirmed cases by age',
         caption = 'Eix-x: setmana; Eix-y: edat. Color: casos nous.\nDades de https://analisi.transparenciacatalunya.cat/. Gràfic: @joethebrew')
ggsave('~/Desktop/by_age.png', height = 4, width = 7)  

```

```{r}
# Age by month
pd <- edat %>%
  mutate(date = as.Date(TipusCasData, format = '%d/%m/%Y'),
         age_group = EdatRang) %>%
  filter(!TipusCasDescripcio %in% c('Sospitós'),
         !EdatRang %in% c('No classificat')) %>%
  mutate(date_month = paste0(format(date, '%Y-%m'), '-01')) %>%
  mutate(age = ifelse(age_group %in% c('60-69', '70-79', '80-89', '90+'), '>60', '<60')) %>%
  group_by(date_month, age) %>%
  summarise(n = sum(NumCasos, na.rm = TRUE),
            days = length(unique(date))) %>%
  group_by(date_month) %>% 
  mutate(p = n / sum(n) * 100) %>%
  filter(date_month >= '2020-03-01') %>%
  mutate(date_label = format(as.Date(date_month), '%b')) %>%
  mutate(date_label = factor(date_label, 
                             levels = c('Mar','Apr','May','Jun','Jul', 'Aug'),
                             labels = c('Mar','Abr','Mai','Jun','Jul', 'Ago'))) %>%
  mutate(n = n / days) %>%
  dplyr::select(-days) %>%
  gather(key, value, n:p) %>%
  mutate(key = ifelse(key == 'n',
                      'Casos per dia',
                      'Percentatge de tots els casos'))
# pd <- pd %>% filter(date_label != 'Ago')

ggplot(data = pd,
       aes(x = date_label,
           y = value,
           color = age,
           group = age)) +
  geom_line(size = 2, alpha = 0.6) +
  # geom_point(size = 2) +
  databrew::theme_simple() +
  facet_wrap(~key, scales = 'free_y') +
  labs(y = '',
       x = 'Mes',
       title = 'Casos confirmats i edat: Covid-19 a Catalunya') +
  scale_color_manual(name = 'Edat',
                     values = c('darkorange',
                                'purple')) +
  theme(legend.position = 'top',
        strip.text = element_text(size = 18))
ggsave('~/Desktop/age2.png', width = 8.5, height = 5)
```

```{r}
# Age by percentage
bwp <- bw %>%
  group_by(date) %>%
  mutate(p = per_day / sum(per_day) * 100) %>%
  ungroup
ggplot(data = bwp,
       aes(x = date,
           y = age_group,
           fill = p)) +
    geom_tile() +
    databrew::theme_simple() +
    scale_fill_gradientn(name = '', colors =
                           colorRampPalette(c('lightyellow', 
                                                                # 'yellow',
                                                                # 'lightblue', 
                                                                # 'white',
                                                                'darkorange', 
                                                                'darkred'))(20)) +
    labs(x = '',
         y = '',
         title = 'Catalunya: casos confirmats per edat',
         subtitle = 'Catalonia: confirmed cases by age',
         caption = 'Aix-x: setmana; Aix-y: edat. Color: casos nous.\nDades de https://analisi.transparenciacatalunya.cat/. Gràfic: @joethebrew')
ggsave('~/Desktop/by_age.png', height = 4, width = 7) 
```

```{r}
fl <- df %>%
  filter(district == 'Florida')

ggplot(data = df,
       aes(x = date,
           y = cases)) +
  geom_line()
```

Municipalities

```{r}
pd <- muni %>%
  filter(date == max(date)) %>%
    # Get confirmed cases by population
    mutate(p = confirmed_cases / pop * 100000) %>%
  arrange(desc(p))
```

```{r}

c25 <- c(
  "dodgerblue2", "#E31A1C", # red
  "green4",
  "#6A3D9A", # purple
  "#FF7F00", # orange
  "black", "gold1",
  "skyblue2", "#FB9A99", # lt pink
  "palegreen2",
  "#CAB2D6", # lt purple
  "#FDBF6F", # lt orange
  "gray70", "khaki2",
  "maroon", "orchid1", "deeppink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3",
  "darkorange4", "brown"
)

comarcas <- sort(unique(muni$ComarcaDescripcio))
for(i in 1:length(comarcas)){
  this_comarca <- comarcas[i]
  pd <- muni %>%
    # Get confirmed cases by population
    mutate(p = confirmed_cases / pop * 100000) %>%
    dplyr::filter(ComarcaDescripcio %in% this_comarca)
  
  # pie(rep(1, 25), col = c25)
  cols <- colorRampPalette(c25)(length(unique(pd$MunicipiDescripcio)))
  
  g <- ggplot(data = pd,
         aes(x = date,
             y = p)) +
    geom_line(aes(group = MunicipiDescripcio,
                  color = MunicipiDescripcio,
                  size = pop),
              alpha = 0.8) +
    # scale_y_log10() +
    # facet_wrap(~MunicipiDescripcio, scales = 'free_y') +
      scale_color_manual(name = '', values = cols) +
    databrew::theme_simple() +
    labs(x = 'Data',
         y = 'Casos per 100.000',
         title = paste0(this_comarca),
         subtitle = 'Incidència acumulada per 100.000 habitants') +
    scale_size(name = 'Població') +
    guides(color = guide_legend(override.aes = list(size = 2)))
  print(g)
  # Sys.sleep(10)
}

```

Cumulative incidence map, Catalonia, by municipality

```{r}
# map <- municipios[municipios@data$id %in% muni$MunicipiCodi,]
catalan_codes <- c('43', '08', '25', '17')
map <- municipios[substr(municipios@data$id, 1, 2) %in% catalan_codes, ]

map_fortified <- fortify(map, region = 'id')
# Define function for getting each day
get_day <- function(the_date = Sys.Date()-3){
  pd <- muni %>% filter(date == the_date)
  joined <- left_join(map_fortified, pd %>% dplyr::rename(id = MunicipiCodi)) %>%
    mutate(p = confirmed_cases / pop * 100000)
  
  ggplot(data = joined,
         aes(x = long,
             y = lat,
             fill = p)) +
    geom_polygon(aes(group = id)) +
    scale_fill_gradientn(name = '',
                         colors = c('white', 'yellow', 'darkorange', 'red', 'darkred', 'black'),
                       # colors = rev(RColorBrewer::brewer.pal(n = 9, name = 'Spectral')),
                       limits = c(0, 10000)) +
    xlim(1,2.5) +
    ylim(41, 42.1)
}
dates <- seq(as.Date('2020-03-08'),
             max(muni$date),
             by = 1)

for(i in 1:length(dates)){
  this_date <- dates[i]
  get_day(the_date = this_date)
}
```




```{r}
# Excess mortality spain

pd <- excess %>% dplyr::rename(ccaa = region) %>% 
  filter(ccaa != 'Spain') %>%
  left_join(esp_pop %>%
              mutate(ccaa = ifelse(ccaa == 'Baleares', 'Balears, Illes',
                                   ifelse(ccaa == 'CLM', 'Castile-La Mancha',
                                          ifelse(ccaa == 'CyL', 'Castile & León',
                                                 ifelse(ccaa == 'C. Valenciana', 'Comunitat Valenciana', ccaa)))))) %>%
  mutate(ccaa = ifelse(ccaa == 'Cataluña', 'Catalonia', ccaa)) %>%
  mutate(ccaa = ifelse(ccaa == 'Comunitat Valenciana', 'Valencia', ccaa)) %>%
  mutate(ccaa = ifelse(ccaa == 'País Vasco', 'Basque country', ccaa)) %>%
  mutate(date = end_date)

# Calculate percentage of normal
pd <- pd %>%
  mutate(p = total_deaths / expected_deaths * 100) %>% filter(!ccaa %in% c('Ceuta', 'Melilla', 'Canarias')) %>%
  mutate(type = ifelse(p > 100, 'Above\nexpected\n', 'Below\nexpected\n'))

# pd <- pd %>% filter(ccaa %in% c('Castile & León',
#                                 'Castile-La Mancha',
#                                 'Catalonia',
#                                 'Madrid'))
pd <- pd %>% filter(date > '2020-01-01')

ggplot(data = pd,
       aes(x = date,
           y = p,
           group = ccaa)) +
  geom_hline(yintercept = 100, color = 'darkgrey') +
  # geom_line() +
  # geom_area(aes(ymin = 100)) +
  geom_ribbon(aes(ymin=100, ymax=p), fill = 'darkgrey', alpha = 0.5) +
  geom_line() +
  geom_point(aes(color = type), alpha = 0.7) +
  facet_wrap(~ccaa) +
  scale_color_manual(name = '',
                     values = c('darkred', 'blue')) +
  # scale_fill_manual(name = '',
  #                    values = c('darkred', 'blue')) +
  databrew::theme_simple() +
  theme(legend.position = 'right',
        strip.text = element_text(size = 12, hjust = 0.5)) +
  labs(x = '',
       y = 'Percent',
       title = 'Excess mortality by region',
       caption = 'Raw from the Economist: https://github.com/TheEconomist/covid-19-excess-deaths-tracker/blob/master/output-data/historical-deaths/spain_weekly_deaths.csv\nChart by Joe Brew. Ceuta, Melilla, and Canarias removed due to low numbers.') 
  # geom_hline(yintercept = 100)
ggsave('~/Desktop/plot.png')
```

```{r}
pd <- excess %>% dplyr::rename(ccaa = region) %>% 
  filter(ccaa != 'Spain') %>%
  left_join(esp_pop %>%
              mutate(ccaa = ifelse(ccaa == 'Baleares', 'Balears, Illes',
                                   ifelse(ccaa == 'CLM', 'Castile-La Mancha',
                                          ifelse(ccaa == 'CyL', 'Castile & León',
                                                 ifelse(ccaa == 'C. Valenciana', 'Comunitat Valenciana', ccaa)))))) %>%
  mutate(ccaa = ifelse(ccaa == 'Cataluña', 'Catalonia', ccaa)) %>%
  mutate(ccaa = ifelse(ccaa == 'Comunitat Valenciana', 'Valencia', ccaa)) %>%
  mutate(ccaa = ifelse(ccaa == 'País Vasco', 'Basque country', ccaa)) %>%
  mutate(date = end_date)

# Calculate percentage of normal
pd <- pd %>%
  mutate(p = total_deaths / expected_deaths * 100) %>% filter(!ccaa %in% c('Ceuta', 'Melilla', 'Canarias')) %>%
  mutate(type = ifelse(p > 100, 'Above\nexpected\n', 'Below\nexpected\n'))

# pd <- pd %>% filter(ccaa %in% c('Castile & León',
#                                 'Castile-La Mancha',
#                                 'Catalonia',
#                                 'Madrid'))
pd <- pd %>% filter(date > '2020-01-01')
pd$grp <- ifelse(grepl('Madrid|Castil|Rioj|Catal|Navarr', pd$ccaa), pd$ccaa, 'Others')
cols <- c(RColorBrewer::brewer.pal(n = length(unique(pd$grp))-1, 'Set1'), 'black')
cols[length(cols)-1] <- 'pink'
pd <- pd %>% arrange(desc(p))
pd$ccaa <- factor(pd$ccaa, levels = unique(pd$ccaa))
ggplot(data = pd,
       aes(x = date,
           y = p,
           group = ccaa)) +
  geom_hline(yintercept = 100, color = 'darkgrey') +
  geom_line(data = pd %>% filter(grp == 'Others'),
            aes(color = grp), alpha = 0.7) +
    geom_line(data = pd %>% filter(grp != 'Others'),
            aes(color = grp),
            size = 2, alpha = 0.8) +
  databrew::theme_simple() +
  theme(legend.position = 'right',
        strip.text = element_text(size = 12, hjust = 0.5)) +
  scale_color_manual(name = '',
                     values = cols) +
  labs(x = '',
       y = 'Percent',
       title = 'Excess mortality by region',
       caption = 'Raw from the Economist: https://github.com/TheEconomist/covid-19-excess-deaths-tracker/blob/master/output-data/historical-deaths/spain_weekly_deaths.csv\nChart by Joe Brew. Ceuta, Melilla, and Canarias removed due to low numbers.') 
ggsave('~/Desktop/plot2.png')
```

```{r}
pd <- df_country %>%
  filter(country == 'India')

ggplot(data = pd,
       aes(x = date,
           y = cases_non_cum)) +
  geom_point() +
  geom_segment(aes(xend = date, yend = 0)) +
  theme_simple() +
  labs(x = 'Date',
       y = 'Incident cases',
       title = 'INDIA: Confirmed COVID-19 cases')
```

```{r}
pd <- esp_df %>%
  filter(ccaa %in% c('Cataluña', 'Madrid'))
ggplot(data = pd,
       aes(x = date,
           y = cases_non_cum)) +
  geom_point() +
  geom_segment(aes(xend = date, yend = 0)) +
  theme_simple() +
  labs(x = 'Date',
       y = 'Incident cases',
       title = 'Confirmed COVID-19 cases') +
  facet_wrap(~ccaa)
```

```{r}
# Africa cases
# Number of Africa countries
pd <- df_country
pd %>% left_join(world_pop) %>%
  # filter(sub_region %in% c('Sub-Saharan Africa')) %>%
  filter(region %in% 'Africa') %>%
  filter(cases > 0) %>%
  group_by(date) %>%
  summarise(cases = sum(cases))
```

```{r}
# Number of Africa countries
pd <- df_country
pd <- pd %>% left_join(world_pop) %>%
  filter(sub_region %in% c('Sub-Saharan Africa')) %>%
  filter(cases > 0)
pd <- pd %>% group_by(date) %>% tally

ggplot(data = pd,
       aes(x = date,
           y = n)) +
  # geom_line() +
  geom_area(fill = 'darkorange',
            color = 'black',
            alpha = 0.6) +
  theme_simple() +
  labs(x = 'Date',
       y = 'Countries',
       title = 'Sub-Saharan African countries with confirmed COVID-19 cases')
```

```{r}
# Overall Africa cases
pd <- df_country
pd <- pd %>% left_join(world_pop) %>%
  filter(sub_region %in% c('Sub-Saharan Africa'))
x = pd %>%
  group_by(date) %>%
  summarise(cases = sum(cases),
            deaths = sum(deaths),
            n = length(country[cases > 0]))

x %>% filter(date == '2020-04-13' | date == '2020-05-13' |  date == max(date) | date == '2020-03-13')
```

```{r}
# Tests
pd <- testing
entity_split <- strsplit(pd$Entity, split = ' - ')
pd$country <- unlist(lapply(entity_split, function(x){x[1]}))
pd$key <- unlist(lapply(entity_split, function(x){x[2]}))
# pd <- pd %>% filter(key == 'tests performed')
pd <- pd %>% left_join(world_pop)# %>%
  # filter(sub_region %in% c('Sub-Saharan Africa', 'Southern Europe', 'Northern Europe', 'Western Europe'))

ggplot(data = pd,
       aes(x = Date,
           y = `Cumulative total`,
           group = country)) +
  geom_line(aes(color = country))

```

```{r}
x = pd %>% group_by(country) %>%
  filter(Date == max(Date)) %>%
  ungroup %>%
  mutate(sub_region = ifelse(sub_region == 'Sub-Saharan Africa',
                             'SSA', 'Other')) %>%
  filter(!is.na(sub_region)) %>%
  arrange(`Cumulative total`)
x$sub_region <- factor(x$sub_region, levels = unique(x$sub_region))
x$country <- factor(x$country, levels = unique(x$country))
ggplot(data = x,
       aes(x = country,
           y = `Cumulative total`)) +
  geom_bar(aes(fill = sub_region), stat = 'identity') +
  theme_simple() +
  scale_fill_manual(name = '',
                    values = c('grey', 'darkorange')) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x= 'Country',
       title = 'Tests administered')
ggsave('~/Desktop/man/3.png')
```

```{r}
pd = esp_df %>%
  left_join(esp_pop) %>%
  mutate(p = deaths_non_cum / pop * 1000000)

ggplot(data = pd,
       aes(x = date,
           y = p)) +
  # geom_step() +
  geom_bar(stat = 'identity',
           fill = 'red',
           alpha = 0.6,
           color = NA) +
  # geom_ribbon(aes(x = date, ymin = 0, ymax = p), data = pd, fill = 'blue') +
  facet_wrap(~ccaa) +
  theme_minimal() +
  labs(x = 'Date',
       y = 'Daily deaths per 1,000,000',
       title = 'Daily deaths per 1,000,000 population')
```

# Daily cases Spain

```{r}
pd <- df_country %>%
  filter(country == 'Spain')

ggplot(data = pd,
       aes(x = date,
           y = cases_non_cum)) +
  geom_bar(stat = 'identity') +
  theme_simple() +
  labs(x = 'Fecha',
       y = 'Casos diarios',
       title = 'Casos confirmados nuevos')
```

# Daily deahts in Spain

```{r}
pd <- df_country %>%
  filter(country == 'Spain')

ggplot(data = pd,
       aes(x = date,
           y = deaths_non_cum)) +
  geom_bar(stat = 'identity') +
  theme_simple() +
  labs(x = 'Fecha',
       y = 'Muertes diarias',
       title = 'Muertes')
```


# Daily cases world / population-adjusted

```{r}
covid19::plot_day_zero(countries = c('Italy', 'Spain', 'US', 'Germany',
                                     'Canada', 'UK', 'Netherlands'
                                     ),
                       ylog = F,
                       day0 = 1,
                       cumulative = F,
                       calendar = T,
                       pop = T,
                       point_alpha = 0,
                       color_var = 'geo')
```

# Cases per pop last week

```{r}
pd <- df_country %>%
  left_join(world_pop %>% dplyr::select(iso, pop)) %>%
  group_by(country) %>%
  mutate(max_date = max(date)) %>%
  mutate(week_ago = max_date - 6) %>%
  # filter(date == max(date)) %>%
  filter(date >= week_ago, date <= max_date) %>%
  group_by(country) %>%
  summarise(y = sum(cases_non_cum),
            pop = dplyr::first(pop),
            date_range = paste0(min(date), ' - ', max(date)),
            yp = sum(cases_non_cum) / dplyr::first(pop) * 1000000) %>%
  ungroup %>%
  filter(pop > 1000000) %>%
  arrange(desc(yp)) %>%
  head(15) %>%
  mutate(country = ifelse(country == 'United Kingdom', 'UK', country))
pd$country <- factor(pd$country, levels = unique(pd$country))

ggplot(data = pd,
       aes(x = country,
           y = yp)) +
  geom_bar(stat = 'identity') +
  theme_simple() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 12)) +
  geom_text(aes(label = round(yp, digits = 0)),
            nudge_y = -50,
            color = 'white') +
  labs(x = '',
       y = 'Cases per 1,000,000 (last 7 days)',
       title = 'New confirmed COVID-19 cases per million population, last 7 days')
```

# Deaths per pop last week


```{r}
pd <- df_country %>%
  left_join(world_pop %>% dplyr::select(iso, pop)) %>%
  group_by(country) %>%
  mutate(max_date = max(date)) %>%
  mutate(week_ago = max_date - 6) %>%
  # filter(date == max(date)) %>%
  filter(date >= week_ago, date <= max_date) %>%
  group_by(country) %>%
  summarise(y = sum(deaths_non_cum),
            pop = dplyr::first(pop),
            date_range = paste0(min(date), ' - ', max(date)),
            yp = sum(deaths_non_cum) / dplyr::first(pop) * 1000000) %>%
  ungroup %>%
  filter(pop > 1000000) %>%
  arrange(desc(yp)) %>%
  head(10) %>%
  mutate(country = ifelse(country == 'United Kingdom', 'UK', country))
pd$country <- factor(pd$country, levels = unique(pd$country))

ggplot(data = pd,
       aes(x = country,
           y = yp)) +
  geom_bar(stat = 'identity') +
  theme_simple() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 12)) +
  geom_text(aes(label = round(yp, digits = 0)),
            nudge_y = -10,
            color = 'white') +
  labs(x = '',
       y = 'Deaths per 1,000,000 (last 7 days)',
       title = 'New confirmed COVID-19 deaths per million population, last 7 days')
```

# Lombardia, Catalonia, Madrid

## New cases in last week

```{r}
pd <- esp_df %>%
  left_join(esp_pop %>% dplyr::select(ccaa, pop)) %>%
  mutate(country = 'Spain') %>%
  bind_rows(
    ita %>% left_join(ita_pop %>% dplyr::select(ccaa, pop)) %>% mutate(country = 'Italy')
  ) %>%
  bind_rows(
    df %>% filter(country == 'US') %>% mutate(ccaa = district) %>% left_join(regions_pop %>% dplyr::select(ccaa, pop)) %>% mutate(country = 'US')) %>%
  group_by(ccaa) %>%
  mutate(max_date = max(date)) %>%
  mutate(week_ago = max_date - 6) %>%
  # filter(date == max(date)) %>%
  filter(date >= week_ago, date <= max_date) %>%
  group_by(ccaa) %>%
  summarise(y = sum(cases_non_cum),
            country = dplyr::first(country),
            pop = dplyr::first(pop),
            date_range = paste0(min(date), ' - ', max(date))) %>%
  ungroup %>%
  mutate(yp = y / pop * 1000000) %>%
  ungroup %>%
  # filter(pop > 1000000) %>%
  arrange(desc(yp)) 

#Get country totals
pd %>%
  group_by(country) %>%
  summarise(y = sum(y, na.rm = T),
            pop = sum(pop, na.rm = T)) %>%
  ungroup %>%
  mutate(yp = y / pop * 1000000)


library(knitr)
library(kableExtra)
pd <- pd %>%
  mutate(yp = round(yp, digits = 1)) %>%
  mutate(Rank = 1:nrow(pd)) %>%
  dplyr::select(Rank, 
                Región = ccaa,
                `Casos nuevos, 7 días` = y,
                Población = pop,
                `Casos nuevos 7 días por millón` = yp)
kable(pd) %>%
  kable_styling("striped", full_width = F) %>%
  column_spec(which(names(pd) == 'Casos nuevos 7 días por millón'), bold = T) %>%
  row_spec(which(pd$`Región` %in% esp_df$ccaa), bold = T, color = "white", background = "#D7261E")
```

# Just Italy and Spain

```{r}
xpd = pd %>% filter(`Región` %in% c(ita$ccaa, esp_df$ccaa))
xpd$Rank <- 1:nrow(xpd)

kable(xpd) %>%
  kable_styling("striped", full_width = F) %>%
  column_spec(which(names(xpd) == 'Casos nuevos 7 días por millón'), bold = T) %>%
  row_spec(which(xpd$`Región` %in% esp_df$ccaa), bold = T, color = "white", background = "#D7261E")
```

```{r}
x = esp_df %>% left_join(esp_pop) %>%
  bind_rows(ita %>% left_join(ita_pop)) %>%
  filter(date == '2020-04-09') %>%
  filter(ccaa %in% c('Madrid',
                     'Cataluña', 'Lombardia')) %>%
  dplyr::select(ccaa, deaths, cases, pop) %>%
  mutate(deathsp = deaths / pop * 100000,
         casesp = cases / pop * 100000)


covid19::plot_day_zero(countries = c('Italy', 'Spain', 'China', 'South Korea', 'Sinagpore'),
                       districts = c('Madrid', #'Hubei',
                     'Cataluña', 'Lombardia'),
                     by_district = T,
                     roll = 7,
                     deaths = F,
                     pop = T,
                     day0 = 0,
                     ylog = F,
                     calendar = T,
                     cumulative = F) +
  labs(x = 'Data',
       y = 'Casos diaris (mitjana mòbil de 7 dies)',
       title = 'Casos diaris per 1.000.000',
       subtitle = 'Mitjana mòbil de 7 dies') 

```

```{r}
covid19::plot_day_zero(countries = c('Italy', 'Spain'),
                     roll = 7,
                     deaths = F,
                     pop = T,
                     day0 = 0,
                     ylog = F,
                     calendar = T,
                     cumulative = F) +
  labs(x = 'Data',
       y = 'Casos diaris per 1.000.000 (mitjana mòbil de 7 dies)',
       title = 'Casos diaris per 1.000.000',
       subtitle = 'Mitjana mòbil de 7 dies') +
  theme(legend.direction = 'horizontal',
        legend.position = 'top')
```

```{r}
covid19::plot_day_zero(countries = c('Italy', 'Spain'),
                     roll = 7,
                     deaths = T,
                     pop = T,
                     day0 = 0,
                     ylog = F,
                     calendar = T,
                     cumulative = F) +
  labs(x = 'Data',
       y = 'Morts diaris per 1.000.000 (mitjana mòbil de 7 dies)',
       title = 'Morts diaris per 1.000.000',
       subtitle = 'Mitjana mòbil de 7 dies') +
  theme(legend.direction = 'horizontal',
        legend.position = 'top')
```

```{r}
covid19::plot_day_zero(countries = c('Italy', 'Spain'),
                       by_district = T,
                       districts = c('Madrid', 'Emilia-Romagna',
                     'Cataluña', 'Lombardia'),
                     roll = 7,
                     deaths = F,
                     pop = T,
                     day0 = 0,
                     ylog = F,
                     calendar = T,
                     cumulative = F) +
  labs(x = 'Data',
       y = 'Casos diaris per 1.000.000 (mitjana mòbil de 7 dies)',
       title = 'Casos diaris per 1.000.000',
       subtitle = 'Mitjana mòbil de 7 dies') +
  theme(legend.direction = 'horizontal',
        legend.position = 'top')
```

<!-- # Personas susceptibles -->

<!-- ```{r} -->
<!-- left <- esp_df %>% -->
<!--   group_by(date) %>% -->
<!--   summarise(cases = sum(cases)) -->
<!-- left$pop <- esp_pop %>% summarise(pop = sum(pop)) %>% .$pop -->
<!-- left$susceptible <- left$pop - (left$cases * 10) -->
<!-- left$p <- left$susceptible / left$pop * 100 -->

<!-- ggplot(data = left, -->
<!--        aes(x = date, -->
<!--            y = p)) + -->
<!--   geom_line() + -->
<!--   ylim(0, 100) -->
<!-- ``` -->


## Asia

```{r}

covid19::plot_day_zero(countries = c('Japan', 'South Korea', 'Singapore', 'Hong Kong'),
                     roll = 7,
                     deaths = F,
                     # pop = T,
                     day0 = 0,
                     ylog = F,
                     calendar = T,
                     cumulative = F) +
    labs(x = 'Data',
       y = 'Casos diaris (mitjana mòbil de 3 dies)',
       title = 'Casos diaris',
       subtitle = 'Mitjana mòbil de 3 dies') +
  facet_wrap(~geo, scales = 'free_y') +
  theme(legend.position = 'none')
```

## Day of week analysis

```{r}
pd <- esp_df %>%
  arrange(date) %>%
  group_by(date) %>%
  summarise(deaths_non_cum = sum(deaths_non_cum),
            cases_non_cum = sum(cases_non_cum)) %>%
  ungroup %>%  
  mutate(dow = weekdays(date)) %>%
  mutate(week = isoweek(date)) %>%
  group_by(week) %>%
  mutate(start_date = min(date)) %>%
  ungroup %>%
  filter(date >= '2020-03-09')
pd$dow <- factor(pd$dow,
                 levels = c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday',
                            'Saturday', 'Sunday'),
                 labels = c('Lunes', 'Martes', 'Miércoles', 'Jueves', 'Viernes',
                            'Sábado', 'Domingo'))
n_cols <- length(unique(pd$start_date))
cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = 'Spectral'))(n_cols)
pd$start_date <- factor(pd$start_date)
ggplot(data = pd,
       aes(x = dow,
           y = cases_non_cum,
           group = week,
           color = start_date)) +
  geom_line(size = 4) +
  geom_point(size = 4) +
  scale_color_manual(name = 'Primer día\nde la semana',
                     values = cols) +
  theme_simple() +
  labs(x = 'Día de la semana',
       y = 'Muertes')
```

## Basque country vs rest of Spain

```{r}
pd <- esp_df %>%
  mutate(geo = ifelse(ccaa == 'País Vasco', 'Basque country', 'Rest of Spain')) %>%
  group_by(geo, date) %>%
  summarise(deaths = sum(deaths)) %>%
  ungroup
pp <- esp_pop %>%
    mutate(geo = ifelse(ccaa == 'País Vasco', 'Basque country', 'Rest of Spain')) %>%
  group_by(geo) %>%
  summarise(pop = sum(pop))
pd <- left_join(pd, pp) %>% mutate(pk = deaths / pop * 100000)

ggplot(data = pd %>% filter(pk > 0.1),
       aes(x = date,
           y = pk,
           color = geo)) +
  geom_line() +
  labs(x = 'Date',
       y = 'Cumulative deaths per 100,000') +
  scale_color_manual(name = '',
                     values = c('red', 'black')) +
  theme_simple() 
```

## Country trajectories, population adjusted

```{r}
countries <- c(
  'Spain',
  'US',
  'France',
  # 'Portugal',
  'Italy',
  'China'
)
districts <- c('Lombardia', 'Cataluña', 
               'New York', 
               # 'Hubei',
               'CyL', 
               'CLM', 
               # 'Washington',
               'La Rioja',
               'Madrid')

plot_day_zero(countries = countries,
              districts = districts,
              ylog = F,
              day0 = 1,
              cumulative = F,
              time_before = 0,
              roll = 3,
              deaths = T,
              pop = T,
              by_district = T,
              point_alpha = 0,
              line_size = 3,
              color_var = 'geo')
```

## Italy and Spain

```{r}
dir.create('/tmp/totesmou')
plot_day_zero(countries = c('Spain', 'Italy', max_date = Sys.Date()-1),
              point_size = 2, calendar = T)
ggsave('/tmp/totesmou/1_italy_vs_spain.png',
       height = 5.6,
       width = 9.6)
```

```{r}
plot_day_zero(countries = c('Spain', 'Italy'),
              point_size = 2, calendar = F)
ggsave('/tmp/totesmou/2_italy_vs_spain_temps_ajustat.png',
       height = 5.6,
       width = 9.6)
```


```{r}
plot_day_zero(countries = c('Spain', 'Italy'),
              point_size = 2, calendar = T, deaths = T, day0 = 10)
```


```{r}
plot_day_zero(countries = c('Spain', 'Italy'),
              point_size = 2, calendar = F, deaths = T, day0 = 10)
```

```{r}
plot_day_zero(countries = c('Spain', 'Italy'),
              point_size = 2, calendar = F, deaths = T, day0 = 10, pop = T)
```

```{r}
plot_day_zero(countries = c('Spain', 'Italy'),
              point_size = 2, calendar = F, deaths = T, day0 = 1, pop = T, roll = 3, roll_fun = 'mean')
```

```{r}
plot_day_zero(countries = c('Spain', 'Italy'),
              districts = c('Cataluña', 'Madrid',
                            'Lombardia', 'Emilia-Romagna'),
              by_district = T,
              day0 = 10,
              pop = T)
```

```{r}
plot_day_zero(countries = c('Spain', 'Italy'),
              districts = c('Cataluña', 'Madrid',
                            'Lombardia', 'Emilia-Romagna'),
              by_district = T,
              deaths = T,
              day0 = 1,
              pop = T)
```

```{r}
plot_day_zero(countries = c('Spain', 'Italy'),
              districts = c('Cataluña', 'Madrid',
                            'Lombardia', 'Emilia-Romagna'),
              by_district = T,
              deaths = T,
              roll = 3,
              roll_fun = 'mean',
              day0 = 1,
              ylog = F,
              pop = T, calendar = T)
```

```{r}
plot_day_zero(countries = c('Spain', 'Italy', 'US'),
              districts = c('Cataluña', 'Madrid', 
                            'New York', 
                            'Lombardia', 'Emilia-Romagna'),
              by_district = T,
              deaths = T,
              roll = 3,
              roll_fun = 'mean',
              day0 = 1,
              pop = T)
```


```{r}
plot_day_zero(countries = c('Spain', 'Italy', 'US'),
              districts = c('Cataluña', 'Madrid', 
                            'New York', 
                            'Lombardia', 'Emilia-Romagna'),
              by_district = T,
              deaths = T,
              # roll = 7,
              roll_fun = 'mean',
              day0 = 1,
              pop = F)
```

```{r}
plot_day_zero(color_var = 'iso', by_district = T,
              deaths = T,
              day0 = 1,
              alpha = 0.6,
              point_alpha = 0,
              calendar = T,
              countries = c('Spain', 'Italy'),
              pop = T)

```


```{r}
place_transform <- function(x){ifelse(x == 'Madrid', 'Madrid',
                                      # ifelse(x == 'Cataluña', 'Cataluña',
                                             'Altres CCAA')
  # )
}
place_transform_ita <- function(x){
  ifelse(x == 'Lombardia', 'Lombardia', 
         # ifelse(x == 'Emilia Romagna', 'Emilia Romagna', 
                'Altres regions italianes')
  # )
}
pd <- esp_df %>% mutate(country = 'Espanya') %>%
  mutate(ccaa = place_transform(ccaa)) %>%
  bind_rows(ita %>% mutate(ccaa = place_transform_ita(ccaa),
                           country = 'Itàlia')) %>%
  group_by(country, ccaa, date) %>% 
  summarise(cases = sum(cases),
            uci = sum(uci),
            deaths = sum(deaths),
            cases_non_cum = sum(cases_non_cum),
            deaths_non_cum = sum(deaths_non_cum),
            uci_non_cum = sum(uci_non_cum)) %>%
  left_join(esp_pop %>%
              mutate(ccaa = place_transform(ccaa)) %>%
              bind_rows(ita_pop %>% mutate(ccaa = place_transform_ita(ccaa))) %>%
              group_by(ccaa) %>%
              summarise(pop = sum(pop))) %>%
  mutate(deaths_non_cum_p = deaths_non_cum / pop * 1000000) %>%
  group_by(country, date) %>%
  mutate(p_deaths_non_cum_country = deaths_non_cum / sum(deaths_non_cum) * 100,
         p_deaths_country = deaths / sum(deaths) * 100)
pd$ccaa <- factor(pd$ccaa,
                  levels = c('Madrid',
                             # 'Cataluña',
                             'Altres CCAA',
                             'Lombardia',
                             # 'Emilia Romagna',
                             'Altres regions italianes'))
cols <- c(
  rev(brewer.pal(n = 3, 'Reds'))[1:2],
  rev(brewer.pal(n = 3, 'Blues'))[1:2]
)

label_data <- pd %>%
  filter(country  %in% c('Itàlia', 'Espanya')) %>%
  group_by(country) %>%
  filter(date == max(date))  %>%
  mutate(label = gsub('\nitalianas', '',  gsub(' ', '\n', ccaa))) %>%
  mutate(x = date - 2,
         y = p_deaths_country + 
           ifelse(p_deaths_country > 50, 10, -9))
ggplot(data = pd %>% group_by(country) %>% mutate(start_day = dplyr::first(date[deaths >=50])) %>% filter(date >= start_day),
       aes(x = date,
           y = p_deaths_country,
           color = ccaa,
           group = ccaa)) +
  # geom_bar(stat = 'identity',
  #          position = position_dodge(width = 0.99)) +
  scale_fill_manual(name = '', values = cols) +
  scale_color_manual(name = '', values = cols) +
  geom_line(size = 2,
            aes(color = ccaa)) +
  # geom_point(size = 3,
  #            aes(color = ccaa)) +
  facet_wrap(~country) +
  # xlim(as.Date('2020-03-09'),
  #      Sys.Date()-1) +
  theme_simple() +
  geom_hline(yintercept = 50, lty = 2, alpha = 0.6) +
  # geom_line(lty = 2, alpha = 0.6) +
  labs(x = 'Data',
       y = 'Percentatge de morts',
       title = 'Percentatge de morts: regió més afectada vs resta del pais',
       subtitle = 'A partir del primer día a cada país amb 50 morts acumulades') +
  theme(legend.position = 'top',
        strip.text = element_text(size = 30),
        legend.text = element_text(size = 10))  +
  guides(color = guide_legend(nrow = 2,
                              keywidth = 5)) +
  geom_text(data = label_data,
            aes(x = x-2,
                y = y,
                label = label,
                color = ccaa),
            size = 7,
            show.legend = FALSE)
```

```{r}
# Same cahrt as previous, but one shared axis
place_transform <- function(x){ifelse(x == 'Madrid', 'Madrid',
                                      # ifelse(x == 'Cataluña', 'Cataluña',
                                             'Rest of Spain')
  # )
}
place_transform_ita <- function(x){
  ifelse(x == 'Lombardia', 'Lombardia', 
         # ifelse(x == 'Emilia Romagna', 'Emilia Romagna', 
                'Rest of Italy')
  # )
}
pd <- esp_df %>% mutate(country = 'Spain') %>%
  mutate(ccaa = place_transform(ccaa)) %>%
  bind_rows(ita %>% mutate(ccaa = place_transform_ita(ccaa),
                           country = 'Italy')) %>%
  group_by(country, ccaa, date) %>% 
  summarise(cases = sum(cases),
            uci = sum(uci),
            deaths = sum(deaths),
            cases_non_cum = sum(cases_non_cum),
            deaths_non_cum = sum(deaths_non_cum),
            uci_non_cum = sum(uci_non_cum)) %>%
  left_join(esp_pop %>%
              mutate(ccaa = place_transform(ccaa)) %>%
              bind_rows(ita_pop %>% mutate(ccaa = place_transform_ita(ccaa))) %>%
              group_by(ccaa) %>%
              summarise(pop = sum(pop))) %>%
  mutate(deaths_non_cum_p = deaths_non_cum / pop * 1000000) %>%
  group_by(country, date) %>%
  mutate(p_deaths_non_cum_country = deaths_non_cum / sum(deaths_non_cum) * 100,
         p_deaths_country = deaths / sum(deaths) * 100)
pd$ccaa <- factor(pd$ccaa,
                  levels = c('Madrid',
                             # 'Cataluña',
                             'Rest of Spain',
                             'Lombardia',
                             # 'Emilia Romagna',
                             'Rest of Italy'))
cols <- c(
  rev(brewer.pal(n = 3, 'Reds'))[1:2],
  rev(brewer.pal(n = 3, 'Blues'))[1:2]
)

label_data <- pd %>%
  filter(country  %in% c('Italy', 'Spain')) %>%
  group_by(country) %>%
  filter(date == max(date))  %>%
  # mutate(label = gsub('\nitalianas', '',  gsub(' ', '\n', ccaa))) %>%
  mutate(x = date - 2,
         y = p_deaths_country + 
           ifelse(p_deaths_country > 50, 10, -9)) %>%
  ungroup
ggplot(data = pd %>% group_by(country) %>% mutate(start_day = dplyr::first(date[deaths >=30])) %>% filter(date >= start_day),
       aes(x = date,
           y = p_deaths_country,
           color = ccaa,
           group = ccaa)) +
  # geom_bar(stat = 'identity',
  #          position = position_dodge(width = 0.99)) +
  scale_fill_manual(name = '', values = cols) +
  scale_color_manual(name = '', values = cols) +
  geom_line(size = 2,
            aes(color = ccaa)) +
  # geom_point(size = 3,
  #            aes(color = ccaa)) +
  # facet_wrap(~country, scales = 'free_x') +
  # xlim(as.Date('2020-03-09'),
  #      Sys.Date()-1) +
  theme_simple() +
  geom_hline(yintercept = 50, lty = 2, alpha = 0.6) +
  # geom_line(lty = 2, alpha = 0.6) +
  labs(x = 'Date',
       y = 'Percentage of deaths',
       title = 'Percentage of country\'s cumulative COVID-19 deaths by geography',
       subtitle = 'Starting at first day with 50 or more cumulative deaths') +
  theme(legend.position = 'top',
        strip.text = element_text(size = 30),
        legend.text = element_text(size = 10))  +
  guides(color = guide_legend(nrow = 2,
                              keywidth = 5))# +
  # geom_text(data = label_data,
  #           aes(x = x-2,
  #               y = y,
  #               label = label,
  #               color = ccaa),
  #           size = 7,
  #           show.legend = FALSE)
```

Same chart as above but absolute numbers


```{r}
# Same cahrt as previous, but one shared axis
place_transform <- function(x){ifelse(x == 'Madrid', 'Madrid',
                                      # ifelse(x == 'Cataluña', 'Cataluña',
                                             'Rest of Spain')
  # )
}
place_transform_ita <- function(x){
  ifelse(x == 'Lombardia', 'Lombardia', 
         # ifelse(x == 'Emilia Romagna', 'Emilia Romagna', 
                'Rest of Italy')
  # )
}
pd <- esp_df %>% mutate(country = 'Spain') %>%
  mutate(ccaa = place_transform(ccaa)) %>%
  bind_rows(ita %>% mutate(ccaa = place_transform_ita(ccaa),
                           country = 'Italy')) %>%
  group_by(country, ccaa, date) %>% 
  summarise(cases = sum(cases),
            uci = sum(uci),
            deaths = sum(deaths),
            cases_non_cum = sum(cases_non_cum),
            deaths_non_cum = sum(deaths_non_cum),
            uci_non_cum = sum(uci_non_cum)) %>%
  left_join(esp_pop %>%
              mutate(ccaa = place_transform(ccaa)) %>%
              bind_rows(ita_pop %>% mutate(ccaa = place_transform_ita(ccaa))) %>%
              group_by(ccaa) %>%
              summarise(pop = sum(pop))) %>%
  mutate(deaths_non_cum_p = deaths_non_cum / pop * 1000000) %>%
  group_by(country, date) %>%
  mutate(p_deaths_non_cum_country = deaths_non_cum / sum(deaths_non_cum) * 100,
         p_deaths_country = deaths / sum(deaths) * 100)
pd$ccaa <- factor(pd$ccaa,
                  levels = rev(c('Madrid',
                             # 'Cataluña',
                             'Rest of Spain',
                             'Lombardia',
                             # 'Emilia Romagna',
                             'Rest of Italy')))
cols <- c(
  rev(brewer.pal(n = 3, 'Reds'))[1:2],
  rev(brewer.pal(n = 3, 'Blues'))[1:2]
)

label_data <- pd %>%
  filter(country  %in% c('Italy', 'Spain')) %>%
  group_by(country) %>%
  filter(date == max(date))  %>%
  # mutate(label = gsub('\nitalianas', '',  gsub(' ', '\n', ccaa))) %>%
  mutate(x = date - 2,
         y = p_deaths_country + 
           ifelse(p_deaths_country > 50, 10, -9)) %>%
  ungroup

# Get moving
ma <- function(x, n = 2){
    
    if(length(x) >= n){
      stats::filter(x, rep(1 / n, n), sides = 1)
    } else {
      new_n <- length(x)
      stats::filter(x, rep(1 / new_n, new_n), sides = 1)
    }
}


ggplot(data = pd %>% group_by(country) %>% 
         mutate(start_day = dplyr::first(date[deaths >=1])) %>% 
         filter(date >= start_day) %>% 
         mutate(days_since = as.numeric(date - start_day)) %>%
         ungroup %>% arrange(date) %>%
         group_by(country, ccaa) %>%
         mutate(rolling_average = ma(deaths_non_cum, n = 3)) %>%
         ungroup,
       aes(x = date,
           y = rolling_average,
           color = ccaa,
           group = ccaa)) +
  # geom_bar(stat = 'identity',
  #          position = position_dodge(width = 0.99)) +
  scale_fill_manual(name = '', values = cols) +
  scale_color_manual(name = '', values = cols) +
  geom_line(size = 2,
            aes(color = ccaa)) +
  # geom_point(size = 3,
  #            aes(color = ccaa)) +
  # scale_y_log10(limits = c(1.5, 1000)) +
  # scale_y_log10() +
  facet_wrap(~country) +
  # xlim(as.Date('2020-03-09'),
  #      Sys.Date()-1) +
  theme_simple() +
  # geom_hline(yintercept = 50, lty = 2, alpha = 0.6) +
  # geom_line(lty = 2, alpha = 0.6) +
  labs(x = 'Date',
       y = 'Deaths (log-scale)',
       title = 'Daily COVID-19 deaths by geography',
       subtitle = '3-day rolling average') +
  theme(legend.position = 'top',
        plot.title = element_text(size = 30),
        plot.subtitle = element_text(size = 24),
        strip.text = element_text(size = 30, hjust = 0.5),
        legend.text = element_text(size = 20))  +
  guides(color = guide_legend(nrow = 2,
                              keywidth = 5))
ggsave('~/Desktop/madlom.png')
```



```{r}
roll_curve <- function(data,
                       n = 4,
                       deaths = FALSE,
                       scales = 'fixed',
                       nrow = NULL,
                       ncol = NULL,
                       pop = FALSE){

  # Create the n day rolling avg
  ma <- function(x, n = 5){
    
    if(length(x) >= n){
      stats::filter(x, rep(1 / n, n), sides = 1)
    } else {
      new_n <- length(x)
      stats::filter(x, rep(1 / new_n, new_n), sides = 1)
    }
    
    
  }
  
  pd <- data
  if(deaths){
    pd$var <- pd$deaths_non_cum
  } else {
    pd$var <- pd$cases_non_cum
  }
  
  if('ccaa' %in% names(pd)){
    pd$geo <- pd$ccaa
  } else {
    pd$geo <- pd$iso
  }
  
  if(pop){
    # try to get population
    if(any(pd$geo %in% unique(esp_df$ccaa))){
      right <- esp_pop
      right_var <- 'ccaa'
    } else {
      right <- world_pop
      right_var <- 'iso'
    }
    pd <- pd %>% left_join(right %>% dplyr::select(all_of(right_var), pop),
                           by = c('geo' = right_var)) %>%
      mutate(var = var / pop * 100000)
  }
  
  pd <- pd %>%
    arrange(date) %>%
    group_by(geo) %>%
    mutate(with_lag = ma(var, n = n))
  
  
  ggplot() +
    geom_bar(data = pd,
         aes(x = date,
             y = var),
             stat = 'identity',
         fill = 'grey',
         alpha = 0.8) +
    geom_area(data = pd,
              aes(x = date,
                  y = with_lag),
              color = 'red',
              fill = 'red',
              alpha = 0.6) +
    facet_wrap(~geo, scales = scales, nrow = nrow, ncol = ncol) +
    theme_simple() +
    labs(x = '',
         y = ifelse(deaths, 'Deaths', 'Cases'),
         title = paste0('Daily (non-cumulative) ', ifelse(deaths, 'deaths', 'cases'),
                        ifelse(pop, ' (per 100,000)', '')),
         subtitle = paste0('Data as of ', max(data$date),
                           '.\nRed line = ', n, ' day rolling average. Grey bar = day-specific value.')) +
    theme(axis.text.x = element_text(size = 7,
                                     angle = 90,
                                     hjust = 0.5, vjust = 1)) #+
    # scale_x_date(name ='',
    #              breaks = sort(unique(pd$date)),
    #              labels = format(sort(unique(pd$date)), '%b %d'))
    # scale_y_log10()
}
this_ccaa <- 'Cataluña'
plot_data <- esp_df %>% mutate(geo = ccaa) %>% filter(ccaa == this_ccaa)
roll_curve(plot_data, scales = 'fixed')  + theme(strip.text = element_text(size = 30))

plot_data <- esp_df %>% mutate(geo = ccaa) %>% filter(ccaa == this_ccaa)
roll_curve(plot_data, deaths = T, scales = 'fixed') + theme(strip.text = element_text(size = 30))

```


```{r}
african_countries <-  world_pop$country[world_pop$sub_region %in% c('Sub-Saharan Africa')]

pd <- plot_day_zero(countries = c(african_countries),
                    day0 = 1,
                    max_date = Sys.Date() - 46,
                    ylog = F) +
  ylim(0, 5000)
pd

pd <- plot_day_zero(countries = c(african_countries),
                    day0 = 1,
                    max_date = Sys.Date(),
                    ylog = F) +
  ylim(0, 5000)  
pd

pd <- plot_day_zero(countries = c(african_countries),
                    day0 = 1,
                    calendar = T,
                    max_date = Sys.Date(),
                    ylog = T) + theme(legend.position = 'none')
pd

pd <- plot_day_zero(countries = c(african_countries),
                    day0 = 10,
                    max_date = Sys.Date() - 13)
pd


pd <- plot_day_zero(countries = c(african_countries),
                    day0 = 10,
                    max_date = Sys.Date() - 6)
pd


pd <- plot_day_zero(countries = c(african_countries),
                    day0 = 10,
                    max_date = Sys.Date(),
                    ylog = F)
pd



latam_countries <-  world_pop$country[world_pop$sub_region %in% c('Latin America and the Caribbean')]
latam_countries <- latam_countries[!latam_countries %in% c('Guyana')]

pd <- plot_day_zero(countries = c(latam_countries),
                    day0 = 10,
                    max_date = Sys.Date() - 13)
pd


pd <- plot_day_zero(countries = c(latam_countries),
                    day0 = 10,
                    max_date = Sys.Date() - 6)
pd


pd <- plot_day_zero(countries = c(latam_countries),
                    day0 = 10,
                    max_date = Sys.Date())
pd
```

Latin America and Africa vs Europe


```{r}
isos <- sort(unique(world_pop$sub_region))
keep_countries <- world_pop$country[world_pop$sub_region %in% c('Latin America and the Caribbean', 'Sub-Saharan Africa') |
                                      world_pop$region %in% 'Europe']
keep_countries <- keep_countries[!keep_countries %in% c('Guyana')]
pd <- df_country %>% ungroup %>%
  filter(country %in% keep_countries) %>%
  dplyr::select(-country) %>%
  left_join(world_pop) %>%
  group_by(iso) %>%
  mutate(day0 = min(date[cases >= 10])) %>%
  ungroup %>%
  mutate(days_since = date - day0) %>%
  filter(days_since >= 0)


cols <- c( 'black')
g <- ggplot(data = pd,
       aes(x = days_since,
           y = cases,
           group = country,
           color = region)) +
  geom_line(data = pd %>% filter(region == 'Europe'),
            alpha = 0.6) +
  scale_y_log10() +
  scale_color_manual(name = '', values = cols) +
  theme_simple() +
  labs(x = 'Days since first day at 10 cases') +
  theme(legend.position = 'top')
g


cols <- c('darkred', 'black')

g + 
    geom_line(data = pd %>% filter(region == 'Africa'),
            # alpha = 1,
            size = 1.5,
            alpha = 0.8) +
    scale_y_log10() +
    scale_color_manual(name = '', values = cols) 


cols <- c( 'darkorange', 'black')
g + 
    geom_line(data = pd %>% filter(sub_region == 'Latin America and the Caribbean'),
            # alpha = 1,
            size = 1.5,
            alpha = 0.8) +
    scale_color_manual(name = '', values = cols) 


cols <- c('darkred', 'darkorange', 'black')
g + 
    geom_line(data = pd %>% filter(sub_region != 'Europe'),
            # alpha = 1,
            size = 1.5,
            alpha = 0.8) +
    scale_color_manual(name = '', values = cols) 
```

```{r, eval = F}
# Assets
pyramid_dir <- '../../data-raw/pyramids/'
pyramid_files <- dir(pyramid_dir)
out_list <- list()
for(i in 1:length(pyramid_files)){
  out_list[[i]] <- read_csv(paste0(pyramid_dir, pyramid_files[i])) %>%
    mutate(region = gsub('.csv', '', pyramid_files[i]))
}
pyramid <- bind_rows(out_list)
make_pyramid <- function(the_region = 'AFRICA-2019'){
  sub_data <- pyramid %>% filter(region == the_region)
  sub_data$Age <- factor(sub_data$Age, levels = sub_data$Age)
  sub_data <- tidyr::gather(sub_data, key, value, M:F)
  ggplot(data = sub_data,
         aes(x = Age,
             y = value,
             fill = key)) +
    geom_bar(stat = 'identity',
             position = position_dodge(),
             alpha = 0.6,
             color = 'black') +
    scale_fill_manual(name = '', values = c('darkorange', 'lightblue')) +
    theme_simple() +
    labs(x = 'Age group',
         y = 'Population') +
    theme(legend.position = 'top')
}
make_pyramid_overlay <- function(){
  sub_data <- pyramid %>% filter(region %in% c('EUROPE-2019',
                                               'AFRICA-2019',
                                               'LATIN AMERICA AND THE CARIBBEAN-2019')) %>%
    mutate(region = gsub('-2019', '', region))
   sub_data$Age <- factor(sub_data$Age, levels = unique(sub_data$Age))
  sub_data <- tidyr::gather(sub_data, key, value, M:F) %>%
    group_by(Age, region) %>%
    summarise(value = sum(value)) %>%
    ungroup %>%
    group_by(region) %>%
    mutate(p = value / sum(value) * 100) %>%
    ungroup
  ggplot(data = sub_data,
         aes(x = Age,
             y = p,
             color = region,
             group = region,
             fill = region)) +
    geom_area(position = position_dodge(),
              alpha = 0.6) +
    scale_fill_manual(name = '',
                      values = c('darkred', 'darkorange', 'black')) +
    scale_color_manual(name = '',
                      values = c('darkred', 'darkorange', 'black')) +
    theme_simple() +
    theme(legend.position = 'top') +
    labs(x = 'Age group', y = 'Percentage')
}

make_pyramid(the_region = 'Spain-2019') + labs(title = 'Spain')
make_pyramid(the_region = 'Italy-2019') + labs(title = 'Italy')

make_pyramid(the_region = 'EUROPE-2019') + labs(title = 'Europe')
make_pyramid(the_region = 'Kenya-2019') + labs(title = 'Kenya')
make_pyramid(the_region = 'Mozambique-2019') + labs(title = 'Mozambique')
make_pyramid(the_region = 'Tanzania-2019') + labs(title = 'Tanzania')

make_pyramid(the_region = 'Guatemala-2019') + labs(title = 'Guatemala')

make_pyramid(the_region = 'AFRICA-2019') + labs(title = 'Africa')

make_pyramid(the_region = 'LATIN AMERICA AND THE CARIBBEAN-2019') + labs(title = 'Latin America and the Caribbean')

make_pyramid_overlay() + labs(title = 'Population distribution by region')

pyramid <- pyramid %>%
  mutate(old = Age %in% c('80-84', '85-89', '90-94','95-99', '100+'))

pyramid %>%
  group_by(region, old) %>%
  summarise(n = sum(M) + sum(F)) %>%
  ungroup %>%
  group_by(region) %>%
  mutate(p = n / sum(n) * 100) %>%
  filter(old)

```

```{r}
plot_day_zero(countries = c('South Africa', 'Spain'), day0 = 1, calendar = T)
```

```{r}
plot_day_zero(countries = c('Kenya', 'Italy'), day0 = 10, calendar = F)
```


## Pics

```{r}
plot_day_zero(countries = c('China', 'Italy', 'Spain'),
              districts = c('Hubei', 'Lombardia', 'Cataluña', 'Madrid'),
              by_district = T,
              point_alpha = 0,
              day0 = 5,
              pop = F,
              deaths = T,
              ylog = T,
              calendar = F,
              roll = 5)
```

## Map of portugal, france, spain

```{r}
# cat_transform <- function(x){ifelse(x == 'Catalunya', 'Cataluña', x)}
cat_transform <- function(x){return(x)}
pd <- por_df %>% mutate(country = 'Portugal') %>%
  bind_rows(esp_df %>% mutate(country = 'Spain')) %>%
  bind_rows(fra_df %>% mutate(country = 'France')) %>%
  bind_rows(ita %>% mutate(country = 'Italy')) %>%
  bind_rows(
    df %>% filter(country == 'Andorra') %>%
      mutate(ccaa = 'Andorra')
  )
keep_date = pd %>% group_by(country) %>% summarise(max_date= max(date)) %>% summarise(x = min(max_date)) %>% .$x
pd <- pd %>%
  mutate(ccaa = cat_transform(ccaa)) %>%
  group_by(ccaa) %>%
  filter(date == keep_date) %>%
  # filter(date == '2020-03-27') %>%
  ungroup %>%
  dplyr::select(date, ccaa, deaths, deaths_non_cum,
                cases, cases_non_cum) %>%
  left_join(regions_pop %>%
              bind_rows(
                world_pop %>% filter(country == 'Andorra') %>% dplyr::mutate(ccaa = country) %>%
                  dplyr::select(-region, -sub_region)
              )) %>%
  mutate(cases_per_million = cases / pop * 1000000,
         deaths_per_million = deaths / pop * 1000000) %>%
  mutate(cases_per_million_non_cum = cases_non_cum / pop * 1000000,
         deaths_per_million_non_cum = deaths_non_cum / pop * 1000000)

map_esp1 <- map_esp
map_esp1@data <- map_esp1@data %>% dplyr::select(ccaa)
map_fra1 <- map_fra
map_fra1@data <- map_fra1@data %>% dplyr::select(ccaa = NAME_1)
map_por1 <- map_por
map_por1@data <- map_por1@data %>% dplyr::select(ccaa = CCDR)
map_ita1 <- map_ita 
map_ita1@data <- map_ita1@data %>% dplyr::select(ccaa)
map_and1 <- map_and
map_and1@data <- map_and1@data %>% dplyr::select(ccaa = NAME_0)


map <- 
  rbind(map_esp1,
        map_por1,
        map_fra1,
        map_ita1,
        map_and1)

# Remove areas not of interest
centroids <- coordinates(map)
centroids <- data.frame(centroids)
names(centroids) <- c('x', 'y')
centroids$ccaa <- map@data$ccaa
centroids <- left_join(centroids, pd)
# map <- map_sp <- map[centroids$y >35 & centroids$x > -10 &
#              centroids$x < 8 & (centroids$y < 43  | map@data$ccaa %in% c('Occitanie', 'Nouvelle-Aquitaine') |
#                                   map@data$ccaa %in% esp_df$ccaa),]
# map_sp <- map <-  map[centroids$x > -10 & centroids$y <47,]
map_sp <- map <-  map[centroids$x > -10 & centroids$y <77,]

# map_sp <- map <-  map[centroids$x > -10,]

# fortify
map <- fortify(map, region = 'ccaa')

# join data
map$ccaa <- map$id
map <- left_join(map, pd)
var <- 'deaths_per_million'
map$var <- as.numeric(unlist(map[,var]))
centroids <- centroids[,c('ccaa', 'x', 'y', var)]
centroids <- centroids %>%
  filter(ccaa %in% map_sp@data$ccaa)

# cols <- rev(RColorBrewer::brewer.pal(n = 9, name = 'Spectral'))
# cols <- c('#A16928','#bd925a','#d6bd8d','#edeac2','#b5c8b8','#79a7ac','#2887a1')
# cols <- c('#009392','#39b185','#9ccb86','#e9e29c','#eeb479','#e88471','#cf597e')
# cols <- c('#008080','#70a494','#b4c8a8','#f6edbd','#edbb8a','#de8a5a','#ca562c')
cols <- rev(colorRampPalette(c('darkred', 'red', 'darkorange', 'orange', 'yellow', 'lightblue'))(10))
g <- ggplot(data = map,
         aes(x = long,
             y = lat,
             group = group)) +
    geom_polygon(aes(fill = var),
                 lwd = 0.3,
                 # color = 'darkgrey',
                 color = NA,
                 size = 0.6) +
      scale_fill_gradientn(name = '',
                           colours = cols) +
  # scale_fill_() +
  ggthemes::theme_map() +
  theme(legend.position = 'bottom',
        plot.title = element_text(size = 16)) +
  guides(fill = guide_colorbar(direction= 'horizontal',
                               barwidth = 50,
                               barheight = 1,
                               label.position = 'bottom')) +
  labs(title = 'Cumulative COVID-19 deaths per million population, western Mediterranean',
       subtitle = paste0('Data as of ', format(max(pd$date), '%B %d, %Y'))) +
  geom_text(data = centroids,
            aes(x = x,
                y = y,
                group = NA,
                label = paste0(ccaa, '\n',
                               round(deaths_per_million, digits = 2))),
            alpha = 0.8,
            size = 3)
g
ggsave('/tmp/map_with_borders.png',
       height = 8, width = 13)
```



## Animation, Portugal, France, Spain, Italy

```{r}
dir.create('/tmp/animation_map/')
pd <- por_df %>% mutate(country = 'Portugal') %>%
  bind_rows(esp_df %>% mutate(country = 'Spain')) %>%
  bind_rows(fra_df %>% mutate(country = 'France')) %>%
  bind_rows(ita %>% mutate(country = 'Italy')) %>%
  bind_rows(
    df %>% filter(country == 'Andorra') %>%
      mutate(ccaa = 'Andorra')
  )
pd %>% group_by(country) %>% summarise(max_date= max(date))
unique_dates <- sort(unique(pd$date))
unique_dates <- unique_dates[unique_dates >= '2020-03-01']
popper <- regions_pop %>%
                bind_rows(
                  world_pop %>% filter(country == 'Andorra') %>% dplyr::mutate(ccaa = country) %>%
                    dplyr::select(-region, -sub_region)
                )


map_esp1 <- map_esp
map_esp1@data <- map_esp1@data %>% dplyr::select(ccaa)
map_fra1 <- map_fra
map_fra1@data <- map_fra1@data %>% dplyr::select(ccaa = NAME_1)
map_por1 <- map_por
map_por1@data <- map_por1@data %>% dplyr::select(ccaa = CCDR)
map_ita1 <- map_ita 
map_ita1@data <- map_ita1@data %>% dplyr::select(ccaa)
map_and1 <- map_and
map_and1@data <- map_and1@data %>% dplyr::select(ccaa = NAME_0)


map <- 
  rbind(map_esp1,
        map_por1,
        map_fra1,
        map_ita1,
        map_and1)

# Remove areas not of interest
centroids <- coordinates(map)
centroids <- data.frame(centroids)
names(centroids) <- c('x', 'y')
centroids$ccaa <- map@data$ccaa
# map <- map_sp <- map[centroids$y >35 & centroids$x > -10 &
#              # centroids$x < 8 &
#                (centroids$y < 43  | map@data$ccaa %in% c('Occitanie', 'Nouvelle-Aquitaine') |
#                                   map@data$ccaa %in% esp_df$ccaa),]
# map_sp <- map <-  map[centroids$x > -10 & centroids$y <47,]
map_sp <- map <-  map[centroids$x > -10 & centroids$y <477,]


# fortify
map <- fortify(map, region = 'ccaa')



for(i in 1:length(unique_dates)){
  this_date <- unique_dates[i]
    today_map <- map
    today_centroids <- centroids
    today_pd <- pd

  today_pd <- today_pd %>%
      mutate(ccaa = cat_transform(ccaa)) %>%
    group_by(ccaa) %>%
    # filter(date == max(date)) %>%
    filter(date == this_date) %>%
    ungroup %>%
    dplyr::select(date, ccaa, deaths, deaths_non_cum,
                  cases, cases_non_cum) %>%
    left_join(popper) %>%
    mutate(cases_per_million = cases / pop * 1000000,
           deaths_per_million = deaths / pop * 1000000) %>%
    mutate(cases_per_million_non_cum = cases_non_cum / pop * 1000000,
           deaths_per_million_non_cum = deaths_non_cum / pop * 1000000)
  
  today_centroids <- left_join(today_centroids, today_pd)

  
  # join data
  today_map$ccaa <- today_map$id
  today_map <- left_join(today_map, today_pd)
  var <- 'deaths_per_million'
  today_map$var <- as.numeric(unlist(today_map[,var]))
  today_map$var <- ifelse(is.na(today_map$var), 0, today_map$var)
  today_centroids <- today_centroids[,c('ccaa', 'x', 'y', var)]
  today_centroids <- today_centroids %>%
    filter(ccaa %in% today_map$ccaa)
  today_centroids$var <- today_centroids[,var]
  today_centroids$var <- ifelse(is.na(today_centroids$var), 0, today_centroids$var)

  
  # cols <- rev(RColorBrewer::brewer.pal(n = 9, name = 'Spectral'))
  # cols <- c('#A16928','#bd925a','#d6bd8d','#edeac2','#b5c8b8','#79a7ac','#2887a1')
  # cols <- c('#009392','#39b185','#9ccb86','#e9e29c','#eeb479','#e88471','#cf597e')
  # cols <- c('#008080','#70a494','#b4c8a8','#f6edbd','#edbb8a','#de8a5a','#ca562c')
  cols <- rev(colorRampPalette(c('darkred', 'red', 'darkorange', 'orange', 'yellow', 'white'))(17))
  g <- ggplot(data = today_map,
           aes(x = long,
               y = lat,
               group = group)) +
      geom_polygon(aes(fill = var),
                   lwd = 0.3,
                   # color = 'darkgrey',
                   color = NA,
                   size = 0.6) +
        scale_fill_gradientn(name = '',
                             colours = cols,
                             breaks = seq(0, 1100, 50),
                             limits = c(0, 1100)) +
    # scale_fill_() +
    ggthemes::theme_map() +
    theme(legend.position = 'right',
          plot.title = element_text(size = 24)) +
    guides(fill = guide_colorbar(direction= 'vertical',
                                 barwidth = 1,
                                 barheight = 30,
                                 label.position = 'left')) +
    labs(subtitle = 'Cumulative COVID-19 deaths per million population',
         title = paste0(format(this_date, '%B %d, %Y'))) +
    geom_text(data = today_centroids,
              aes(x = x,
                  y = y,
                  group = NA,
                  label = paste0(ccaa, '\n',
                                 round(var, digits = 2))),
              alpha = 0.8,
              size = 1.5)
  
  ggsave(paste0('/tmp/animation_map/', this_date, '.png'),
         plot = g,
         height = 6, width = 9)
}
```

```
# Command line
cd /tmp/animation_map
mogrify -resize 50% *.png
convert -delay 20 -loop 0 *.png result.gif

```

## Deaths overall over time Spain

```{r}
df_country %>% filter(country == 'Spain') %>% arrange(date) %>% tail
```

## Deaths yesterday

```{r}
pd <- df_country
pd$value <- pd$deaths_non_cum
the_date <- Sys.Date() - 1
pd <- pd %>%
  filter(date == the_date) %>%
  dplyr::select(country, iso, cases, cases_non_cum,
                deaths, value) %>%
  dplyr::arrange(desc(value)) %>%
  left_join(world_pop %>% dplyr::select(-country)) %>%
  mutate(value_per_million = value / pop * 1000000) #%>% 
  # arrange(desc(value_per_million))
pd <- pd[1:10,]
pd$country <- factor(pd$country, levels = pd$country)
ggplot(data = pd,
       aes(x = country,
           y = value)) +
  geom_bar(stat = 'identity',
           fill = 'black') +
  theme_simple() +
  geom_text(aes(label = value),
            nudge_y = -20,
            size = 4,
            color = 'white') +
  labs(title = paste0('Confirmed COVID-19 deaths on ', the_date),
       x = '', y = '')
pd

```

Deaths per million yesterday per CCAA

```{r}
pd <- esp_df
pd$value <- pd$deaths_non_cum
the_date <- max(pd$date)
pd <- pd %>%
  filter(date == max(date)) %>%
  dplyr::select(ccaa, cases, cases_non_cum,
                deaths, value) %>%
  dplyr::arrange(desc(value)) %>%
  left_join(esp_pop)%>%
  mutate(value_per_million = value / pop * 1000000) #%>% 
  # arrange(desc(value_per_million))
pd <- pd[1:10,]
pd$ccaa <- factor(pd$ccaa, levels = pd$ccaa)
ggplot(data = pd,
       aes(x = ccaa,
           y = value)) +
  geom_bar(stat = 'identity',
           fill = 'black') +
  theme_simple() +
  geom_text(aes(label = value),
            nudge_y = -20,
            size = 4,
            color = 'white')
pd
```

## Deaths yesterday animation

```{r, eval = T}
dir.create('/tmp/animation_deaths')
dates <- seq(as.Date('2020-03-17'), (Sys.Date()-1), by = 1)
for(i in 1:length(dates)){
  this_date <- dates[i]
  pd <- df_country
  pd$value <- pd$deaths_non_cum
  pd <- pd %>%
    filter(date == max(this_date)) %>%
    dplyr::select(country, cases, cases_non_cum,
                  deaths, value) %>%
    dplyr::arrange(desc(value))
  pd <- pd[1:10,]
  pd <- pd %>% filter(value > 0)
  pd$country <- gsub(' ', '\n', pd$country)
  pd$country <- factor(pd$country, levels = pd$country)
  pd$color_var <- pd$country == 'Spain'
  if('Spain' %in% pd$country){
    cols <- rev(c('darkred', 'black'))
  } else {
    cols <- 'black'
  }
  g <- ggplot(data = pd,
         aes(x = country,
             y = value)) +
    geom_bar(stat = 'identity',
             aes(fill = color_var),
             alpha = 0.8,
             show.legend = FALSE) +
    theme_simple() +
    geom_text(aes(label = value),
              nudge_y = max(pd$value) * .05,
              size = 5,
              color = 'black') +
    labs(title = 'Daily (non-cumulative) COVID-19 deaths',
         subtitle = format(this_date, '%B %d')) +
    labs(x = 'Country',
         y = 'Deaths') +
    theme(axis.text = element_text(size = 15),
          plot.subtitle = element_text(size = 20)) +
    scale_fill_manual(name ='',
                      values = cols) +
    ylim(0, 900)
  ggsave(filename = paste0('/tmp/animation_deaths/', this_date, '.png'),
         g)
}
# Command line
```

```
cd /tmp/animation_deaths
mogrify -resize 50% *.png
convert -delay 50 -loop 0 *.png result.gif
```

## Heatmap

```{r, fig.width = 10, fig.height = 8}
pd <- by_country <-  esp_df %>% mutate(country = 'Spain') %>% 
  bind_rows(ita %>% mutate(country = 'Italy')) %>%
  bind_rows(por_df %>% mutate(country = 'Portugal')) %>%
  bind_rows(fra_df %>% mutate(country = 'France'))
pd$value <- pd$deaths_non_cum
max_date <- pd %>% group_by(country) %>% summarise(d = max(date)) %>% ungroup %>% summarise(d = min(d)) %>% .$d
# pd$value <- ifelse(is.na(pd$value), 0, pd$value)
left <- expand.grid(date = seq(min(pd$date),
                               max(pd$date),
                               by = 1),
                    ccaa = sort(unique(pd$ccaa)))
right <- pd %>% dplyr::select(date, ccaa, value)
pd <- left_join(left, right) %>% mutate(value = ifelse(is.na(value), NA, value))
pd <- left_join(pd, by_country %>% dplyr::distinct(country, ccaa)) %>%
  filter(date <= max_date) %>%
  filter(value > 0)
the_limits <- c(1, 600)
the_breaks <- c(1, seq(100, 600, length = 6)) #seq(0, 600, length = 7)
pd$ccaa <- factor(pd$ccaa, levels = rev(unique(sort(pd$ccaa))))
ggplot(data = pd,
       aes(x = date,
           y = ccaa,
           color = value,
           size = value)) +
  # geom_tile(color = 'white') +
  geom_point(alpha = 0.8) +
  # geom_tile() +
  scale_color_gradientn(colors = rev(colorRampPalette(brewer.pal(n = 8, 'Spectral'))(5)),
                        name = '',
                        limits = the_limits,
                        breaks = the_breaks) +
    # scale_fill_gradientn(colors = rev(colorRampPalette(brewer.pal(n = 8, 'Spectral'))(5)),
    #                     name = '',
    #                     limits = the_limits,
    #                     breaks = the_breaks) +
  scale_size_area(name = '', limits = the_limits, breaks = the_breaks, max_size = 10) +
  
  theme_simple() +
  facet_wrap(~country, scales = 'free_y') +
  theme(strip.text = element_text(size = 20),
        axis.title = element_blank(),
        axis.text = element_text(size = 10),
        axis.text.x = element_text(size = 12)) +
  guides(color = guide_legend(), size = guide_legend()) +
  labs(title = 'Daily (non-cumulative) COVID-19 deaths by sub-state regions',
       caption = paste0('Data as of ', max_date))
ggsave('/tmp/1.png',
       width = 10,
       height = 8)
```




## Heatmap per population

```{r, fig.width = 12, fig.height = 8}
pd <- by_country <-  esp_df %>% mutate(country = 'Spain') %>%  bind_rows(ita %>% mutate(country = 'Italy'))
poppy <- bind_rows(ita_pop, esp_pop)
pd <- pd %>% left_join(poppy)
pd$value <- pd$deaths_non_cum / pd$pop * 1000000
max_date <- pd %>% group_by(country) %>% summarise(d = max(date)) %>% ungroup %>% summarise(d = min(d)) %>% .$d
# pd$value <- ifelse(is.na(pd$value), 0, pd$value)
left <- expand.grid(date = seq(min(pd$date),
                               max(pd$date),
                               by = 1),
                    ccaa = sort(unique(pd$ccaa)))
right <- pd %>% dplyr::select(date, ccaa, value)
pd <- left_join(left, right) %>% mutate(value = ifelse(is.na(value), NA, value))
pd <- left_join(pd, by_country %>% dplyr::distinct(country, ccaa)) %>%
  filter(date <= max_date) %>%
  filter(value > 0)
the_limits <- c(1, 60)
the_breaks <- c(1, seq(10, 60, length = 6)) #seq(0, 600, length = 7)
pd$ccaa <- factor(pd$ccaa, levels = rev(unique(sort(pd$ccaa))))
ggplot(data = pd,
       aes(x = date,
           y = ccaa,
           color = value,
           size = value)) +
  # geom_tile(color = 'white') +
  geom_point(alpha = 0.8) +
  scale_color_gradientn(colors = rev(colorRampPalette(brewer.pal(n = 8, 'Spectral'))(5)),
                        name = '',
                        limits = the_limits,
                        breaks = the_breaks) +
  scale_size_area(name = '', limits = the_limits, breaks = the_breaks, max_size = 10) +
  theme_simple() +
  facet_wrap(~country, scales = 'free') +
  theme(strip.text = element_text(size = 26),
        axis.title = element_blank(),
        axis.text = element_text(size = 16),
        axis.text.x = element_text(size = 12)) +
  guides(color = guide_legend(), size = guide_legend()) +
  labs(title = 'Daily COVID-19 deaths per 1,000,000 population by sub-state regions',
       caption = paste0('Data as of ', max_date))
ggsave('/tmp/2.png',
       width = 10,
       height = 8)
```

## Madrid vs rest of state

```{r}
place_transform <- function(x){ifelse(x == 'Madrid', 'Madrid',
                                      # ifelse(x == 'Cataluña', 'Cataluña',
                                             'Otras CCAA')
  # )
}
place_transform_ita <- function(x){
  ifelse(x == 'Lombardia', 'Lombardia', 
         # ifelse(x == 'Emilia Romagna', 'Emilia Romagna', 
                'Otras regiones italianas')
  # )
}
pd <- esp_df %>% mutate(country = 'España') %>%
  mutate(ccaa = place_transform(ccaa)) %>%
  bind_rows(ita %>% mutate(ccaa = place_transform_ita(ccaa),
                           country = 'Italia')) %>%
  group_by(country, ccaa, date) %>% 
  summarise(cases = sum(cases),
            uci = sum(uci),
            deaths = sum(deaths),
            cases_non_cum = sum(cases_non_cum),
            deaths_non_cum = sum(deaths_non_cum),
            uci_non_cum = sum(uci_non_cum)) %>%
  left_join(esp_pop %>%
              mutate(ccaa = place_transform(ccaa)) %>%
              bind_rows(ita_pop %>% mutate(ccaa = place_transform_ita(ccaa))) %>%
              group_by(ccaa) %>%
              summarise(pop = sum(pop))) %>%
  mutate(deaths_non_cum_p = deaths_non_cum / pop * 1000000) %>%
  group_by(country, date) %>%
  mutate(p_deaths_non_cum_country = deaths_non_cum / sum(deaths_non_cum) * 100,
         p_deaths_country = deaths / sum(deaths) * 100)
pd$ccaa <- factor(pd$ccaa,
                  levels = c('Madrid',
                             # 'Cataluña',
                             'Otras CCAA',
                             'Lombardia',
                             # 'Emilia Romagna',
                             'Otras regiones italianas'))
cols <- c(
  rev(brewer.pal(n = 3, 'Reds'))[1:2],
  rev(brewer.pal(n = 3, 'Blues'))[1:2]
)
ggplot(data = pd,
       aes(x = date,
           y = deaths_non_cum_p,
           fill = ccaa,
           group = ccaa)) +
  geom_bar(stat = 'identity',
           position = position_dodge(width = 0.99)) +
  scale_fill_manual(name = '', values = cols) +
  scale_color_manual(name = '', values = cols) +
  # geom_line(size = 0.2,
  #           aes(color = ccaa)) +
  xlim(as.Date('2020-03-09'),
       Sys.Date()-1) +
  theme_simple() +
  labs(x = 'Fecha',
       y = 'Muertes diarias por 1.000.000',
       title = 'Muertes por 1.000.000 habitantes') +
  theme(legend.position = 'top') +
  geom_text(aes(label = round(deaths_non_cum_p, digits = 1),
                color = ccaa,
                y = deaths_non_cum_p + 2,
                group = ccaa),
            size = 2.5,
            angle = 90,
            position = position_dodge(width = 0.99))


label_data <- pd %>%
  filter(country  %in% c('Italia', 'España')) %>%
  group_by(country) %>%
  filter(date == max(date))  %>%
  mutate(label = gsub('\nitalianas', '',  gsub(' ', '\n', ccaa))) %>%
  mutate(x = date - 2,
         y = p_deaths_country + 
           ifelse(p_deaths_country > 50, 10, -10))
ggplot(data = pd %>% group_by(country) %>% mutate(start_day = dplyr::first(date[deaths >=50])) %>% filter(date >= start_day),
       aes(x = date,
           y = p_deaths_country,
           color = ccaa,
           group = ccaa)) +
  # geom_bar(stat = 'identity',
  #          position = position_dodge(width = 0.99)) +
  scale_fill_manual(name = '', values = cols) +
  scale_color_manual(name = '', values = cols) +
  geom_line(size = 2,
            aes(color = ccaa)) +
  geom_point(size = 3,
             aes(color = ccaa)) +
  facet_wrap(~country, scales = 'free_x') +
  # xlim(as.Date('2020-03-09'),
  #      Sys.Date()-1) +
  theme_simple() +
  geom_hline(yintercept = 50, lty = 2, alpha = 0.6) +
  # geom_line(lty = 2, alpha = 0.6) +
  labs(x = 'Fecha',
       y = 'Porcentaje de muertes',
       title = 'Porcentaje de muertes del país: región más afectada vs. resto del país',
       subtitle = 'A partir del primer día en cada país con 50 o más muertes') +
  theme(legend.position = 'top',
        strip.text = element_text(size = 30),
        legend.text = element_text(size = 10))  +
  guides(color = guide_legend(nrow = 2,
                              keywidth = 5)) +
  geom_text(data = label_data,
            aes(x = x - 2,
                y = y,
                label = label,
                color = ccaa),
            size = 6,
            show.legend = FALSE)
```

## Italy regions, Spanish regions, Chinese regions (adjusted for population)


```{r}
# Spanish data
a <- esp_df %>%
  left_join(esp_pop) %>%
  mutate(country = 'Spain')
# Italian data
b <- ita %>%
  left_join(ita_pop) %>%
  mutate(country = 'Italy')
# Chinese data
d <- df %>% filter(country == 'China') %>%
  mutate(cases = cases) %>%
  mutate(ccaa = district) %>%
  mutate(country = 'China') %>%
  left_join(chi_pop)
# join
joined <- bind_rows(a, b, d)
# Get deaths per milllion
joined$deaths_pm <- joined$deaths / joined$pop * 1000000
joined$cases_pm <- joined$cases / joined$pop * 1000000

# Get the days since paradigm
x_deaths <- 5
x_deaths_pm <- 5
x_cases <- 50
x_cases_pm <- 50
joined <- joined %>%
  arrange(date) %>%
  group_by(ccaa) %>%
  mutate(start_deaths = min(date[deaths >= x_deaths]),
         start_cases = min(date[cases >= x_cases]),
         start_deaths_pm = min(date[deaths_pm >= x_deaths_pm]),
         start_cases_pm = min(date[cases_pm >= x_cases_pm])) %>%
  ungroup %>%
  mutate(days_since_start_deaths = date - start_deaths,
         days_since_start_cases = date - start_cases,
         days_since_start_deaths_pm = date - start_deaths_pm,
         days_since_start_cases_pm = date - start_cases_pm) 

# Define plot data
pd <- joined %>% filter(days_since_start_deaths_pm >= 0) %>%
  mutate(country = ifelse(country == 'China',
                          'Hubei (China)',
                          ifelse(country == 'Italy', 'Italia', 'España')))

lombardy_location <- (max(pd_big$days_since_start_deaths_pm[pd_big$ccaa == 'Lombardia']))

# Define label data
label_data <- pd %>% group_by(ccaa) %>% filter(
                                                          (
                                                            (country == 'Hubei (China)' & days_since_start_deaths_pm == 22) |
                                                            (date == max(date) & country == 'España' & deaths_pm > 40 & days_since_start_deaths_pm >= 7) & ccaa != 'CyL' |
                                                              (date == max(date) & country == 'Italia' &
                                                                 ccaa != 'Liguria' & days_since_start_deaths_pm > 15)
                                                          ))
# Get differential label data based on what to be emphasized
bigs <- c('Madrid', 'Lombardia', 'Hubei')
label_data_big <- label_data %>%
  filter(ccaa %in% bigs)
label_data_small <- label_data %>%
  filter(!ccaa %in% bigs)

pd_big <- pd %>%
  filter(ccaa %in% bigs)
pd_small <- pd %>%
  filter(!ccaa %in% bigs)

# cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = 'Set2'))(length(unique(pd$country)))
# cols <- rainbow(3)
cols <- c('darkred', '#FF6633', '#006666')

ggplot(data = pd_big,
       aes(x = days_since_start_deaths_pm,
           y = deaths_pm,
           group = ccaa)) +
  geom_line(aes(color = country),
            alpha = 0.9,
            size = 2) +
  geom_line(data = pd_small,
            aes(x = days_since_start_deaths_pm,
                y = deaths_pm,
                color = country),
            alpha = 0.7,
            size = 1) +
  scale_y_log10() +
  scale_color_manual(name = '',
                     values = c(cols)) +
  theme_simple() +
  theme(legend.position = 'top') +
  labs(x = 'Dias desde "el comienzo del brote"',
       y = 'Muertes por millón de habitantes',
       title = 'Muertes por 1.000.000 habitantes',
       subtitle = paste0('Dia 0 ("comienzo del brote") = primer día a ', x_deaths_pm, ' o más muertes acumuladas por milión de población\nLíneas rojas: CCAA; líneas verde-azules: regiones italianas; línea naranja: Hubei, China'),
       caption = '@joethebrew | www.databrew.cc') +
  geom_text(data = label_data_big,
            aes(x = days_since_start_deaths_pm + 0.6,
                y = (deaths_pm + 50),
                label = gsub(' ', '\n', ccaa),
                color = country),
            size = 8,
            alpha = 0.9,
            show.legend = FALSE) +
    geom_text(data = label_data_small,
            aes(x = days_since_start_deaths_pm + 0.6,
                y = deaths_pm  + (log(deaths_pm)/2),
                label = gsub(' ', '\n', ccaa),
                color = country),
            size = 5,
            alpha = 0.6,
            show.legend = FALSE) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        legend.text = element_text(size = 16),
        plot.title = element_text(size = 25))  +
  xlim(0, lombardy_location + 5)
```



## Italy regions, Spanish regions, Chinese regions (raw numbers, not adjusted for population)


```{r}
# Spanish data
a <- esp_df %>%
  left_join(esp_pop) %>%
  mutate(country = 'Spain')
# Italian data
b <- ita %>%
  left_join(ita_pop) %>%
  mutate(country = 'Italy')
# Chinese data
d <- df %>% filter(country == 'China') %>%
  mutate(cases = cases) %>%
  mutate(ccaa = district) %>%
  mutate(country = 'China') %>%
  left_join(chi_pop)
# join
joined <- bind_rows(a, b, d)
# Get deaths per milllion
joined$deaths_pm <- joined$deaths / joined$pop * 1000000
joined$cases_pm <- joined$cases / joined$pop * 1000000

# Get the days since paradigm
x_deaths <- 5
x_deaths_pm <- 5
x_cases <- 50
x_cases_pm <- 50
joined <- joined %>%
  arrange(date) %>%
  group_by(ccaa) %>%
  mutate(start_deaths = min(date[deaths >= x_deaths]),
         start_cases = min(date[cases >= x_cases]),
         start_deaths_pm = min(date[deaths_pm >= x_deaths_pm]),
         start_cases_pm = min(date[cases_pm >= x_cases_pm])) %>%
  ungroup %>%
  mutate(days_since_start_deaths = date - start_deaths,
         days_since_start_cases = date - start_cases,
         days_since_start_deaths_pm = date - start_deaths_pm,
         days_since_start_cases_pm = date - start_cases_pm) 

# Define plot data
pd <- joined %>% filter(days_since_start_deaths >= 0) %>%
  mutate(country = ifelse(country == 'China',
                          'China',
                          ifelse(country == 'Italy', 'Italia', 'España')))

# Define label data
label_data <- pd %>% group_by(ccaa) %>% filter(
                                                          (
                                                            (country == 'China' & deaths >10 & days_since_start_deaths == 29) |
                                                            (date == max(date) & country == 'España' & deaths > 90) |
                                                              (date == max(date) & country == 'Italia' &
                                                                 ccaa != 'Liguria' & days_since_start_deaths > 10)
                                                          ))
# Get differential label data based on what to be emphasized
label_data_big <- label_data %>%
  filter(ccaa %in% c('Madrid', 'Lombardia', 'Hubei'))
label_data_small <- label_data %>%
  filter(!ccaa %in% c('Madrid', 'Lombardia', 'Hubei'))

pd_big <- pd %>%
  filter(ccaa %in% c('Madrid', 'Lombardia', 'Hubei'))
pd_small <- pd %>%
  filter(!ccaa %in% c('Madrid', 'Lombardia', 'Hubei'))

# cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = 'Set2'))(length(unique(pd$country)))
# cols <- rainbow(3)
cols <- c( '#FF6633',  'darkred', '#006666')
ggplot(data = pd_big,
       aes(x = days_since_start_deaths,
           y = deaths,
           group = ccaa)) +
  geom_line(aes(color = country),
            alpha = 0.9,
            size = 2) +
  geom_line(data = pd_small,
            aes(x = days_since_start_deaths,
                y = deaths,
                color = country),
            alpha = 0.7,
            size = 1) +
  scale_y_log10() +
  scale_color_manual(name = '',
                     values = c(cols)) +
  theme_simple() +
  theme(legend.position = 'top') +
  labs(x = 'Dias desde el primer día con 5 o más muertes acumuladas',
       y = 'Muertes',
       title = 'Muertes por COVID-19',
       caption = '@joethebrew | www.databrew.cc') +
  geom_text(data = label_data_big,
            aes(x = days_since_start_deaths + 1.6,
                y = ifelse(ccaa == 'Hubei', (deaths -500),
                           ifelse(ccaa == 'Lombardia',  (deaths + 700),
                                   (deaths + 300))),
                label = gsub(' ', '\n', ccaa),
                color = country),
            size = 8,
            alpha = 0.9,
            show.legend = FALSE) +
    geom_text(data = label_data_small,
            aes(x = days_since_start_deaths + 1.6,
                align = 'left',
                y = deaths ,
                label = ccaa,
                # label = gsub(' ', '\n', ccaa),
                color = country),
            size = 5,
            alpha = 0.6,
            show.legend = FALSE) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 16),
        plot.title = element_text(size = 30))  +
  xlim(0, 50)
```



## ANIMATION: Italy regions, Spanish regions, Chinese regions (raw numbers, not adjusted for population)


```{r, eval = F}
# Spanish data
a <- esp_df %>%
  left_join(esp_pop) %>%
  mutate(country = 'Spain')
# Italian data
b <- ita %>%
  left_join(ita_pop) %>%
  mutate(country = 'Italy')
# Chinese data
d <- df %>% filter(country == 'China') %>%
  mutate(cases = cases) %>%
  mutate(ccaa = district) %>%
  mutate(country = 'China') %>%
  left_join(chi_pop)
# join
joined <- bind_rows(a, b, d)
# Get deaths per milllion
joined$deaths_pm <- joined$deaths / joined$pop * 1000000
joined$cases_pm <- joined$cases / joined$pop * 1000000

# Get the days since paradigm
x_deaths <- 5
x_deaths_pm <- 5
x_cases <- 50
x_cases_pm <- 50
joined <- joined %>%
  arrange(date) %>%
  group_by(ccaa) %>%
  mutate(start_deaths = min(date[deaths >= x_deaths]),
         start_cases = min(date[cases >= x_cases]),
         start_deaths_pm = min(date[deaths_pm >= x_deaths_pm]),
         start_cases_pm = min(date[cases_pm >= x_cases_pm])) %>%
  ungroup %>%
  mutate(days_since_start_deaths = date - start_deaths,
         days_since_start_cases = date - start_cases,
         days_since_start_deaths_pm = date - start_deaths_pm,
         days_since_start_cases_pm = date - start_cases_pm) 

# Define plot data
pd <- joined %>% filter(days_since_start_deaths >= 0) %>%
  mutate(country = ifelse(country == 'China',
                          'China',
                          ifelse(country == 'Italy', 'Italia', 'España')))


add_zero <- function(x, n){
  x <- as.character(x)
  adders <- n - nchar(x)
  adders <- ifelse(adders < 0, 0, adders)
  for (i in 1:length(x)){
    if(!is.na(x[i])){
      x[i] <- paste0(
        paste0(rep('0', adders[i]), collapse = ''),
        x[i],
        collapse = '')  
    } 
  }
  return(x)
}
# # Define label data
# label_data <- pd %>% group_by(ccaa) %>% filter(
#                                                           (
#                                                             (country == 'China' & deaths >10 & days_since_start_deaths == 29) |
#                                                             (date == max(date) & country == 'España' & deaths > 90) |
#                                                               (date == max(date) & country == 'Italia' &
#                                                                  ccaa != 'Liguria' & days_since_start_deaths > 10)
#                                                           ))
# # Get differential label data based on what to be emphasized
# label_data_big <- label_data %>%
#   filter(ccaa %in% c('Madrid', 'Lombardia', 'Hubei'))
# label_data_small <- label_data %>%
#   filter(!ccaa %in% c('Madrid', 'Lombardia', 'Hubei'))
# 
pd_big <- pd %>%
  filter(ccaa %in% c('Madrid', 'Lombardia', 'Hubei'))
pd_small <- pd %>%
  filter(!ccaa %in% c('Madrid', 'Lombardia', 'Hubei'))



# cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = 'Set2'))(length(unique(pd$country)))
# cols <- rainbow(3)
cols <- c( '#FF6633',  'darkred', '#006666')

the_dir <- '/tmp/animation/'
dir.create(the_dir)
the_dates <- sort(unique(c(pd_big$date, pd_small$date)))
for(i in 1:length(the_dates)){
  
  the_date <- the_dates[i]
  pd_big_sub <- pd_big %>% filter(date <= the_date)
  pd_big_current <- pd_big_sub %>% filter(date == the_date)
  pd_small_sub <- pd_small %>% filter(date <= the_date)
  pd_small_current <- pd_small_sub %>% filter(date == the_date)

  label_data_big <-
    pd_big_sub %>%
    filter(ccaa %in% c('Lombardia', 'Madrid', 'Hubei')) %>%
    group_by(ccaa) %>%
    filter(date == max(date)) %>%
    ungroup %>%
    mutate(days_since_start_deaths = ifelse(ccaa == 'Hubei' &
                                              days_since_start_deaths >32,
                                            32,
                                            days_since_start_deaths))
  
  label_data_small <-
    pd_small_sub %>%
    filter(ccaa %in% c('Emilia Romagna',
                       'Cataluña',
                       'CLM',
                       'País Vasco',
                       'Veneto',
                       'Piemonte',
                       'Henan',
                       'Heilongjiang')) %>%
    group_by(ccaa) %>%
    filter(date == max(date))

  n_countries <- length(unique(pd_big_sub$country))
  if(n_countries == 3){
    sub_cols  <- cols
  }
  if(n_countries == 2){
    sub_cols <- cols[c(1,3)]
  }
   if(n_countries == 1){
    sub_cols <- cols[1]
  }
  g <- ggplot(data = pd_big_sub,
       aes(x = days_since_start_deaths,
           y = deaths,
           group = ccaa)) +
  geom_line(aes(color = country),
            alpha = 0.9,
            size = 2) +
  geom_line(data = pd_small_sub,
            aes(x = days_since_start_deaths,
                y = deaths,
                color = country),
            alpha = 0.7,
            size = 1) +
    geom_point(data = pd_big_current,
               aes(x = days_since_start_deaths,
                y = deaths,
                color = country),
               size = 3) +
    geom_point(data = pd_small_current,
               aes(x = days_since_start_deaths,
                y = deaths,
                color = country),
               size = 1, alpha = 0.6) +
    scale_y_log10(limits = c(5, 4500)) +
  scale_color_manual(name = '',
                     values = sub_cols) +
  theme_simple() +
  theme(legend.position = 'top') +
  labs(x = 'Dias desde el primer día con 5 o más muertes acumuladas',
       y = 'Muertes',
       title = format(the_date, '%d %b'),
       subtitle = 'Muertes por COVID-19',
       caption = '@joethebrew | www.databrew.cc') +
  geom_text(data = label_data_big,
            aes(x = days_since_start_deaths + 1,
                y = deaths,
                hjust = 0,
                label = gsub(' ', '\n', ccaa),
                color = country),
            size = 8,
            alpha = 0.9,
            show.legend = FALSE) +
    geom_text(data = label_data_small,
            aes(x = days_since_start_deaths + 1.6,
                y = deaths ,
                label = ccaa,
                # label = gsub(' ', '\n', ccaa),
                color = country),
            size = 5,
            alpha = 0.6,
            show.legend = FALSE) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 16),
        plot.title = element_text(size = 35),
        plot.subtitle = element_text(size = 24))  +
  xlim(0, 38) 
  message(i)
  ggsave(paste0(the_dir, add_zero(i, 3), '.png'),
         height = 7,
         width = 10.5)
}
```




```
# Command line
cd /tmp/animation
mogrify -resize 50% *.png
convert -delay 20 -loop 0 *.png result.gif

```


## ANIMATION: Spain only


```{r, eval = F}
# Spanish data
a <- esp_df %>%
  left_join(esp_pop) %>%
  mutate(country = 'Spain')
joined <- a
# Get deaths per milllion
joined$deaths_pm <- joined$deaths / joined$pop * 1000000
joined$cases_pm <- joined$cases / joined$pop * 1000000

# Get the days since paradigm
x_deaths <- 5
x_deaths_pm <- 5
x_cases <- 50
x_cases_pm <- 50
joined <- joined %>%
  arrange(date) %>%
  group_by(ccaa) %>%
  mutate(start_deaths = min(date[deaths >= x_deaths]),
         start_cases = min(date[cases >= x_cases]),
         start_deaths_pm = min(date[deaths_pm >= x_deaths_pm]),
         start_cases_pm = min(date[cases_pm >= x_cases_pm])) %>%
  ungroup %>%
  mutate(days_since_start_deaths = date - start_deaths,
         days_since_start_cases = date - start_cases,
         days_since_start_deaths_pm = date - start_deaths_pm,
         days_since_start_cases_pm = date - start_cases_pm) 

# Define plot data
pd <- joined %>% filter(days_since_start_deaths >= 0) %>%
  mutate(country = ifelse(country == 'China',
                          'China',
                          ifelse(country == 'Italy', 'Italia', 'España')))

bigs <- c('Madrid', 'Cataluña', 'CLM', 'CyL', 'País Vasco', 'La Rioja')
pd_big <- pd %>%
  filter(ccaa %in% bigs)
pd_small <- pd %>%
  filter(!ccaa %in% bigs)



# cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = 'Set2'))(length(unique(pd$country)))
# cols <- rainbow(3)
cols <- colorRampPalette(c('#A16928','#bd925a','#d6bd8d','#edeac2', '#b5c8b8','#79a7ac','#2887a1'))(length(unique(pd$country)))

the_dir <- '/tmp/animation2/'
dir.create(the_dir)
the_dates <- sort(unique(c(pd_big$date, pd_small$date)))
for(i in 1:length(the_dates)){
  
  the_date <- the_dates[i]
  pd_big_sub <- pd_big %>% filter(date <= the_date)
  pd_big_current <- pd_big_sub %>% filter(date == the_date)
  pd_small_sub <- pd_small %>% filter(date <= the_date)
  pd_small_current <- pd_small_sub %>% filter(date == the_date)

  label_data_big <-
    pd_big_sub %>%
    filter(ccaa %in% bigs) %>%
    group_by(ccaa) %>%
    filter(date == max(date)) %>%
    ungroup
  
  label_data_small <-
    pd_small_sub %>%
    group_by(ccaa) %>%
    filter(date == max(date))
# sub_cols <- colorRampPalette(c('#A16928','#bd925a','#d6bd8d','#edeac2', '#b5c8b8','#79a7ac','#2887a1'))(length(unique(pd$ccaa)))
  sub_cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = 'Dark2'))(length(unique(pd$ccaa)))
  # sub_cols <- rainbow((length(unique(pd$ccaa))))
  
  g <- ggplot(data = pd_big_sub,
       aes(x = days_since_start_deaths,
           y = deaths,
           group = ccaa)) +
  geom_line(aes(color = ccaa),
            alpha = 0.9,
            size = 2) +
  geom_line(data = pd_small_sub,
            aes(x = days_since_start_deaths,
                y = deaths,
                color = ccaa),
            alpha = 0.7,
            size = 1) +
    geom_point(data = pd_big_current,
               aes(x = days_since_start_deaths,
                y = deaths,
                color = ccaa),
               size = 3) +
    geom_point(data = pd_small_current,
               aes(x = days_since_start_deaths,
                y = deaths,
                color = ccaa),
               size = 1, alpha = 0.6) +
    geom_point(data = pd,
               aes(x = days_since_start_deaths,
                y = deaths,
                color = ccaa),
               size = 1, alpha = 0.01) +
    scale_y_log10(limits = c(5, max(pd$deaths)*1.2),
                  breaks = c(10, 50, 100, 500, 1000)) +
  scale_color_manual(name = '',
                     values = sub_cols) +
  theme_simple() +
  theme(legend.position = 'top') +
  labs(x = 'Dias desde el primer día con 5 o más muertes acumuladas',
       y = 'Muertes',
       title = format(the_date, '%d %b'),
       subtitle = 'Muertes por COVID-19',
       caption = '@joethebrew | www.databrew.cc')   +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 16),
        plot.title = element_text(size = 35),
        plot.subtitle = element_text(size = 24))  +
  xlim(0, 20) +
    theme(legend.position = 'none')
  message(i)
  if(nrow(label_data_big) > 0){
    g <- g +
      geom_text(data = label_data_big,
            aes(x = days_since_start_deaths + 0.2,
                y = deaths,
                hjust = 0,
                label = gsub(' ', ' ', ccaa),
                color = ccaa),
            size = 8,
            alpha = 0.9,
            show.legend = FALSE) +
    geom_text(data = label_data_small,
            aes(x = days_since_start_deaths + 0.2,
                y = deaths ,
                label = ccaa,
                # label = gsub(' ', '\n', ccaa),
                color = ccaa),
            size = 5,
            alpha = 0.6,
            show.legend = FALSE)
  }
  
  ggsave(paste0(the_dir, add_zero(i, 3), '.png'),
         height = 7,
         width = 12)
}
```

```
# Command line
cd /tmp/animation
mogrify -resize 50% *.png
convert -delay 25 -loop 0 *.png result.gif

```




## Italy regions for Spanish regions


```{r}
# Spanish data
a <- esp_df %>%
  left_join(esp_pop) %>%
  mutate(country = 'Spain')
# Italian data
b <- ita %>%
  left_join(ita_pop) %>%
  mutate(country = 'Italy')
# join
joined <- bind_rows(a, b)
# Get deaths per milllion
joined$deaths_pm <- joined$deaths / joined$pop * 1000000
joined$cases_pm <- joined$cases / joined$pop * 1000000

# Get the days since paradigm
x_deaths <- 5
x_deaths_pm <- 5
x_cases <- 50
x_cases_pm <- 50
joined <- joined %>%
  arrange(date) %>%
  group_by(ccaa) %>%
  mutate(start_deaths = min(date[deaths >= x_deaths]),
         start_cases = min(date[cases >= x_cases]),
         start_deaths_pm = min(date[deaths_pm >= x_deaths_pm]),
         start_cases_pm = min(date[cases_pm >= x_cases_pm])) %>%
  ungroup %>%
  mutate(days_since_start_deaths = date - start_deaths,
         days_since_start_cases = date - start_cases,
         days_since_start_deaths_pm = date - start_deaths_pm,
         days_since_start_cases_pm = date - start_cases_pm) 

ggplot(data = joined %>% filter(days_since_start_deaths_pm >= 0),
       aes(x = days_since_start_deaths_pm,
           y = deaths_pm,
           group = ccaa)) +
  geom_line(aes(color = country),
            alpha = 0.8,
            size = 2) +
  scale_y_log10() +
  scale_color_manual(name = '',
                     values = c('darkorange', 'purple')) +
  theme_simple() +
  theme(legend.position = 'none') +
  labs(x = 'Days since "start out outbreak"',
       y = 'Deaths per million',
       title = 'Deaths per capita, Italian regions vs. Spanish autonomous communities',
       subtitle = paste0('Day 0 ("start of outbreak") = first day at ', x_deaths_pm, ' or greater cumulative deaths per million'),
       caption = '@joethebrew | www.databrew.cc') +
  geom_text(data = joined %>% group_by(ccaa) %>% filter(date == max(date) & 
                                                          (
                                                            (country == 'Spain' & deaths_pm > 25) |
                                                              (country == 'Italy' & days_since_start_deaths_pm > 10)
                                                          )),
            aes(x = days_since_start_deaths_pm + 0.6,
                y = deaths_pm,
                label = gsub(' ', '\n', ccaa),
                color = country),
            size = 6) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 20)) +
  xlim(0, 23)
ggsave('/tmp/italy_comparison.png',
       height = 6,
       width = 10)


# Separate for Catalonia
pd <- joined %>% filter(days_since_start_deaths_pm >= 0) %>%
         mutate(country = ifelse(ccaa == 'Cataluña',
                                 'Catalonia',
                                 country)) %>%
  mutate(ccaa = ifelse(ccaa == 'Cataluña', 'Catalunya', ccaa))
pdcat <- pd %>% filter(country == 'Catalonia')
label_data <- pd %>% group_by(ccaa) %>% filter(date == max(date) & 
                                                          (
                                                            (country == 'Catalonia') |
                                                            (country == 'Spain' & deaths_pm > 25) |
                                                              (country == 'Italy' & days_since_start_deaths_pm > 10)
                                                          ))
ggplot(data = pd,
       aes(x = days_since_start_deaths_pm,
           y = deaths_pm,
           group = ccaa)) +
  geom_line(aes(color = country),
            alpha = 0.3,
            size = 1.5) +
    geom_line(data = pdcat,
              aes(color = country),
            alpha = 0.8,
            size = 2) +
      geom_point(data = pdcat %>% filter(date == max(date)),
              aes(color = country),
            alpha = 0.8,
            size = 4) +
  scale_y_log10() +
  scale_color_manual(name = '',
                     values = c('darkred', 'darkorange', "purple")) +
  theme_simple() +
  theme(legend.position = 'none') +
  labs(x = 'Dies des del "començament del brot"',
       y = 'Morts per milió',
       title = 'Morts per càpita: Catalunya, comunitats autònomes, regions italianes',
       subtitle = paste0('Dia 0 ("començament del brot") = primer dia a ', x_deaths_pm, ' o més morts acumulades per milió de població'),
       caption = '@joethebrew | www.databrew.cc') +
  geom_text(data = label_data,
            aes(x = days_since_start_deaths_pm +0.2 ,
                y = deaths_pm +3,
                hjust = 0,
                label = gsub(' ', '\n', ccaa),
                color = country),
            size = 6,
            alpha = 0.7) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 20)) +
  xlim(0, 24)
ggsave('/tmp/cat_italy_comparison.png',
       height = 6,
       width = 10)


# Straightforward Lombardy, Madrid, Cat comparison
specials <- c('Lombardia', 'Madrid')
pd <- joined %>% filter(days_since_start_deaths_pm >= 0) %>%
         mutate(country = ifelse(ccaa == 'Cataluña',
                                 'Catalonia',
                                 country)) %>%
  mutate(ccaa = ifelse(ccaa == 'Cataluña', 'Catalunya', ccaa))
pdcat <- pd %>% filter(ccaa %in%  specials)
label_data <- pd %>% group_by(ccaa) %>% filter(date == max(date) & 
                                                          (
                                                            # (country == 'Catalonia') |
                                                            (country == 'Spain' & deaths_pm > 20) |
                                                              (country == 'Italy' & days_since_start_deaths_pm >= 10)
                                                          ))
ggplot(data = pd,
       aes(x = days_since_start_deaths_pm,
           y = deaths_pm,
           group = ccaa)) +
  geom_line(aes(color = country),
            alpha = 0.3,
            size = 1.5) +
    geom_line(data = pdcat,
              aes(color = country),
            alpha = 0.8,
            size = 2) +
  scale_y_log10() +
  scale_color_manual(name = '',
                     values = c('darkred', 'darkorange', "purple")) +
  theme_simple() +
  theme(legend.position = 'none') +
  labs(x = 'Dias desde "el comienzo del brote"',
       y = 'Muertes por millón de habitantes',
       title = 'Muertes acumuladas por 1.000.000 habitantes',
       subtitle = paste0('Dia 0 ("comienzo del brote") = primer día a ', x_deaths_pm, ' o más muertes acumuladas por milión de población'),
       caption = '@joethebrew | www.databrew.cc') +
  geom_text(data = label_data %>% filter(!ccaa %in% specials),
            aes(x = days_since_start_deaths_pm + 0.4,
                y = deaths_pm +3,
                label = gsub(' ', '\n', ccaa),
                color = country),
            size = 5,
            alpha = 0.5) +
    geom_text(data = label_data %>% filter(ccaa %in% specials),
            aes(x = days_since_start_deaths_pm ,
                y = deaths_pm +30,
                label = gsub(' ', '\n', ccaa),
                color = country),
            size = 8,
            alpha = 0.8) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 20)) +
  xlim(0, 23)
ggsave('/tmp/mad_lom_italy_comparison.png',
       height = 6,
       width = 10)
```

# Loop for regions of the world
```{r}
isos <- sort(unique(world_pop$sub_region))
isos <- c('Central Asia', 'Eastern Asia', 'Eastern Europe',
          'Latin America and the Caribbean',
          'Northern Africa', 'Northern America',
          'Nothern Europe',
          'South-eastern Asia',
          'Southern Asia', 'Southern Europe',
          'Sub-Saharan Africa', 'Western Asia', 'Western Europe')
dir.create('/tmp/world')
for(i in 1:length(isos)){
  this_iso <- isos[i]
  message(i, ' ', this_iso)
  countries <- world_pop %>% filter(sub_region == this_iso)
  pd <- df %>%
    filter(!country %in% c('Guyan
                           a', 'Bahamas', 'The Bahamas')) %>%
          group_by(country, iso, date) %>%
          summarise(cases = sum(cases, na.rm = TRUE)) %>%
    ungroup %>%
    group_by(country) %>%
         filter(length(which(cases > 0)) > 1) %>%
    ungroup %>%
         filter(iso %in% countries$iso)
  if(nrow(pd) > 0){
    cols <- colorRampPalette(brewer.pal(n = 8, 'Spectral'))(length(unique(pd$country)))
cols <- sample(cols, length(cols))
    # Plot
n_countries <- (length(unique(pd$country)))
ggplot(data = pd,
       aes(x = date,
           # color = country,
           # fill = country,
           y = cases)) +
  theme_simple() +
  # geom_point() +
  # geom_line() +
  geom_area(fill = 'darkred', alpha = 0.3, color = 'darkred') +
  # scale_color_manual(name = '',
  #                    values = cols) +
  # scale_fill_manual(name = '',
  #                   values = cols) +
  theme(legend.position = 'none',
        axis.text = element_text(size = 6),
        strip.text = element_text(size = ifelse(n_countries > 20, 6,
                                                ifelse(n_countries > 10, 10,
                                                       ifelse(n_countries > 5, 11, 12))) ),
        legend.text = element_text(size = 6)) +
  # scale_y_log10() +
  facet_wrap(~country,
             scales = 'free') +
  labs(x = '',
       y = 'Confirmed cases',
       title = paste0('Confirmed cases of COVID-19 in ', this_iso)) 
  ggsave(paste0('/tmp/world/', this_iso, '.png'),
         width = 12, 
         height = 7)
  }



}

```




## Rolling average new events


```{r}
plot_data <- df_country %>% filter(country %in% c('Spain', 'France', 'Italy', 'Germany', 'Belgium', 'Norway')) %>% mutate(geo = country)
roll_curve(plot_data, pop = T)
```

```{r}
dir.create('/tmp/countries')
roll_curve_country <- function(the_country = 'Spain'){
  plot_data <- df_country %>% filter(country %in% the_country) %>% mutate(geo = country)
  g1 <- roll_curve(plot_data, pop = F)
  g2 <- roll_curve(plot_data, pop = T)
  g3 <- roll_curve(plot_data, pop = F, deaths = T)
  g4 <- roll_curve(plot_data, pop = T, deaths = T)
  ggsave(paste0('/tmp/countries/', the_country, '1.png'), g1)
  ggsave(paste0('/tmp/countries/', the_country, '2.png'), g2)
  ggsave(paste0('/tmp/countries/', the_country, '3.png'), g3)
  ggsave(paste0('/tmp/countries/', the_country, '4.png'), g4)
}


countries <- c('Spain', 'France', 'Italy', 'Germany', 'Belgium', 'Norway', 'US', 'United Kingdom', 'Korea, South',
  'China', 'Japan', 'Switzerland', 'Sweden', 'Denmark', 'Netherlands', 'Iran', 'Canada')
for(i in 1:length(countries)){
  roll_curve_country(the_country = countries[i])
}
```




```{r}
# Cases
plot_data <- df_country %>% filter(country == 'Spain') %>% mutate(geo = country)
roll_curve(plot_data)

# Cases adjusted
plot_data <- df_country %>% filter(country == 'Spain') %>% mutate(geo = country)
roll_curve(plot_data, pop = T)

# Deaths
plot_data <- df_country %>% filter(country == 'Spain') %>% mutate(geo = country)
roll_curve(plot_data, deaths = T)

# Cases adjusted
plot_data <- df_country %>% filter(country == 'Spain') %>% mutate(geo = country)
roll_curve(plot_data, pop = T, deaths = T)
```


```{r}
plot_data <- esp_df  %>% mutate(geo = ccaa)

roll_curve(plot_data, pop = T, deaths = T)
```


```{r}
plot_data <- df_country %>% filter(country == 'Spain') %>% mutate(geo = country)

roll_curve(plot_data, deaths = T)
```

```{r}

# Latest in Spain
pd <- esp_df %>%
  filter(date == max(date)) %>%
  mutate(p = deaths / sum(deaths) * 100)
text_size <- 12

# deaths
ggplot(data = pd,
       aes(x = ccaa,
           y = deaths)) +
  geom_bar(stat = 'identity',
           fill = 'black') +
  theme_simple() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = '',
       y = 'Deaths | Muertes',
       title = 'COVID-19 deaths in Spain',
       subtitle = paste0('Data as of ', max(pd$date)),
       caption = 'github.com/databrew/covid19 | joe@databrew.cc') +
  theme(legend.position = 'top',
        legend.text = element_text(size = text_size * 2),
        axis.title = element_text(size = text_size * 2),
        plot.title = element_text(size = text_size * 2.3),
        axis.text.x = element_text(size = text_size * 1.5)) +
  geom_text(data = pd %>% filter(deaths > 0),
            aes(x = ccaa,
                y = deaths,
                label = paste0(deaths, '\n(',
                               round(p, digits = 1), '%)')),
            size = text_size * 0.3,
            nudge_y = 180) +
  ylim(0, max(pd$deaths * 1.1))
ggsave('/tmp/spain.png')

```


Muertes relativas por CCAA


```{r}

# Latest in Spain
pd <- esp_df %>%
  filter(date == max(date)) %>%
  mutate(p = deaths / sum(deaths) * 100)

pd <- pd %>% left_join(esp_pop)
text_size <- 12
pd$value <- pd$deaths / pd$pop * 100000

# deaths
ggplot(data = pd,
       aes(x = ccaa,
           y = value)) +
  geom_bar(stat = 'identity',
           fill = 'black') +
  theme_simple() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = '',
       y = 'Deaths per 100,000',
       title = 'COVID-19 deaths per 100.000',
       subtitle = paste0('Data as of ', max(pd$date)),
       caption = 'github.com/databrew/covid19 | joe@databrew.cc') +
  theme(legend.position = 'top',
        legend.text = element_text(size = text_size * 2),
        axis.title = element_text(size = text_size * 2),
        plot.title = element_text(size = text_size * 2.3),
        axis.text.x = element_text(size = text_size * 1.5)) +
  geom_text(data = pd %>% filter(value > 0),
            aes(x = ccaa,
                y = value,
                label = paste0(round(value, digits = 2), '\n(',
                               deaths, '\ndeaths)')),
            size = text_size * 0.3,
            nudge_y = 4.5) +
  ylim(0, max(pd$value) * 1.2)
ggsave('/tmp/spai2.png')

```


# Just yesterday

```{r}

# Latest in Spain
pd <- esp_df %>%
  filter(date == max(date)) %>%
  mutate(p = deaths_non_cum / sum(deaths_non_cum) * 100)
text_size <- 12

# deaths
ggplot(data = pd,
       aes(x = ccaa,
           y = deaths_non_cum)) +
  geom_bar(stat = 'identity',
           fill = 'black') +
  theme_simple() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = '',
       y = 'Deaths',
       title = 'COVID-19 deaths in Spain',
       subtitle = paste0('Data only for ', max(pd$date)),
       caption = 'github.com/databrew/covid19 | joe@databrew.cc') +
  theme(legend.position = 'top',
        legend.text = element_text(size = text_size * 2),
        axis.title = element_text(size = text_size * 2),
        plot.title = element_text(size = text_size * 2.3),
        axis.text.x = element_text(size = text_size * 1.5)) +
  geom_text(data = pd %>% filter(deaths_non_cum > 0),
            aes(x = ccaa,
                y = deaths_non_cum,
                label = paste0(deaths_non_cum, '\n(',
                               round(p, digits = 1), '%)')),
            size = text_size * 0.3,
            nudge_y = 30) +
  ylim(0, max(pd$deaths_non_cum * 1.1))
ggsave('/tmp/spain_non_cum.png')

```


Muertes relativas por CCAA ayer SOLO


```{r}

# Latest in Spain
pd <- esp_df %>%
  filter(date == max(date)) %>%
  mutate(p = deaths_non_cum / sum(deaths_non_cum) * 100)

pd <- pd %>% left_join(esp_pop)
text_size <- 12
pd$value <- pd$deaths_non_cum / pd$pop * 100000

# deaths
ggplot(data = pd,
       aes(x = ccaa,
           y = value)) +
  geom_bar(stat = 'identity',
           fill = 'black') +
  theme_simple() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = '',
       y = 'Deaths per 100,000',
       title = 'COVID-19 deaths per 100.000',
       subtitle = paste0('Data as of ', max(pd$date)),
       caption = 'github.com/databrew/covid19 | joe@databrew.cc') +
  theme(legend.position = 'top',
        legend.text = element_text(size = text_size * 2),
        axis.title = element_text(size = text_size * 2),
        plot.title = element_text(size = text_size * 2.3),
        axis.text.x = element_text(size = text_size * 1.5)) +
  geom_text(data = pd %>% filter(value > 0),
            aes(x = ccaa,
                y = value,
                label = paste0(round(value, digits = 2), '\n(',
                               deaths_non_cum, '\ndeaths)')),
            size = text_size * 0.3,
            nudge_y = 1) +
  ylim(0, max(pd$value) * 1.3)
ggsave('/tmp/spain_ayer_adj.png')

```



```{r}
plot_data <- esp_df %>% mutate(geo = ccaa) %>% filter(!ccaa %in% c('Melilla'))
roll_curve(plot_data, scales = 'fixed')
ggsave('/tmp/a.png',
       width = 1280 / 150,
       height = 720 / 150)
```

Loop for everywhere (see desktop)

```{r}
dir.create('/tmp/ccaas')
ccaas <- sort(unique(esp_df$ccaa))
for(i in 1:length(ccaas)){
  this_ccaa <- ccaas[i]
  plot_data <- esp_df %>% mutate(geo = ccaa) %>% filter(ccaa == this_ccaa)
  roll_curve(plot_data, scales = 'fixed')  + theme(strip.text = element_text(size = 30))
  ggsave(paste0('/tmp/ccaas/', i, this_ccaa, '_cases.png'),
         width = 1280 / 150,
         height = 720 / 150)
}

ccaas <- sort(unique(esp_df$ccaa))
for(i in 1:length(ccaas)){
  this_ccaa <- ccaas[i]
  plot_data <- esp_df %>% mutate(geo = ccaa) %>% filter(ccaa == this_ccaa)
  roll_curve(plot_data, scales = 'fixed', pop = TRUE)  + theme(strip.text = element_text(size = 30))
  ggsave(paste0('/tmp/ccaas/', i, this_ccaa, '_cases_pop.png'),
         width = 1280 / 150,
         height = 720 / 150)
}

# Deaths too
for(i in 1:length(ccaas)){
  this_ccaa <- ccaas[i]
  plot_data <- esp_df %>% mutate(geo = ccaa) %>% filter(ccaa == this_ccaa)
  roll_curve(plot_data, deaths = T, scales = 'fixed') + theme(strip.text = element_text(size = 30))
  ggsave(paste0('/tmp/ccaas/', i, this_ccaa, '_deaths.png'),
         width = 1280 / 150,
         height = 720 / 150)
}

# Deaths too
for(i in 1:length(ccaas)){
  this_ccaa <- ccaas[i]
  plot_data <- esp_df %>% mutate(geo = ccaa) %>% filter(ccaa == this_ccaa)
  roll_curve(plot_data, deaths = T, scales = 'fixed', pop = TRUE)  + theme(strip.text = element_text(size = 30))
  ggsave(paste0('/tmp/ccaas/', i, this_ccaa, '_deaths_pop.png'),
         width = 1280 / 150,
         height = 720 / 150)
}

```


```{r}
plot_data <- esp_df %>% mutate(geo = ccaa) %>% filter(!ccaa %in% c('Melilla'))
roll_curve(plot_data, scales = 'free_y')
ggsave('/tmp/b.png',
       width = 1280 / 150,
       height = 720 / 150)
```

```{r}
plot_data <- esp_df %>% mutate(geo = ccaa) %>% filter(!ccaa %in% c('Melilla'))
roll_curve(plot_data, deaths = T, scales = 'free_y')
ggsave('/tmp/c.png',
       width = 1280 / 150,
       height = 720 / 150)
```


```{r}
plot_data <- esp_df %>% mutate(geo = ccaa) %>% filter(!ccaa %in% c('Melilla'))
roll_curve(plot_data, deaths = T, scales = 'fixed')
ggsave('/tmp/d.png',
       width = 1280 / 150,
       height = 720 / 150)
```


```{r}
plot_data <- df_country %>% filter(country %in% c('Spain', 'Italy', 'Germany', 'France', 'US',
                                                  'China', 'Korea, South', 'Japan', 'Singapore')) %>% mutate(geo = country)
roll_curve(plot_data, scales = 'free_y')
ggsave('/tmp/z.png',
       width = 1280 / 150,
       height = 720 / 150)

```



## World at large

```{r}
pd <- df_country %>%
  group_by(date) %>%
  summarise(n = sum(cases)) %>%
  filter(date < max(date))
ggplot(data = pd,
       aes(x = date,
           y = n)) +
  geom_point() +
  theme_simple() +
  labs(x = 'Date',
       y = 'Cases',
       title = 'COVID-19 cases')
ggsave('~/Videos/update/a.png',
       width = 1280 / 150,
       height = 720 / 150)
```


## China vs world

```{r}
pd <- df_country %>%
  group_by(date,
           country = ifelse(country == 'China', 'China', 'Other countries')) %>%
  summarise(n = sum(cases))  %>%
  ungroup %>%
  filter(date < max(date))
ggplot(data = pd,
       aes(x = date,
           y = n,
           color = country)) +
  geom_line(size = 2) +
  # geom_point() +
  theme_simple() +
  labs(x = 'Date',
       y = 'Cases',
       title = 'COVID-19 cases') +
  scale_color_manual(name = '',
                     values = c('red', 'black')) +
  theme(legend.text = element_text(size = 25),
        legend.position = 'top')
ggsave('~/Videos/update/b.png',
       width = 1280 / 150,
       height = 720 / 150)
```

## NOn china only


```{r}
pd <- df_country %>%
  group_by(date,
           country = ifelse(country == 'China', 'China', 'Other countries')) %>%
  summarise(n = sum(cases)) %>%
  filter(country == 'Other countries')  %>%
  ungroup %>%
  filter(date < max(date))
ggplot(data = pd,
       aes(x = date,
           y = n)) +
  geom_line(size = 2) +
  # geom_point() +
  theme_simple() +
  labs(x = 'Date',
       y = 'Cases',
       title = 'COVID-19 cases, outside of China') 
ggsave('~/Videos/update/c.png',
       width = 1280 / 150,
       height = 720 / 150)
```

## Case numbers across countries

```{r}
plot_day_zero(countries = c('France', 'Germany', 'Italy', 'Spain', 'Switzerland', 'Sweden', 'Norway', 'Netherlands'))
# ggsave('~/Videos/update/d.png',
#        width = 1280 / 150,
#        height = 720 / 150)
```


## World at large - deaths

```{r}
pd <- df_country %>%
  group_by(date) %>%
  summarise(n = sum(deaths)) %>%
  filter(date < max(date))
ggplot(data = pd,
       aes(x = date,
           y = n)) +
  geom_point() +
  theme_simple() +
  labs(x = 'Date',
       y = 'Deaths',
       title = 'COVID-19 deaths')
# ggsave('~/Videos/update/e.png',
#        width = 1280 / 150,
#        height = 720 / 150)
```




## China vs world deaths

```{r}
pd <- df_country %>%
  group_by(date,
           country = ifelse(country == 'China', 'China', 'Other countries')) %>%
  summarise(n = sum(deaths))  %>%
  ungroup %>%
  filter(date < max(date))
ggplot(data = pd,
       aes(x = date,
           y = n,
           color = country)) +
  geom_line(size = 2) +
  # geom_point() +
  theme_simple() +
  labs(x = 'Date',
       y = 'Deaths',
       title = 'COVID-19 deaths') +
  scale_color_manual(name = '',
                     values = c('red', 'black')) +
  theme(legend.text = element_text(size = 25),
        legend.position = 'top')
# ggsave('~/Videos/update/f.png',
#        width = 1280 / 150,
#        height = 720 / 150)
```


## Asian hope

```{r}
plot_day_zero(countries = c('Korea, South', 'Japan', 'China', 'Singapore'))
# ggsave('~/Videos/update/g.png',
#        width = 1280 / 150,
#        height = 720 / 150)
```



Since trajectories are very unstable when cases are low, we'll exclude from our analysis the first few days, and will only count as "outbreak" once a country reaches 150 or more cumulative cases.

```{r}
# Doubling time
n_cases_start = 150
countries = c('Italy', 'Spain', 'France', 'Germany', 'Italy', 'Switzerland', 'Denmark', 'US', 'United Kingdom', 'Norway')
# countries <- sort(unique(df_country$country))
out_list <- curve_list <-  list()
counter <- 0
for(i in 1:length(countries)){
  message(i)
  this_country <- countries[i]
  sub_data <-df_country %>% filter(country == this_country)
  # Only calculate on countries with n_cases_start or greater cases,
  # starting at the first day at n_cases_start or greater
  ok <- max(sub_data$cases, na.rm = TRUE) >= n_cases_start
  if(ok){
    counter <- counter + 1
    pd <- sub_data %>%
      filter(!is.na(cases)) %>%
      mutate(start_date = min(date[cases >= n_cases_start])) %>%
      mutate(days_since = date - start_date) %>%
      filter(days_since >= 0) %>%
      mutate(days_since = as.numeric(days_since))
    fit <- lm(log(cases) ~ days_since, data = pd) 
    # plot(pd$days_since, log(pd$cases))
    # abline(fit)
    ## Slope
    curve <- fit$coef[2]
    
    # Predict days ahead
    fake <- tibble(days_since = seq(0, max(pd$days_since) + 5, by = 1))
    fake <- left_join(fake, pd %>% dplyr::select(days_since, cases, date))
    fake$predicted <- exp(predict(fit, newdata = fake))
    
    # Doubling time
    dt <- log(2)/fit$coef[2]
    out <- tibble(country = this_country,
                  doubling_time = dt,
                  slope = curve)
    out_list[[counter]] <- out
    curve_list[[counter]] <- fake %>% mutate(country = this_country)
  }
}
done <- bind_rows(out_list)
print(done)
curves <- bind_rows(curve_list)
# Get curves back in exponential form
# curves$curve <- exp(curves$curve)

# Join doubling time to curves
joined <- left_join(curves, done)

# Get rid of Italy future (since it's the "leader")
joined <- joined %>%
  filter(country != 'Italy' |
           date <= (Sys.Date() -1))


# Make long format
long <- joined %>% 
  dplyr::select(date, days_since, country, cases, predicted, doubling_time) %>%
  tidyr::gather(key, value, cases:predicted) %>%
  mutate(key = Hmisc::capitalize(gsub('_', ' ', key))) %>%
  mutate(key = ifelse(key == 'Predicted', 'Predicted (based on current doubling time)', key))
```


The below chart shows the trajectories in terms of number of cases in Europe in red, and the predicted trajectories in black. The black line assumes that the doubling rate will stay constant.



```{r, fig.height = 9}
cols <- c('red', 'black')
ggplot(data = long,
       aes(x = days_since,
           y = value,
           lty = key,
           color = key)) +
  geom_line(data = long %>% filter(key != 'Confirmed cases'),
            size = 1.2, alpha = 0.8) +
  geom_point(data = long %>% filter(key == 'Confirmed cases')) +
  geom_line(data = long %>% filter(key == 'Confirmed cases'),
            size = 0.8) +
  facet_wrap(~paste0(country, '\n',
                     '(doubling time: ', 
                     round(doubling_time, digits = 1), ' days)'), scales = 'free') +
  theme_simple() +
  scale_linetype_manual(name ='',
                        values = c(1,2)) +
  scale_color_manual(name = '',
                     values = cols) +
  theme(legend.position = 'top') +
  labs(x = 'Days since first day at >150 cumulative cases',
       y = 'Cases',
       title = 'COVID-19 CASES: ("predicted" assumes no change in doubling time)',
       caption = 'Data from Johns Hopkins. Processing: Joe Brew @joethebrew. Code: github.com/databrew/covid19',
       subtitle = '(Doubling time calculated since first day at >150 cumulative cases)') +
    theme(strip.text = element_text(size = 13),
          plot.title = element_text(size = 15))
```

Since Italy is "leading the way", it's helpful to also compare each country to Italy. Let's see that. 

```{r, fig.height = 9}
# Overlay Italy
ol1 <- joined %>% filter(!country %in% 'Italy')
ol2 <- joined %>% filter(country == 'Italy') %>% dplyr::rename(Italy = cases) %>%
  dplyr::select(Italy, days_since)
ol <- left_join(ol1, ol2) %>%
  dplyr::select(days_since, date, country, cases, predicted, Italy,doubling_time)
ol <- tidyr::gather(ol, key, value, cases: Italy) %>%
  mutate(key = Hmisc::capitalize(gsub('_', ' ', key))) %>%
  mutate(key = ifelse(key == 'Predicted', 'Predicted (based on current doubling time)', key))

cols <- c('red', 'blue', 'black')
ggplot(data = ol,
       aes(x = days_since,
           y = value,
           lty = key,
           color = key)) +
  geom_line(data = ol %>% filter(!key %in% c('Confirmed cases', 'Italy')),
            size = 1.2, alpha = 0.8) +
    geom_line(data = ol %>% filter(key %in% c('Italy')),
            size = 0.8, alpha = 0.8) +
  geom_point(data = ol %>% filter(key == 'Confirmed cases')) +
  geom_line(data = ol %>% filter(key == 'Confirmed cases'),
            size = 0.8) +
  facet_wrap(~paste0(country, '\n',
                     '(doubling time: ', 
                     round(doubling_time, digits = 1), ' days)'), scales = 'free') +
  theme_simple() +
  scale_linetype_manual(name ='',
                        values = c(1,6,2)) +
  scale_color_manual(name = '',
                     values = cols) +
  theme(legend.position = 'top') +
  labs(x = 'Days since first day at >150 cumulative cases',
       y = 'Cases',
       title = 'COVID-19 CASES: ("predicted" assumes no change in doubling time)',
       caption = 'Data from Johns Hopkins. Processing: Joe Brew @joethebrew. Code: github.com/databrew/covid19',
       subtitle = '(Doubling time calculated since first day at >150 cumulative cases)') +
    theme(strip.text = element_text(size = 13),
          plot.title = element_text(size = 15))
```

In the above, what's striking is how many places have trajectories that are _worse_ than Italy's. Yes, Italy has more cases, but it's doubling time is less. Either that changes soon, or these other countries will soon have more cases than Italy.


## Deaths or cases?

The number of cases is not necessarily the best indicator for the severity of an outbreak of this nature. Why? Because (a) testing rates and protocols are different by place and (b) testing rates are different by time (since health services are changing their approaches as things develop). In other words, when we compare the number of cases by place and time, we are introducing significant bias.

Using deaths to gauge the magnitude of the outbreak is also problematic. Death rates are differential by age, so the number of deaths depends on a country's population period, or age structure. Also, death rates will be a function of health services, which are not of the same quality every where. And, of course, like cases, we don't necessarily know about all of the deaths that occur because of COVID-19.

Still, there's an argument that death rates have _less_ bias than case rates because deaths are easier to identify than cases. Let's accept that argument, for the time being, and have a look at death rates by country.

```{r}
# Doubling time
n_deaths_start = 5
countries = c('Italy', 'Spain', 'France', 'Italy', 'Switzerland', 'Denmark', 'US', 'United Kingdom', 'Norway', 'Germany')
# countries <- sort(unique(df_country$country))

make_double_time <- function(data = df_country,
                             the_country = 'Spain',
                             n_deaths_start = 5,
                             time_ahead = 7){
   sub_data <-data %>% filter(country == the_country)
  # Only calculate on countries with n_cases_start or greater cases,
  # starting at the first day at n_cases_start or greater
  ok <- max(sub_data$deaths, na.rm = TRUE) >= n_deaths_start
  if(ok){
    counter <- counter + 1
    pd <- sub_data %>%
      filter(!is.na(deaths)) %>%
      mutate(start_date = min(date[deaths >= n_deaths_start])) %>%
      mutate(days_since = date - start_date) %>%
      filter(days_since >= 0) %>%
      mutate(days_since = as.numeric(days_since)) %>%
      mutate(the_weight = 1/(1 + (as.numeric(max(date) - date))))
    fit <- lm(log(deaths) ~ days_since,
              weights = the_weight,
              data = pd) 
    # fitlo <- loess(deaths ~ days_since, data = pd)
    # plot(pd$days_since, log(pd$cases))
    # abline(fit)
    ## Slope
    # curve <- fit$coef[2]
    
    # Predict days ahead
    day0 <- pd$date[pd$days_since == 0]
    fake <- tibble(days_since = seq(0, max(pd$days_since) + time_ahead, by = 1))
    fake <- fake %>%mutate(date = seq(day0, day0+max(fake$days_since), by = 1))
    fake <- left_join(fake, pd %>% dplyr::select(days_since, deaths, date))
    fake$predicted <- exp(predict(fit, newdata = fake))
    # fake$predictedlo <- predict(fitlo, newdata = fake)
    ci <- exp(predict(fit, newdata = fake, interval = 'prediction'))
    # cilo <- predict(fitlo, newdata = fake, interval = 'prediction')

    fake$lwr <- ci[,'lwr']
    fake$upr <- ci[,'upr']
    # fake$lwrlo <- ci[,'lwr']
    # fake$uprlo <- ci[,'upr']
    # Doubling time
    dt <- log(2)/fit$coef[2]
    fake %>% mutate(country = the_country) %>% mutate(doubling_time = dt)
  }
}

plot_double_time <- function(data, ylog = F){
  the_labs <- labs(x = 'Date',
                   y = 'Deaths',
                   title = paste0('Predicted deaths in ', data$country[1]))
  long <- data %>%
    tidyr::gather(key, value, deaths:predicted) %>%
    mutate(key = Hmisc::capitalize(key))
  g <- ggplot() +
        geom_ribbon(data = data %>% filter(date > max(long$date[!is.na(long$value) & long$key == 'Deaths'])),
                aes(x = date,
                    ymax = upr,
                    ymin = lwr),
                alpha =0.6,
                fill = 'darkorange') +
    geom_line(data = long,
              aes(x = date,
                  y = value,
                  group = key,
                  lty = key)) +
    geom_point(data = long %>% filter(key == 'Deaths'),
               aes(x = date,
                   y = value)) +
    theme_simple() +
    theme(legend.position = 'right',
          legend.title = element_blank()) +
    the_labs
  if(ylog){
    g <- g + scale_y_log10()
  }
  return(g)
}
options(scipen = '999')
data <- make_double_time(n_deaths_start = 150, time_ahead = 7)
data
dir.create('/tmp/ccaa_predictions')

plot_double_time(data, ylog = T) +
  labs(subtitle = 'Basic log-linear model weighted at (1 + (1/ days ago)),\nassuming no change in growth trajectory since first day at >150 deaths')
ggsave('/tmp/ccaa_predictions/spain.png')
# All ccaas
ccaas <- sort(unique(esp_df$ccaa))
for(i in 1:length(ccaas)){
  message(i)
  this_ccaa <- ccaas[i]
  sub_data <- esp_df %>% mutate(country = ccaa) 
  try({
    data <- make_double_time(
    data = sub_data,
    the_country = this_ccaa,
    n_deaths_start = 5,
    time_ahead = 7)
  plot_double_time(data, ylog = T) +
  labs(subtitle = 'Basic log-linear model weighted at (1 + (1/ days ago)), assuming no change in growth trajectory since first day at >5 deaths')
  ggsave(paste0('/tmp/ccaa_predictions/',
                this_ccaa, '.png'),
         height = 4.9,
         width = 8.5)
  })

}

```

```{r}
# all_countries <- sort(unique(df_country$country))
# for(i in 1:length(all_countries)){
#   this_country <- all_countries[i]
#   data <- make_double_time(the_country = this_country, n_deaths_start = 5)
#   if(!is.null(data)){
#     # print(this_country)
#     g <- plot_double_time(data, ylog = F) +
#   labs(subtitle = 'Basic log-linear model assuming no change in growth trajectory since first day at >5 deaths')
#     ggsave(paste0('/tmp/', this_country, '.png'), height = 5, width = 8)
#     print(data)
#   }
# }
```

```{r}
counter <- 0
# Africa
data <- make_double_time(the_country = 'South Africa',
                         n_deaths_start = 5, time_ahead = 7)
data
dir.create('/tmp/africa_predictions')

plot_double_time(data, ylog = T) +
  labs(subtitle = 'Basic log-linear model weighted at (1 + (1/ days ago)),\nassuming no change in growth trajectory since first day at >150 deaths')

```

```{r}
out_list <- curve_list <-  list()
counter <- 0
for(i in 1:length(countries)){
  message(i)
  this_country <- countries[i]
  sub_data <-df_country %>% filter(country == this_country)
  # Only calculate on countries with n_cases_start or greater cases,
  # starting at the first day at n_cases_start or greater
  ok <- max(sub_data$deaths, na.rm = TRUE) >= n_deaths_start
  if(ok){
    counter <- counter + 1
    pd <- sub_data %>%
      filter(!is.na(deaths)) %>%
      mutate(start_date = min(date[deaths >= n_deaths_start])) %>%
      mutate(days_since = date - start_date) %>%
      filter(days_since >= 0) %>%
      mutate(days_since = as.numeric(days_since))
    fit <- lm(log(deaths) ~ days_since, data = pd) 
    # plot(pd$days_since, log(pd$cases))
    # abline(fit)
    ## Slope
    # curve <- fit$coef[2]
    
    # Predict days ahead
    fake <- tibble(days_since = seq(0, max(pd$days_since) + 5, by = 1))
    fake <- left_join(fake, pd %>% dplyr::select(days_since, deaths, date))
    fake$predicted <- exp(predict(fit, newdata = fake))
    
    # Doubling time
    dt <- log(2)/fit$coef[2]
    out <- tibble(country = this_country,
                  doubling_time = dt)
    out_list[[counter]] <- out
    curve_list[[counter]] <- fake %>% mutate(country = this_country)
  }
}
done <- bind_rows(out_list)
curves <- bind_rows(curve_list)
# Get curves back in exponential form
# curves$curve <- exp(curves$curve)

# Join doubling time to curves
joined <- left_join(curves, done)

# Get rid of Italy future (since it's the "leader")
joined <- joined %>%
  filter(country != 'Italy' |
           date <= (Sys.Date() -1))


# Make long format
long <- joined %>% 
  dplyr::select(date, days_since, country, deaths, predicted, doubling_time) %>%
  tidyr::gather(key, value, deaths:predicted) %>%
  mutate(key = Hmisc::capitalize(gsub('_', ' ', key))) %>%
  mutate(key = ifelse(key == 'Predicted', 'Predicted (based on current doubling time)', key))
```

```{r, fig.height = 7}
cols <- c('red', 'black')
sub_data <-  long %>% filter(country != 'US')
ggplot(data = sub_data,
       aes(x = days_since,
           y = value,
           lty = key,
           color = key)) +
  geom_line(data = sub_data %>% filter(key != 'Deaths'),
            size = 1.2, alpha = 0.8) +
  geom_point(data = sub_data %>% filter(key == 'Deaths')) +
  geom_line(data = sub_data %>% filter(key == 'Deaths'),
            size = 0.8) +
  facet_wrap(~paste0(country, '\n',
                     '(doubling time: ', 
                     round(doubling_time, digits = 1), ' days)'), scales = 'free') +
  theme_simple() +
  scale_linetype_manual(name ='',
                        values = c(1,2)) +
  scale_color_manual(name = '',
                     values = cols) +
  theme(legend.position = 'top') +
  labs(x = 'Days since first day at >5 cumulative deaths',
       y = 'Deaths',
       title = 'COVID-19 DEATHS: ("predicted" assumes no change in doubling time)',
       caption = 'Data from Johns Hopkins. Processing: Joe Brew @joethebrew. Code: github.com/databrew/covid19',
       subtitle = '(Doubling time calculated since first day at >5 cumulative deaths)') +
    theme(strip.text = element_text(size = 13),
          plot.title = element_text(size = 15))
```




Let's again overlay Italy.

```{r, fig.height = 7}
# Overlay Italy
ol1 <- joined %>% filter(!country %in% 'Italy')
ol2 <- joined %>% filter(country == 'Italy') %>% dplyr::rename(Italy = deaths) %>%
  dplyr::select(Italy, days_since)
ol <- left_join(ol1, ol2) %>%
  dplyr::select(days_since, date, country, deaths, predicted, Italy,doubling_time)
ol <- tidyr::gather(ol, key, value, deaths: Italy) %>%
  mutate(key = Hmisc::capitalize(gsub('_', ' ', key))) %>%
  mutate(key = ifelse(key == 'Predicted', 'Predicted (based on current doubling time)', key))

cols <- c('red', 'blue', 'black')
sub_data <- ol %>% 
  filter(!(key == 'Predicted (based on current doubling time)' &
             country == 'Spain' &
             days_since > 13))
ggplot(data = sub_data,
       aes(x = days_since,
           y = value,
           lty = key,
           color = key)) +
  geom_line(data = sub_data %>% filter(!key %in% c('Deaths', 'Italy')),
            size = 1.2, alpha = 0.8) +
    geom_line(data = sub_data %>% filter(key %in% c('Italy')),
            size = 0.8, alpha = 0.8) +
  geom_point(data = sub_data %>% filter(key == 'Deaths')) +
  geom_line(data = sub_data %>% filter(key == 'Deaths'),
            size = 0.8) +
  facet_wrap(~paste0(country, '\n',
                     '(doubling time: ', 
                     round(doubling_time, digits = 1), ' days)'), scales = 'free') +
  theme_simple() +
  scale_linetype_manual(name ='',
                        values = c(1,6,2)) +
  scale_color_manual(name = '',
                     values = cols) +
  scale_y_log10() +
  theme(legend.position = 'top') +
  labs(x = 'Days since first day at >5 deaths',
       y = 'Deaths',
       title = 'COVID-19 DEATHS: ("predicted" assumes no change in doubling time)',
       caption = 'Data from Johns Hopkins. Processing: Joe Brew @joethebrew. Code: github.com/databrew/covid19',
       subtitle = '(Doubling time calculated since first day at >5 cumulative deaths)') +
    theme(strip.text = element_text(size = 13),
          plot.title = element_text(size = 15)) 
```




Let's look just at Spain

```{r, fig.height = 7}
# Overlay Italy
ol1 <- joined %>% filter(!country %in% 'Italy',
                         country == 'Spain')
ol2 <- joined %>% filter(country == 'Italy') %>% dplyr::rename(Italy = deaths) %>%
  dplyr::select(Italy, days_since)
ol <- left_join(ol1, ol2) %>%
  dplyr::select(days_since, date, country, deaths, predicted, Italy,doubling_time)
ol <- tidyr::gather(ol, key, value, deaths: Italy) %>%
  mutate(key = Hmisc::capitalize(gsub('_', ' ', key))) %>%
  mutate(key = ifelse(key == 'Predicted', 'Predicted (based on current doubling time)', 
                      ifelse(key == 'Deaths', 'Spain', key)))

cols <- c('blue',  'black', 'red')
ggplot(data = ol,
       aes(x = days_since,
           y = value,
           lty = key,
           color = key)) +
  geom_line(data = ol %>% filter(!key %in% c('Deaths', 'Italy')),
            size = 1.2, alpha = 0.8) +
    geom_line(data = ol %>% filter(key %in% c('Italy')),
            size = 0.8, alpha = 0.8) +
  # geom_point(data = ol %>% filter(key == 'Deaths')) +
    geom_point(data = ol %>% filter(country == 'Spain',
                                    key == 'Spain'), size = 4, alpha = 0.6) +

  geom_line(data = ol %>% filter(key == 'Deaths'),
            size = 0.8) +
  # facet_wrap(~paste0(country, '\n',
  #                    '(doubling time: ', 
  #                    round(doubling_time, digits = 1), ' days)'), scales = 'free') +
  theme_simple() +
  scale_linetype_manual(name ='',
                        values = c(1,6,1)) +
  scale_color_manual(name = '',
                     values = cols) +
  scale_y_log10() +
  theme(legend.position = 'top') +
  labs(x = 'Days since first day at >5 deaths',
       y = 'Deaths',
       title = 'COVID-19 DEATHS: ("predicted" assumes no change in doubling time)',
       caption = 'Data from Johns Hopkins. Processing: Joe Brew @joethebrew. Code: github.com/databrew/covid19',
       subtitle = '(Doubling time calculated since first day at >5 cumulative deaths)') +
    theme(strip.text = element_text(size = 13),
          plot.title = element_text(size = 15),
          axis.title = element_text(size = 18))
```



## The importance of lag

Things are changing very rapidly. And measures being taken by these countries will have an impact on the outbreak.

But it's important to remember that there is a lag between when an intervention takes place and when its effect is notable. Because of the incubation period - the number of days between someone getting infected and becoming sick - what we do today won't really have an effect until next weekend. And the clinical cases that present today are among people who got infected a week ago. 

Disease control measures work. We can see that clearly in the case of Hubei, Wuhan, Iran, Japan. And they will work in Europe too. But because many of these measures were implemented very recently, we won't likely see a major effect for at least a few more days.

In the mean time, it's important to practice social distancing. Stay away from others to keep both you and others safe. Listen to Health Authorities. Take this very seriously.

# Spain and Italy regions

```{r}
# Madrid vs Lombardy deaths
n_death_start <- 5
pd <- esp_df %>%
  # filter(ccaa == 'Madrid') %>%
  dplyr::select(date, ccaa, cases, deaths) %>%
  bind_rows(ita %>%
              # filter(ccaa == 'Lombardia') %>%
              dplyr::select(date, ccaa, cases, deaths)) %>%
  arrange(date) %>%
  group_by(ccaa) %>%
  mutate(first_n_death = min(date[deaths >= n_death_start])) %>%
  ungroup %>%
  mutate(days_since_n_deaths = date - first_n_death) %>%
  filter(is.finite(days_since_n_deaths))

pd$country <- pd$ccaa
pd$cases <- pd$cases
countries <- sort(unique(pd$country))
out_list <- curve_list <-  list()
counter <- 0
for(i in 1:length(countries)){
  message(i)
  this_country <- countries[i]
  sub_data <- pd %>% filter(country == this_country)
  # Only calculate on countries with n_cases_start or greater cases,
  # starting at the first day at n_cases_start or greater
  # ok <- max(sub_data$deaths, na.rm = TRUE) >= n_deaths_start
  ok <- length(which(sub_data$deaths >= n_deaths_start))
  if(ok){
    counter <- counter + 1
    sub_pd <- sub_data %>%
      filter(!is.na(deaths)) %>%
      mutate(start_date = min(date[deaths >= n_deaths_start])) %>%
      mutate(days_since = date - start_date) %>%
      filter(days_since >= 0) %>%
      mutate(days_since = as.numeric(days_since))
    fit <- lm(log(deaths) ~ days_since, data = sub_pd) 
    # plot(pd$days_since, log(pd$cases))
    # abline(fit)
    ## Slope
    # curve <- fit$coef[2]
    
    # Predict days ahead
    fake <- tibble(days_since = seq(0, max(sub_pd$days_since) + 5, by = 1))
    fake <- left_join(fake, sub_pd %>% dplyr::select(days_since, deaths, date))
    fake$predicted <- exp(predict(fit, newdata = fake))
    
    # Doubling time
    dt <- log(2)/fit$coef[2]
    out <- tibble(country = this_country,
                  doubling_time = dt)
    out_list[[counter]] <- out
    curve_list[[counter]] <- fake %>% mutate(country = this_country)
  }
}
done <- bind_rows(out_list)
curves <- bind_rows(curve_list)
# Get curves back in exponential form
# curves$curve <- exp(curves$curve)

# Join doubling time to curves
joined <- left_join(curves, done)


# Make long format
long <- joined %>% 
  dplyr::select(date, days_since, country, deaths, predicted, doubling_time) %>%
  tidyr::gather(key, value, deaths:predicted) %>%
  mutate(key = Hmisc::capitalize(gsub('_', ' ', key))) %>%
  mutate(key = ifelse(key == 'Predicted', 'Predicted (based on current doubling time)', key))

# Remove those with not enough data to have a doubling time yet
long <- long %>% filter(!is.na(doubling_time))
```

```{r, fig.height = 7}
text_size <- 12

cols <- c('red', 'black')
ggplot(data = long,
       aes(x = days_since,
           y = value,
           lty = key,
           color = key)) +
  geom_line(data = long %>% filter(key != 'Deaths'),
            size = 1.2, alpha = 0.8) +
  geom_point(data = long %>% filter(key == 'Deaths')) +
  geom_line(data = long %>% filter(key == 'Deaths'),
            size = 0.8) +
  facet_wrap(~paste0(country, '\n',
                     '(doubling time: ', 
                     round(doubling_time, digits = 1), ' days)'), scales = 'free') +
  theme_simple() +
  scale_y_log10() +
  scale_linetype_manual(name ='',
                        values = c(1,2)) +
  scale_color_manual(name = '',
                     values = cols) +
  theme(legend.position = 'top') +
  labs(x = 'Days since first day at >150 cumulative cases',
       y = 'Deaths',
       title = 'COVID-19 DEATHS: ("predicted" assumes no change in doubling time)',
       caption = 'Data from Johns Hopkins. Processing: Joe Brew @joethebrew. Code: github.com/databrew/covid19',
       subtitle = '(Doubling time calculated since first day at >5 cumulative deaths)') +
    theme(strip.text = element_text(size = text_size * 0.5),
          plot.title = element_text(size = 15))
```


Let's overlay Lombardy


```{r, fig.height = 9, fig.width = 10}
# Overlay Italy
ol1 <- joined %>% filter(!country %in% 'Lombardia')
ol2 <- joined %>% filter(country == 'Lombardia') %>% dplyr::rename(Lombardia = deaths) %>%
  dplyr::select(Lombardia, days_since)
ol <- left_join(ol1, ol2) %>%
  dplyr::select(days_since, date, country, deaths, predicted, Lombardia,doubling_time)
ol <- tidyr::gather(ol, key, value, deaths: Lombardia) %>%
  mutate(key = Hmisc::capitalize(gsub('_', ' ', key))) %>%
  mutate(key = ifelse(key == 'Predicted', 'Predicted (based on current doubling time)', key))

# Remove those with not enough data to have a doubling time yet
ol <- ol %>% filter(!is.na(doubling_time))

cols <- c('red', 'blue', 'black')
ggplot(data = ol,
       aes(x = days_since,
           y = value,
           lty = key,
           color = key)) +
  scale_y_log10() +
  geom_line(data = ol %>% filter(!key %in% c('Deaths', 'Italy')),
            size = 1.2, alpha = 0.8) +
    geom_line(data = ol %>% filter(key %in% c('Lombardia')),
            size = 0.5, alpha = 0.8) +
  geom_point(data = ol %>% filter(key == 'Deaths')) +
  geom_line(data = ol %>% filter(key == 'Deaths'),
            size = 0.8) +
  facet_wrap(~paste0(country, '\n',
                     '(doubling time: ', 
                     round(doubling_time, digits = 1), ' days)'), scales = 'free') +
  theme_simple() +
  scale_linetype_manual(name ='',
                        values = c(1,6,2)) +
  scale_color_manual(name = '',
                     values = cols) +
  theme(legend.position = 'top') +
  labs(x = 'Days since first day at >5 deaths',
       y = 'Deaths',
       title = 'COVID-19 DEATHS: ("predicted" assumes no change in doubling time)',
       caption = 'Data from Johns Hopkins. Processing: Joe Brew @joethebrew. Code: github.com/databrew/covid19',
       subtitle = '(Doubling time calculated since first day at >5 cumulative deaths)') +
    theme(strip.text = element_text(size = text_size * 0.5),
          plot.title = element_text(size = 15))
```

Show only Spanish regions vs. Lombardy


```{r, fig.height = 9, fig.width = 10}
text_size <- 14

# Overlay Italy
ol1 <- joined %>% filter(!country %in% 'Lombardia')
ol2 <- joined %>% filter(country == 'Lombardia') %>% dplyr::rename(Lombardia = deaths) %>%
  dplyr::select(Lombardia, days_since)
ol <- left_join(ol1, ol2) %>%
  dplyr::select(days_since, date, country, deaths, predicted, Lombardia,doubling_time)
ol <- tidyr::gather(ol, key, value, deaths: Lombardia) %>%
  mutate(key = Hmisc::capitalize(gsub('_', ' ', key))) %>%
  mutate(key = ifelse(key == 'Predicted', 'Predicted (based on current doubling time)', key))

# Remove those with not enough data to have a doubling time yet
ol <- ol %>% filter(!is.na(doubling_time))

# Only Spain
ol <- ol %>% filter(country %in% esp_df$ccaa) %>%
  filter(!country %in% 'Aragón')

cols <- c('red', 'blue', 'black')
ggplot(data = ol,
       aes(x = days_since,
           y = value,
           lty = key,
           color = key)) +
  scale_y_log10() +
  geom_line(data = ol %>% filter(!key %in% c('Deaths', 'Lombardia')),
            size = 1.2, alpha = 0.8) +
    geom_line(data = ol %>% filter(key %in% c('Lombardia')),
            size = 0.5, alpha = 0.8) +
  geom_point(data = ol %>% filter(key == 'Deaths')) +
  geom_line(data = ol %>% filter(key == 'Deaths'),
            size = 0.8) +
  facet_wrap(~paste0(country, '\n',
                     '(doubling time: ', 
                     round(doubling_time, digits = 1), ' days)'), scales = 'free') +
  theme_simple() +
  scale_linetype_manual(name ='',
                        values = c(1,6,2)) +
  scale_color_manual(name = '',
                     values = cols) +
  theme(legend.position = 'top') +
  labs(x = 'Days since first day at >5 deaths',
       y = 'Deaths',
       title = 'COVID-19 DEATHS: ("predicted" assumes no change in doubling time)',
       caption = 'Data from Johns Hopkins. Processing: Joe Brew @joethebrew. Code: github.com/databrew/covid19',
       subtitle = '(Doubling time calculated since first day at >5 cumulative deaths)') +
    theme(strip.text = element_text(size = text_size * 0.6),
          plot.title = element_text(size = 15))
```

Same plot but overlayed


Same as above, but overlaid

```{r, fig.height = 7}
text_size <-10

# cols <- c('red', 'black')
long <- long %>% filter(country %in% c('Lombardia',
                                       'Emilia Romagna') |
                          country %in% esp_df$ccaa) %>%
  filter(country != 'Aragón')
places <- sort(unique(long$country))

cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 7, 'Spectral'))(length(places))
cols[which(places == 'Madrid')] <- 'red'
cols[which(places == 'Cataluña')] <- 'purple'
cols[which(places == 'Lombardia')] <- 'darkorange'
cols[which(places == 'Emilia Romagna')] <- 'darkgreen'

long$key <- ifelse(long$key != 'Deaths', 'Predicted', long$key)
long$key <- ifelse(long$key == 'Predicted', 'Muertes\nprevistas',
                   'Muertes\nobservadas')


# Keep only Madrid, Lombardy, Emilia Romagna
long <- long %>%
  filter(country %in% c('Madrid',
                        'Lombardia',
                        'Emilia Romagna'))

ggplot(data = long,
       aes(x = days_since,
           y = value,
           lty = key,
           color = country)) +
  geom_point(data = long %>% filter(key == 'Muertes\nobservadas'), size = 2, alpha = 0.8) +
  geom_line(data = long %>% filter(key == 'Muertes\nprevistas'), size = 1, alpha = 0.7) +
  geom_line(data = long %>% filter(key != 'Muertes\nprevistas'), size = 0.8) +
  theme_simple() +
  scale_y_log10() +
  scale_linetype_manual(name ='',
                        values = c(1,4)) +
  scale_color_manual(name = '',
                     values = cols) +
  theme(legend.position = 'top') +
  # labs(x = 'Days since first day at 5 or more cumulative deaths',
  #      y = 'Deaths',
  #      title = 'COVID-19 DEATHS: ("predicted" assumes no change in doubling time)',
  #      caption = 'Data from Johns Hopkins. Processing: Joe Brew @joethebrew. Code: github.com/databrew/covid19',
  #      subtitle = '(Doubling time calculated since first day at >5 cumulative deaths)') +
    labs(x = 'Días desde el primer día a 5 o más muertes acumuladas',
       y = 'Muertes (escala logarítmica)',
       title = 'Muertes por COVID-19',
       caption = 'Data from Johns Hopkins. Processing: Joe Brew @joethebrew. Code: github.com/databrew/covid19',
       subtitle = '(Tasa de crecimiento calculada desde el primer día a 5 o más muertes acumuladas)\n(Muertes "previstas": suponiendo que no hay cambios en la tasa de crecimiento)') +
    theme(strip.text = element_text(size = text_size * 0.75),
          plot.title = element_text(size = text_size * 3),
          legend.text = element_text(size = text_size * 1.5),
          axis.title = element_text(size = text_size * 2),
          axis.text = element_text(size = text_size * 2))
```



```{r}
# cols <- c(cols, 'darkorange')
# ggplot(data = ol,
#        aes(x = days_since,
#            y = value,
#            lty = key,
#            color = key)) +
#   scale_y_log10() +
#   geom_line(aes(color = country)) +
#   
#   # geom_line(data = ol %>% filter(!key %in% c('Deaths', 'Italy')),
#   #           size = 1.2, alpha = 0.8) +
#   #   geom_line(data = ol %>% filter(key %in% c('Lombardia')),
#   #           size = 0.5, alpha = 0.8) +
#   # geom_point(data = ol %>% filter(key == 'Deaths')) +
#   # geom_line(data = ol %>% filter(key == 'Deaths'),
#   #           size = 0.8) +
#   theme_simple() +
#   scale_linetype_manual(name ='',
#                         values = c(1,6,2)) +
#   scale_color_manual(name = '',
#                      values = cols) +
#   theme(legend.position = 'top') +
#   labs(x = 'Days since first day at >5 deaths',
#        y = 'Deaths',
#        title = 'COVID-19 DEATHS: ("predicted" assumes no change in doubling time)',
#        caption = 'Data from Johns Hopkins. Processing: Joe Brew @joethebrew. Code: github.com/databrew/covid19',
#        subtitle = '(Doubling time calculated since first day at >5 cumulative deaths)') +
#     theme(strip.text = element_text(size = text_size * 1),
#           plot.title = element_text(size = 15))
```







```{r}
# Map data preparation

if(!'map.RData' %in% dir()){
  esp1 <- getData(name = 'GADM', country = 'ESP', level = 1)
# Remove canary
esp1 <- esp1[esp1@data$NAME_1 != 'Islas Canarias',]
espf <- fortify(esp1, region = 'NAME_1')

# Standardize names
# Convert names
map_names <- esp1@data$NAME_1
data_names <- sort(unique(esp_df$ccaa))
names_df <- tibble(NAME_1 = c('Andalucía',
 'Aragón',
 'Cantabria',
 'Castilla-La Mancha',
 'Castilla y León',
 'Cataluña',
 'Ceuta y Melilla',
 'Comunidad de Madrid',
 'Comunidad Foral de Navarra',
 'Comunidad Valenciana',
 'Extremadura',
 'Galicia',
 'Islas Baleares',
 'La Rioja',
 'País Vasco',
 'Principado de Asturias',
 'Región de Murcia'),
 ccaa = c('Andalucía',
 'Aragón',
 'Cantabria',
 'CLM',
 'CyL',
 'Cataluña',
 'Melilla',
 'Madrid',
 'Navarra',
 'C. Valenciana',
 'Extremadura',
 'Galicia',
 'Baleares',
 'La Rioja',
 'País Vasco',
 'Asturias',
 'Murcia'))


espf <- left_join(espf %>% dplyr::rename(NAME_1 = id), names_df)
centroids <- data.frame(coordinates(esp1))
names(centroids) <- c('long', 'lat')
centroids$NAME_1 <- esp1$NAME_1
centroids <- centroids %>% left_join(names_df)

# Get random sampling points

  random_list <- list()
  for(i in 1:nrow(esp1)){
    message(i)
    shp <- esp1[i,]
    # bb <- bbox(shp)
    this_ccaa <- esp1@data$NAME_1[i]
    # xs <- runif(n = 500, min = bb[1,1], max = bb[1,2])
    # ys <- runif(n = 500, min = bb[2,1], max = bb[2,2])
    # random_points <- expand.grid(long = xs, lat = ys) %>%
    #   mutate(x = long,
    #          y = lat)
    # coordinates(random_points) <- ~x+y
    # proj4string(random_points) <- proj4string(shp)
    # get ccaa
    message('getting locations of randomly generated points')
    # polys <- over(random_points,polygons(shp))
    # polys <- as.numeric(polys)
    random_points <- spsample(shp, n = 20000, type = 'random')
    random_points <- data.frame(random_points)
    random_points$NAME_1 <-  this_ccaa
    random_points <- left_join(random_points, names_df) %>% dplyr::select(-NAME_1)
    random_list[[i]] <- random_points
  }
  random_points <- bind_rows(random_list)
  random_points <- random_points %>% mutate(long = x,
                                            lat = y)

save(espf,
     esp1,
     names_df,
     centroids,
     random_points,
     file = 'map.RData')
} else {
  load('map.RData')
}

# Define a function for adding zerio
add_zero <- 
  function (x, n) 
  {
    x <- as.character(x)
    adders <- n - nchar(x)
    adders <- ifelse(adders < 0, 0, adders)
    for (i in 1:length(x)) {
      if (!is.na(x[i])) {
        x[i] <- paste0(paste0(rep("0", adders[i]), collapse = ""), 
                       x[i], collapse = "")
      }
    }
    return(x)
  }
```


```{r}
remake_world_map <- FALSE
options(scipen = '999')
if(remake_world_map){
  # World map animation
  world <- map_data('world')
  # world <- ne_countries(scale = "medium", returnclass = "sf")
  
  # Get plotting data
  pd <- df_country %>%
    dplyr::select(date, lng, lat, n = cases)
  dates <- sort(unique(pd$date))
  n_days <- length(dates)
  # # Define vectors for projection
  # vec_lon <- seq(30, -20, length = n_days)
  # vec_lat <- seq(25, 15, length = n_days)
  
  dir.create('animation')
  for(i in 1:n_days){
    message(i, ' of ', n_days)
    this_date <- dates[i]
    # this_lon <- vec_lon[i]
    # this_lat <- vec_lat[i]
    # the_crs <-
    #   paste0("+proj=laea +lat_0=", this_lat,
    #          " +lon_0=",
    #          this_lon,
    #          " +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs ")
    sub_data <- pd %>%
      filter(date == this_date)
    # coordinates(sub_data) <- ~lng+lat
    # proj4string(sub_data) <- proj4string(esp1)
    # # sub_data <- spTransform(sub_data,
    # #                         the_crs)
    # coordy <- coordinates(sub_data)
    # sub_data@data$long <- coordy[,1]
    # sub_data@data$lat <- coordy[,2]
  
    g <- ggplot() +
      geom_polygon(data = world,
                   aes(x = long,
                       y = lat,
                       group = group),
                   fill = 'black',
                   color = 'white',
                   size = 0.1) +
      theme_map() +
          geom_point(data = sub_data %>% filter(n > 0) %>% mutate(Deaths = n),
                 aes(x = lng,
                     y = lat,
                     size = Deaths),
                 color = 'red',
                 alpha = 0.6) +
      geom_point(data = tibble(x = c(0,0), y = c(0,0), Deaths = c(1, 100000)),
                 aes(x = x,
                     y = y,
                     size = Deaths),
                 color = 'red',
                 alpha =0.001) +
      scale_size_area(name = '', breaks = c(100, 1000, 10000, 100000),
                      max_size = 25
                      ) +
    # scale_size_area(name = '', limits = c(1, 10), breaks = c(0, 10, 30, 50, 70, 100, 200, 500)) +
      labs(title = this_date) +
      theme(plot.title = element_text(size = 30),
            legend.text = element_text(size = 15),
            legend.position = 'left')
  
    plot_number <- add_zero(i, 3)
    ggsave(filename = paste0('animation/', plot_number, '.png'),
           plot = g,
           width = 9.5,
           height = 5.1)
  }
  setwd('animation')
  system('convert -delay 30x100 -loop 0 *.png result.gif')
  setwd('..')

}

```

## Maps of Spain

```{r}
make_map <- function(var = 'deaths',
                     data = NULL,
                     pop = FALSE,
                     pop_factor = 100000,
                     points = FALSE,
                     line_color = 'white',
                     add_names = T,
                     add_values = T,
                     text_size = 2.7){
  
  if(is.null(data)){
    data <- esp_df %>%  mutate(ccaa = cat_transform(ccaa))

  }

  left <- espf %>%   mutate(ccaa = cat_transform(ccaa)) 
  right <- data[,c('ccaa', paste0(var, '_non_cum'))]
  

  names(right)[ncol(right)] <- 'var'
  right <- right %>% group_by(ccaa) %>% summarise(var = sum(var, na.rm = T))
  
  if(pop){
    right <- left_join(right, esp_pop)
    right$var <- right$var / right$pop * pop_factor
  }
  map <- left_join(left, right)
  
  if(points){
    the_points <- centroids %>%
      left_join(right)
    g <- ggplot() +
      geom_polygon(data = map,
         aes(x = long,
             y = lat,
             group = group),
         fill = 'black',
         color = line_color,
         lwd = 0.4, alpha = 0.8) +
      geom_point(data = the_points,
                 aes(x = long,
                     y = lat,
                     size = var),
                 color = 'red',
                 alpha = 0.7) +
      scale_size_area(name = '', max_size = 20)
  } else {
    # cols <- c('#008080','#70a494','#b4c8a8','#f6edbd','#edbb8a','#de8a5a','#ca562c')
    cols <- RColorBrewer::brewer.pal(n = 8, name = 'Blues')
    g <- ggplot(data = map,
         aes(x = long,
             y = lat,
             group = group)) +
    geom_polygon(aes(fill = var),
                 lwd = 0.3,
                 color = line_color) +
      scale_fill_gradientn(name = '',
                           colours = cols)
    # scale_fill_viridis(name = '' ,option = 'magma',
    #                    direction = -1) 
  }
  
  # Add names?
  if(add_names){
    centy <- centroids %>% left_join(right)
    if(add_values){
      centy$label <- paste0(centy$ccaa, '\n(', round(centy$var, digits = 2), ')')
    } else {
      centy$label <- centy$ccaa
    }

    g <- g +
      geom_text(data = centy,
                aes(x = long,
                    y = lat,
                    label = label,
                    group = ccaa),
                alpha = 0.7,
                size = text_size)
  }
  
  g +
    theme_map() +
    labs(subtitle = paste0('Data as of ', max(data$date))) +
    theme(legend.position = 'right')
  
}

make_dot_map <- function(var = 'deaths',
                     date = NULL,
                     pop = FALSE,
                     pop_factor = 100,
                     point_factor = 1,
                     points = FALSE,
                     point_color = 'darkred',
                     point_size = 0.6,
                     point_alpha = 0.5){
  
  
  if(is.null(date)){
    the_date <- max(esp_df$date)
  } else {
    the_date <- date
  }
    right <- esp_df[esp_df$date == the_date,c('ccaa', var)]
   names(right)[ncol(right)] <- 'var'
  if(pop){
    right <- left_join(right, esp_pop)
    right$var <- right$var / right$pop * pop_factor
  }
  map_data <- esp1@data %>%
    left_join(names_df) %>%
      left_join(right)
  map_data$var <- map_data$var / point_factor
  out_list <- list()
  for(i in 1:nrow(map_data)){
    sub_data <- map_data[i,]
    this_value = round(sub_data$var)

    if(this_value >= 1){
      this_ccaa = sub_data$ccaa
      # get some points
      sub_points <- random_points %>% filter(ccaa == this_ccaa)
      sampled_points <- sub_points %>% dplyr::sample_n(this_value)
      out_list[[i]] <- sampled_points
    }
  }
  the_points <- bind_rows(out_list)
  
  g <- ggplot() +
    geom_polygon(data = espf,
         aes(x = long,
             y = lat,
             group = group),
         fill = 'white',
         color = 'black',
         lwd = 0.4, alpha = 0.8) +
    geom_point(data = the_points,
               aes(x = long,
                   y = lat),
               color = point_color,
               size = point_size,
               alpha = point_alpha)
  g +
    theme_map() +
    labs(subtitle = paste0('Data as of ', max(esp_df$date)))
  
}

```

### Deaths

#### Absolute number of deaths: points

```{r}
make_map(var = 'deaths',
       points = T) +
  labs(title = 'Number of deaths',
       caption = '@joethebrew')
```

#### Absolute number of deaths: choropleth

```{r}
make_map(var = 'deaths',
         line_color = 'darkgrey',
       points = F) +
  labs(title = 'Number of deaths',
       caption = '@joethebrew')
```


#### Number of deaths adjusted by population: points

```{r}
make_map(var = 'deaths', pop = TRUE, points = T) +
  labs(title = 'Number of deaths per 100,000',
       caption = '@joethebrew')
```


#### Number of deaths adjusted by population: polygons

```{r}
make_map(var = 'deaths', pop = TRUE, points = F, line_color = 'darkgrey') +
  labs(title = 'Number of deaths per 100,000',
       caption = '@joethebrew')
```

#### Number of deaths: 1 dot per death

```{r}
make_dot_map(var = 'deaths', point_size = 0.05) +
  labs(title = 'COVID-19 deaths: 1 point = 1 death\nImportant: points are random within each CCAA; do not reflect exact location',
       caption = '@joethebrew')
```


### Cases

#### Absolute number of cases: points

```{r}
make_map(var = 'cases',
       points = T) +
  labs(title = 'Number of confirmed cases',
       caption = '@joethebrew')
```

#### Absolute number of cases: choropleth

```{r}
make_map(var = 'cases',
         line_color = 'darkgrey',
       points = F) +
  labs(title = 'Number of confirmed cases',
       caption = '@joethebrew')
```


#### Number of cases adjusted by population: points

```{r}
make_map(var = 'cases', pop = TRUE, points = T) +
  labs(title = 'Number of confirmed cases per 100,000',
       caption = '@joethebrew')
```


#### Number of cases adjusted by population: polygons

```{r}
make_map(var = 'cases', pop = TRUE, points = F,
         line_color = 'darkgrey') +
  labs(title = 'Number of confirmed cases per 100,000',
       caption = '@joethebrew')
```

#### Number of cases: points

```{r}
make_dot_map(var = 'cases',
             point_size = 0.05, point_alpha = 0.5, point_factor = 10) +
  labs(title = 'COVID-19 cases: 1 point = 10 cases\nImportant: points are random within each CCAA; do not reflect exact location',
       caption = '@joethebrew')
```

