---
title: "Maps"
author: "www.databrew.cc"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: "hide"
---


```{r setup, include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               # echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.width = 8,
               fig.height = 6,
               fig.path = 'figures/')
```

```{r}
## Load libraries
library(covid19)
library(ggplot2)
library(lubridate)
library(dplyr)
library(ggplot2)
library(sp)
library(raster)
library(viridis)
library(ggthemes)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(rayshader)
add_zero <- function (x, n) {
  x <- as.character(x)
  adders <- n - nchar(x)
  adders <- ifelse(adders < 0, 0, adders)
  for (i in 1:length(x)) {
    if (!is.na(x[i])) {
      x[i] <- paste0(paste0(rep("0", adders[i]), collapse = ""),
                     x[i], collapse = "")
    }
  }
  return(x)
}
```

```{r}
# Get Spanish province level data
df <- covid19::provinces
shp <- raster::getData(name = 'GADM', country = 'ESP', level = 2)

# Join
shp@data$ine <- as.numeric(as.character(shp@data$CC_2))
df$ine <- df$ine_code

# Make sure not less than previous day
df <- df %>%
  arrange(date) %>%
  group_by(ine) %>%
  mutate(cases_accumulated = 
           ifelse(cases_accumulated < dplyr::lag(cases_accumulated, 1),
                  dplyr::lag(cases_accumulated, 1),
                  cases_accumulated))

# Get centroids
centroids <- coordinates(shp)
centroids <- data.frame(centroids)
names(centroids) <- c('x', 'y')
centroids$province <- shp@data$NAME_2
centroids$ine <- shp@data$ine
centroids <- centroids %>%
  mutate(lng = x, lat = y)

# Get fortified shp
shpf <- fortify(shp, region = 'ine')
shpf$ine <- as.numeric(as.character(shpf$id))
# Loop through each day and get the joined data
dir.create('animation')
out_list <- list()
dates <- sort(unique(df$date))
dates <- dates[dates >= '2020-03-08']
counter <- 0
for(i in 1:length(dates)){
  this_date <- dates[i]
  the_data <- df %>% filter(date == this_date)
  the_data <- the_data %>% dplyr::select(-province) %>% left_join(centroids)
  the_data$value <- the_data$cases_accumulated
  the_data$value <- the_data$value / the_data$poblacion * 1000000
  the_data$value <- ifelse(is.na(the_data$value), 0 , the_data$value)
  the_data$value <- ifelse(is.infinite(the_data$value), 0 , the_data$value)
  

  # the_data$value <- the_data$value / sum(the_data$value, na.rm = TRUE) * 100
  the_data <- the_data %>% filter(lng > -12)
  the_shp <- shpf %>% left_join(the_data %>% dplyr::select(-lat, -lng))
  the_shp$value[is.na(the_shp$value)] <- 0
  if(max(the_data$value) > 0){
    counter <- counter + 1
    g <- ggplot() +
      geom_polygon(data = the_shp,
                   aes(x = long, y= lat, group = group,
                       fill = value),
                   color = NA) +
      # geom_point() +
      # geom_segment(aes(x = lng, y = lat, xend = lng, yend = lat + (0.01*(the_data$value^(1)))), alpha=0.5, color='#5A3E37', na.rm=TRUE, size = 4) +
      ggthemes::theme_map() +
      coord_equal(1) +
      labs(title = this_date,
           subtitle = 'Cumulative cases per milllion population') +
      xlim(-10, 4) +
      scale_fill_gradient(name = '',
                           low = 'lightyellow',  high = 'darkred',
                           limits = c(0, 24000),
                           breaks = seq(0, 24000, 6000)) +
      ylim(35, 44) +
      theme(legend.position = 'right')
    file_number <- add_zero(counter, 3) 
    ggsave(g, filename = paste0('animation/', file_number, '.png'))
  }
}

```

```
# Command line
cd animation
mogrify -resize 50% *.png
convert -delay 20 -loop 0 *.png result.gif

```

```{r}
# Map data preparation
if(!'map.RData' %in% dir()){
  esp1 <- getData(name = 'GADM', country = 'ESP', level = 1)
# Remove canary
esp1 <- esp1[esp1@data$NAME_1 != 'Islas Canarias',]
espf <- fortify(esp1, region = 'NAME_1')

# Standardize names
# Convert names
map_names <- esp1@data$NAME_1
data_names <- sort(unique(esp_df$ccaa))
names_df <- tibble(NAME_1 = c('Andalucía',
 'Aragón',
 'Cantabria',
 'Castilla-La Mancha',
 'Castilla y León',
 'Cataluña',
 'Ceuta y Melilla',
 'Comunidad de Madrid',
 'Comunidad Foral de Navarra',
 'Comunidad Valenciana',
 'Extremadura',
 'Galicia',
 'Islas Baleares',
 'La Rioja',
 'País Vasco',
 'Principado de Asturias',
 'Región de Murcia'),
 ccaa = c('Andalucía',
 'Aragón',
 'Cantabria',
 'CLM',
 'CyL',
 'Cataluña',
 'Melilla',
 'Madrid',
 'Navarra',
 'C. Valenciana',
 'Extremadura',
 'Galicia',
 'Baleares',
 'La Rioja',
 'País Vasco',
 'Asturias',
 'Murcia'))


espf <- left_join(espf %>% dplyr::rename(NAME_1 = id), names_df)
centroids <- data.frame(coordinates(esp1))
names(centroids) <- c('long', 'lat')
centroids$NAME_1 <- esp1$NAME_1
centroids <- centroids %>% left_join(names_df)

# Get random sampling points

  random_list <- list()
  for(i in 1:nrow(esp1)){
    message(i)
    shp <- esp1[i,]
    # bb <- bbox(shp)
    this_ccaa <- esp1@data$NAME_1[i]
    # xs <- runif(n = 500, min = bb[1,1], max = bb[1,2])
    # ys <- runif(n = 500, min = bb[2,1], max = bb[2,2])
    # random_points <- expand.grid(long = xs, lat = ys) %>%
    #   mutate(x = long,
    #          y = lat)
    # coordinates(random_points) <- ~x+y
    # proj4string(random_points) <- proj4string(shp)
    # get ccaa
    message('getting locations of randomly generated points')
    # polys <- over(random_points,polygons(shp))
    # polys <- as.numeric(polys)
    random_points <- spsample(shp, n = 20000, type = 'random')
    random_points <- data.frame(random_points)
    random_points$NAME_1 <-  this_ccaa
    random_points <- left_join(random_points, names_df) %>% dplyr::select(-NAME_1)
    random_list[[i]] <- random_points
  }
  random_points <- bind_rows(random_list)
  random_points <- random_points %>% mutate(long = x,
                                            lat = y)

save(espf,
     esp1,
     names_df,
     centroids,
     random_points,
     file = 'map.RData')
} else {
  load('map.RData')
}

# Define a function for adding zerio
add_zero <- function (x, n) {
    x <- as.character(x)
    adders <- n - nchar(x)
    adders <- ifelse(adders < 0, 0, adders)
    for (i in 1:length(x)) {
      if (!is.na(x[i])) {
        x[i] <- paste0(paste0(rep("0", adders[i]), collapse = ""), 
                       x[i], collapse = "")
      }
    }
    return(x)
}
options(scipen = '999')

```


## Maps of Spain

```{r}
make_map <- function(var = 'deaths',
                     data = NULL,
                     pop = FALSE,
                     pop_factor = 100000,
                     points = FALSE,
                     line_color = 'white',
                     add_names = T,
                     add_values = T,
                     text_size = 2.7){
  
  if(is.null(data)){
    data <- esp_df
  }

  left <- espf
  right <- data[,c('ccaa', paste0(var, '_non_cum'))]
  

  names(right)[ncol(right)] <- 'var'
  right <- right %>% group_by(ccaa) %>% summarise(var = sum(var, na.rm = T))
  
  if(pop){
    right <- left_join(right, esp_pop)
    right$var <- right$var / right$pop * pop_factor
  }
  map <- left_join(left, right)
  
  if(points){
    the_points <- centroids %>%
      left_join(right)
    g <- ggplot() +
      geom_polygon(data = map,
         aes(x = long,
             y = lat,
             group = group),
         fill = 'black',
         color = line_color,
         lwd = 0.4, alpha = 0.8) +
      geom_point(data = the_points,
                 aes(x = long,
                     y = lat,
                     size = var),
                 color = 'red',
                 alpha = 0.7) +
      scale_size_area(name = '', max_size = 20)
  } else {
    # cols <- c('#008080','#70a494','#b4c8a8','#f6edbd','#edbb8a','#de8a5a','#ca562c')
    cols <- RColorBrewer::brewer.pal(n = 8, name = 'Blues')
    g <- ggplot(data = map,
         aes(x = long,
             y = lat,
             group = group)) +
    geom_polygon(aes(fill = var),
                 lwd = 0.3,
                 color = line_color) +
      scale_fill_gradientn(name = '',
                           colours = cols)
    # scale_fill_viridis(name = '' ,option = 'magma',
    #                    direction = -1) 
  }
  
  # Add names?
  if(add_names){
    centy <- centroids %>% left_join(right)
    if(add_values){
      centy$label <- paste0(centy$ccaa, '\n(', round(centy$var, digits = 2), ')')
    } else {
      centy$label <- centy$ccaa
    }

    g <- g +
      geom_text(data = centy,
                aes(x = long,
                    y = lat,
                    label = label,
                    group = ccaa),
                alpha = 0.7,
                size = text_size)
  }
  
  g +
    theme_map() +
    labs(subtitle = paste0('Data as of ', max(data$date))) +
    theme(legend.position = 'right')
  
}


make_dot_map <- function(var = 'deaths',
                     date = NULL,
                     pop = FALSE,
                     pop_factor = 100,
                     point_factor = 1,
                     points = FALSE,
                     point_color = 'darkred',
                     point_size = 0.6,
                     point_alpha = 0.5){
  
  
  if(is.null(date)){
    the_date <- max(esp_df$date)
  } else {
    the_date <- date
  }
    right <- esp_df[esp_df$date == the_date,c('ccaa', var)]
   names(right)[ncol(right)] <- 'var'
  if(pop){
    right <- left_join(right, esp_pop)
    right$var <- right$var / right$pop * pop_factor
  }
  map_data <- esp1@data %>%
    left_join(names_df) %>%
      left_join(right)
  map_data$var <- map_data$var / point_factor
  out_list <- list()
  for(i in 1:nrow(map_data)){
    sub_data <- map_data[i,]
    this_value = round(sub_data$var)

    if(this_value >= 1){
      this_ccaa = sub_data$ccaa
      # get some points
      sub_points <- random_points %>% filter(ccaa == this_ccaa)
      sampled_points <- sub_points %>% dplyr::sample_n(this_value)
      out_list[[i]] <- sampled_points
    }
  }
  the_points <- bind_rows(out_list)
  
  g <- ggplot() +
    geom_polygon(data = espf,
         aes(x = long,
             y = lat,
             group = group),
         fill = 'white',
         color = 'black',
         lwd = 0.4, alpha = 0.8) +
    geom_point(data = the_points,
               aes(x = long,
                   y = lat),
               color = point_color,
               size = point_size,
               alpha = point_alpha)
  g +
    theme_map() +
    labs(subtitle = paste0('Data as of ', max(esp_df$date)))
  
}

```

### Deaths

#### Absolute number of deaths: points

```{r}
make_map(var = 'deaths',
       points = T) +
  labs(title = 'Number of deaths')
if(!dir.exists('high_quality')){
  dir.create('high_quality')
}
ggsave('high_quality/points_death_absolute.pdf', width = 8, height = 6)
```

#### Absolute number of deaths: choropleth

```{r}
make_map(var = 'deaths',
         line_color = 'darkgrey',
       points = F) +
  labs(title = 'Number of deaths')
ggsave('high_quality/polygons_death_absolute.pdf', width = 8, height = 6)

```


#### Number of deaths adjusted by population: points

```{r}
make_map(var = 'deaths', pop = TRUE, points = T) +
  labs(title = 'Number of deaths per 100,000')
ggsave('high_quality/points_death_adjusted.pdf', width = 8, height = 6)

```


#### Number of deaths adjusted by population: polygons

```{r}
make_map(var = 'deaths', pop = TRUE, points = F, line_color = 'darkgrey') +
  labs(title = 'Number of deaths per 100,000')
ggsave('high_quality/polygons_death_adjusted.pdf', width = 8, height = 6)

```

#### Number of deaths: 1 dot per death

```{r}
make_dot_map(var = 'deaths', point_size = 0.15) +
  labs(title = 'COVID-19 deaths: 1 point = 1 death\nImportant: points are random within each CCAA; do not reflect exact location')
ggsave('high_quality/point_per_death.pdf', width = 8, height = 6)

```


### Cases

#### Absolute number of cases: points

```{r}
make_map(var = 'cases',
       points = T) +
  labs(title = 'Number of confirmed cases')
ggsave('high_quality/points_case_absolute.pdf', width = 8, height = 6)

```

#### Absolute number of cases: choropleth

```{r}
make_map(var = 'cases',
         line_color = 'darkgrey',
       points = F) +
  labs(title = 'Number of confirmed cases')
ggsave('high_quality/polygon_case_absolute.pdf', width = 8, height = 6)

```


#### Number of cases adjusted by population: points

```{r}
make_map(var = 'cases', pop = TRUE, points = T) +
  labs(title = 'Number of confirmed cases per 100,000')
ggsave('high_quality/points_case_adjusted.pdf', width = 8, height = 6)

```


#### Number of cases adjusted by population: polygons

```{r}
make_map(var = 'cases', pop = TRUE, points = F,
         line_color = 'darkgrey') +
  labs(title = 'Number of confirmed cases per 100,000')
ggsave('high_quality/polygons_case_adjusted.pdf', width = 8, height = 6)

```


#### Number of cases adjusted by population: polygons: JUST 7 MOST RECENT DAYS

```{r}
keep_dates <- as.Date(seq((max(esp_df$date)-6), max(esp_df$date), by = 1))
data <- esp_df %>% filter(date %in% keep_dates)
make_map(var = 'cases', data = data, pop = TRUE, points = F,
         line_color = 'darkgrey') +
  labs(title = 'Number of confirmed cases per 100,000',
       caption = paste0('Cumulative incidence over the week from ',
                        min(keep_dates), ' through ', 
                        max(keep_dates)))
ggsave('high_quality/polygons_case_adjusted_week.pdf', width = 8, height = 6)

```

#### Number of cases: points

```{r}
make_dot_map(var = 'cases',
             point_size = 0.05, point_alpha = 0.5, point_factor = 10) +
  labs(title = 'COVID-19 cases: 1 point = 10 cases\nImportant: points are random within each CCAA; do not reflect exact location')
ggsave('high_quality/point_per_10_cases.pdf', width = 8, height = 6)

```

# Numbers for manuscript

Calculate cumulative incidence per 1.000.000 per week

```{r}
keep_dates <- as.Date(seq((max(esp_df$date)-6), max(esp_df$date), by = 1))
x <- esp_df %>%
  filter(date %in% keep_dates) %>%
  group_by(ccaa) %>%
  summarise(incident_cases_over_7_days = sum(cases_non_cum)) %>%
  ungroup %>%
  left_join(esp_pop) %>%
  mutate(incident_cases_over_7_days_per_10e4 = (incident_cases_over_7_days / pop) * 10e4) %>%
  arrange(desc(incident_cases_over_7_days_per_10e4))
kable(x)
```

